\input{preamble.tex}
\usepackage{tikz-cd}
\hfuzz=5pt

\title{\huge{Groups, Rings and Fields}}
\author{\LARGE{Thobias Høivik}}
\date{\Large{Spring 2026}}

\begin{document}
\maketitle

\newpage
\tableofcontents

\newpage
\section{Groups}

\begin{defn}[Group]
\label{defn:group}
A group is a set $S$ together with a binary operation 
$\circ$ such that the following properties hold: 

\begin{enumerate}
    \item Identity: There exists an element $e \in S$, satisfying
        $$ 
            e \circ a = a \circ e = a 
        $$ 
        for every $a \in S$.

    \item Inverses: For every $a \in S$ there exists $b \in S$ such 
        that 
        $$ 
            a \circ b = b \circ a = e, \text{ the identity element}
        $$ 
        We usually denote this element as $a^{-1}$ or $-a$, depending
        on context.

    \item Associativity: For any $a,b,c \in S$, we require 
        $$ 
            a \circ (b \circ c) = (a \circ b) \circ c
        $$ 
\end{enumerate}

A group is then the tupple $(S,\circ)$. We will often just write 
the set to refer to the group, e.g. refering to the group 
$(\mathbb Z, +)$ as just $\mathbb Z$.
    
\end{defn}

\subsection{Basic Examples}

\begin{ex}[Integers under addition]
The set of integers $\mathbb{Z}$ with the operation $+$ forms a group:
\begin{itemize}
    \item Identity: $0$ since $0 + n = n + 0 = n$ for all $n \in \mathbb{Z}$.
    \item Inverses: For $n \in \mathbb{Z}$, the inverse is $-n$.
    \item Associativity: Addition is associative.
\end{itemize}
Hence $(\mathbb{Z}, +)$ is a group.
\end{ex}

\begin{ex}[Non-example: Natural numbers under addition]
The set $\mathbb{N}$ under $+$ is \emph{not} a group since
there is no inverse for $n > 0$.
\end{ex}

\subsection{Basic Properties}

\begin{thm}[Uniqueness of identity]
\label{thm:identity-unique}
The identity element in a group is unique.
\end{thm}

\begin{proof}
Suppose $e$ and $e'$ are both identities. Then
$$
e = e \circ e' = e',
$$
so the identity is unique.
\end{proof}

\begin{thm}[Uniqueness of inverses]
\label{thm:inverse-unique}
Each element in a group has a unique inverse.
\end{thm}

\begin{proof}
Suppose $b$ and $c$ are inverses of $a$. Then
$$
b = b \circ e = b \circ (a \circ c) = (b \circ a) \circ c = e \circ c = c.
$$
\end{proof}

\subsection{Example Problem}

\begin{prob}
Determine whether the set 
$$
G = \{1, -1, i, -i\} \subset \mathbb{C}
$$ 
with multiplication is a group.
\end{prob}

\begin{proof}[Solution]
We check the group axioms:
\begin{enumerate}
    \item \textbf{Closure:} Multiplying any two elements of $G$ yields another element in $G$. True.
    \item \textbf{Identity:} The element $1$ acts as identity. True.
    \item \textbf{Inverses:} Each element has an inverse in $G$: $1^{-1}=1$, $(-1)^{-1}=-1$, $i^{-1}=-i$, $(-i)^{-1}=i$. True.
    \item \textbf{Associativity:} Multiplication of complex numbers is associative. True.
\end{enumerate}
Hence $(G, \cdot)$ is a group, in fact it is an abelian group 
which we shall describe below in \cref{defn:abelian_group}. 
\end{proof}

\newpage
\subsection{Abelian Groups}

\begin{defn}[Abelian Group]
\label{defn:abelian_group}
A group $(G, \circ)$ is \emph{abelian} (or commutative) if
$$
a \circ b = b \circ a \quad \forall a,b \in G.
$$
\end{defn}

\begin{ex}
The group $(\mathbb{Z}, +)$ is abelian because $m+n = n+m$.
\end{ex}

\newpage 
\section{Important Groups}
First let's define some often used notation. 

For $g \in G$, define: 
\begin{itemize}
    \item $g^n = gg\ldots g(n > 0)$ 
    \item $g^0 = e$
    \item $g^{-n} = (g^{-1})^n$
\end{itemize} 
and recognize the following identities (which are provable 
by induction): 
$$ 
    g^m g^n = g^{m+n}, \quad (g^m)^n = g^{mn}
$$ 

\begin{defn}[Roots of Unity]
    \label{defn:roots_of_unity}
    $$ 
        U_n = \{e^{2\pi ik/n} : 0 \leq k < n\}
    $$ 
    are called the n-th roots of unity.
\end{defn}

\begin{defn}
    An subset $\{g_1,\ldots,g_k\} \subseteq G$ is a generating set 
    if every $g \in G$ can be expressed as 
    $$ 
        g_1^{n_1} \cdots g_k^{n_k}
    $$ 

    If a singleton $g'$ alone generates a group then it is called 
    a generator of $G$ and $G$ is said to be cyclic. 
\end{defn}

The $n$-th roots of unity are finite, with exactly $n$ elements. 
They are cyclic, with generator $e^{2i\pi/n}$.

\begin{defn}[Symmetric Group]
    \label{defn:symmetric_group}
    The symmetric group $S_n$ is the set of all permutations of 
    $n$ elements under composition. 
    In other words it is the set of all $\sigma : [n] \to [n]$ with 
    composition as its operation. 
\end{defn}

The symmetric group is non-abelian for $n \geq 3$. The identity 
corresponds to the identity permutation (doing nothing) and inverses 
are undoing a permutation. 
We have the following notation for permutations in the symmetric Group
(assume $n = 3$ for this example): 
$$ 
    \begin{pmatrix}
        1 & 2 & 3 \\ 
        2 & 3 & 1
    \end{pmatrix}
$$ 
corresponds to $1 \mapsto 2, 2 \mapsto 3, 3 \mapsto 1$. 

Another way to write the same permutation is: 
$$ 
    \left(1 \: 2 \: 3\right)
$$ 

\newpage 
\section{Subgroups}
\begin{defn}[Subgroup]
    \label{defn:subgroup}
    Let $G$ be a group under $\star$.
    A subgroup $H \leq G$ is a subset $H \subseteq G$ that is itself 
    a group under $\star$.
\end{defn}

Having to mechanically check if $H \subseteq G$ satisfies all the 
required axioms can be a bit tedious so next we introduce a 
powerful theorem which let's us easily determine whether some 
subset is a subgroup or not. 

\begin{thm}[Subgroup Test]
    \label{thm:subgroup_test}
    A nonempty subset $H \subseteq G$ is a subgroup if and only 
    if 
    $$ 
        \forall x,y \in H, \; xy^{-1} \in H.
    $$ 
\end{thm}
\begin{proof}[Proof Idea]
    If $H$ is a subgroup the required conditions follows. 

    Conversely: 
    \begin{itemize}
        \item Nonempty so some $h \in H$.
        \item $hh^{-1} = e \in H$. 
        \item Closure under $xy^{-1}$ will give inverses and 
            closure. 
    \end{itemize}
\end{proof}

We get some nice corollaries from this, as well as a nice and tidy 
way to prove whether or not a subset is a subgroup. 

\begin{cor}
    The intersection of any collection of subgroups is a subgroup. 

    Given any subset $S \subseteq G$, there is a smallest subgroup 
    containing it. 
\end{cor}

\begin{defn}[Cyclic Subgroups]
    Let $G$ be a group. 

    For any $g \in G$, define: 
    $$ 
        \langle g \rangle  = \{g^n : n \in \mathbb Z\}.
    $$ 

    $\langle g \rangle$ is called the cyclic subgroup generated 
    by $g$. 
\end{defn}

It is very straightforward to show that the cyclic subgroup generated 
by some $g \in G$ is indeed a subgroup with the use of 
\cref{thm:subgroup_test}.

On page \pageref{defn:roots_of_unity} we say that a group is cyclic 
if it has a signle element which generates it, but now we can simply
say that a group $G$ is cyclic if $G = \langle g \rangle$ for some 
$g \in G$.

\begin{thm}
    Suppose a group $G$ is cyclic, i.e. $G = \langle g \rangle$ 
    for some $g \in G$. 

    Then it is isomorphic to one of the following: 
    \begin{itemize}
        \item $(\mathbb Z, +)$, or 
        \item $(\mathbb Z/n\mathbb Z, +)$ for some $n \geq 1$. 
    \end{itemize}
\end{thm}

\subsection{Structure of Cyclic Subgroups}
\begin{defn}[Order]
    The order of an element $g \in G$ is 
    $$ 
        \text{ord}(g) = |\langle g \rangle|
    $$ 
\end{defn}

Furthermore, if $\text{ord}(g) = n < \infty$, then $g^n = e$ and 
$n$ is minimal, otherwise we say $g$ has infinite order.  

If $\text{ord}(g) = n$, then: 
$$ 
    \langle g \rangle = \{e, g, g^2, \ldots, g^{n-1}\}
$$ 

Moreover: 
\begin{itemize}
    \item $g^k = e \Leftrightarrow n \mid k$
    \item $\langle g^k \rangle = \langle g \rangle \Leftrightarrow 
        \gcd(k,n) = 1$
\end{itemize}

\begin{thm}
    Every subgroup of a cyclic group is cyclic. 

    More precisely: 
    \begin{itemize}
        \item Subgroups of $\mathbb Z$ are exactly $n\mathbb Z$.
        \item Subgroups of $\mathbb Z/n\mathbb Z$ correspond 
            to the divisors of $n$.
    \end{itemize}
\end{thm}

With all of this machinery we can approach group theory with geometric
and algebraic intuition. Cyclic groups are like repeated motion, 
finite cyclic groups corresponding to rotations by rational angles 
and infinite ones corresponding to translations. 

\newpage 
\section{Cosets and Lagrange's Theorem}
\begin{defn}[Left and Right Cosets]
    Let $H \leq G$, and let $g \in G$. 

    The left- and right cosets are: 
    $$ 
        gH = \{gh : h \in H\} \quad Hg = \{hg : h \in H\}
    $$ 
    respectively.
\end{defn}

A key intuition is that a coset is like a copy of $H$ shifted 
by $g$. 
We may also recognize that the cosets are the same size as $H$ by 
identifying $h \mapsto gh$ to be a bijection and so forth. 

The next natural question is when are two cosets equal? 
Like with showing when a subset is a subgroup we have a little trick 
for this problem. 

\begin{thm}
    Let $G$ be a group. 

    For $g_1, g_2 \in G$, the following are equivalent: 
    \begin{align*}
        g_1 H &= g_2 H \\ 
              &\Updownarrow \\ 
        g_2^{-1}g_1 &\in H \\ 
                    &\Updownarrow \\ 
        g_1 &\in g_2 H 
    \end{align*}
\end{thm}

\begin{thm}
    \label{thm:coset_partition}
    The set of all left cosets of $H \leq G$ forms a 
    partition of $G$. 

    That is, 
    \begin{itemize}
        \item Every element of $G$ is in exactly on coset. 
        \item Two cosets are either the same or disjoint. 
    \end{itemize}
\end{thm}

\begin{defn}
    Let $G$ be a group with $H \leq G$. 

    The index of $H$ in $G$, written $[G:H]$, is the number of 
    left cosets of $H$. 
\end{defn}

If $G$ is finite we get $|G| = [G:H] \cdot |H|$.

\newpage 
\begin{thm}[Lagrange's Theorem]
    \label{thm:lagrange} 
    Let $G$ be a group with $H \leq G$. 

    Then: 
    $$ 
        |H| \bigm\vert |G|
    $$ 
\end{thm}
\begin{proof}
    Let $G$ be a group with $H \leq G$. 

    Recall that by \cref{thm:coset_partition} we have that the 
    set of left cosets of $H$, $g_1 H, g_2 H, \ldots, g_k H$ 
    (this set is finite since $G$ is finite) are 
    pairwise disjoint, satisfying 
    $$ 
        G = \bigcup_{i=1}^k g_i H
    $$ 
    meaning 
    $$ 
        |G| = \displaystyle\sum_{i=1}^{k}|g_i H|
    $$ 

    Recall also that $|H| = |g_i H|, \, \forall i \in [k]$. 
    Thus 
    $$ 
        |G| = \displaystyle\sum_{i=1}^{k}|H| = k|H|
    $$ 

    In other words $|G|$ is $|H|$ times some integer $k$, therefore 
    $|H|$ divides $|G|$. 
\end{proof}


\Cref{thm:lagrange} has the following immediate consequences: 
\begin{itemize}
    \item The order of any element divides $|G|$. 
    \item $g^{|G|} = e$ for every $g \in G$. 
\end{itemize}

Beware that it does not follow from Lagrange's Theorem that there 
exists a subgroup with the order of a divisor of $|G|$ for every 
divisor of $G$. The alternating group of order $12$, $A_{4}$, 
has no subgroup of order $6$ for example.

\begin{prop}
    A group of order $p$ where $p$ is prime, is cyclic. 
\end{prop}
\begin{proof}[Proof Sketch]
    Let $G$ be a group with $|G| = p$ (prime). 

    Let $g \in G \setminus \{1\}$. Consider it's generated subgroup 
    $\langle g \rangle$. By Lagrange's Theorem the order of this 
    subgroup divides $p$ so $|\langle g \rangle|$ is $1$ or $p$, but 
    it can't be $1$ as $g \neq 1$ so it's order is $p$, i.e. 
    $g$ generates the entirety of $G$. 
\end{proof}

\subsection{Problems relating to subgroups}
\begin{prob}
    Let $G$ be a group and $g \in G$ with  $\text{ord}(g) = 12$. 

    \begin{itemize}
        \item List all distinct subgroups of $\langle g \rangle$.
        \item For which integers $k$ does $g^k$ generate $\langle g \rangle$?
    \end{itemize}
\end{prob}
\begin{proof}[Solution]
    Recalling that $\langle g \rangle = \{g^k : k \in \mathbb Z\}$, 
    we want to find all the distinct subgroups of $\langle g \rangle$. 
    In other words, which subsets of $\langle g \rangle$ are also 
    groups with respect to the operation of $G$? 

    $\text{ord}(g) = |\langle g \rangle| = 12$ hence $\langle g \rangle = \{g^0, g^1, \ldots, g^{11}\}$.
    We know that there are two trivial subgroups of $\langle g \rangle$, namely $\{g^0\}$ and $\langle g \rangle$. 
    Otherwise, we know that every cyclic subgroub of a cyclic group (which $\langle g\rangle$ is) is itself 
    cyclic, so we can find the subgroups by looking at the generated groups of each element 
    of $\langle g \rangle$. Otherwise, we could use the fact that 
    $\langle g\rangle $ is isomorphic to $\mathbb Z/12\mathbb Z$ together with realizing that the number 
    of subgroups is then the same as the number of positive divisors of $12$. 

    Using the first approach we get: 
    \begin{align*}
        \langle g^2 \rangle &= \{g^n : n = 2k, \: k \in \mathbb Z\} \\ 
                            &= \{e, g^2, \ldots, g^10\} = \langle g^10 \rangle \\ 
        \langle g^3 \rangle &= \{e, g^3, g^6, g^9\} = \langle g^9 \rangle \\ 
        \langle g^4 \rangle &= \{e, g^4, g^8\} = \langle g^8 \rangle \\ 
        \langle g^6 \rangle &= \{e, g^6\} \\ 
    \end{align*}
    
    Those integers which are coprime to $12$, like $1,5,7,11$, will generate $\langle g \rangle$. 
    Thus we have $6$ distinct subgroups, exactly the same as the number of 
    positive divisors of $12$, as expected. 
\end{proof}

\begin{prob}
    Let $G$ be any group $H = \langle g \rangle$ where $\text{ord}(g) < \infty$. 

    Prove that if $g^m \in H$, then $\langle g^m \rangle \leq H$. 
\end{prob}
\begin{proof}
    Assume, for contradiction, that $g^m \in H$, but $\langle g^m \rangle$ 
    is not a subgroup of $H$. 

    Let $s = m \mod \text{ord}(g)$. 
    Then $\langle g^m \rangle = \{g^{sk} : k \in \mathbb Z\} \neq \emptyset$.  
    It is clear that this is a subset of $\langle g \rangle = H$, since 
    $\langle g \rangle$ is closed under the group operation and 
    $g^m \in H$. 
    Then the only way our assumption is true is if 
    there exists some $x,y \in \langle g^m \rangle$ such that 
    $xy^{-1} \not\in \langle g^m \rangle$. 

    Every element $x,y \in \langle g^m \rangle$ is of the form 
    $x = g^{sn}, \, y = g^{st}$ for some $n,l \in \mathbb N$. 
    By assumption, 
    $xy^{-1} = g^{s(n-t)} \not\in \langle g^m \rangle$ 
    which would necessarily mean that $n-t \not \in \mathbb Z$, 
    impossible. 
\end{proof}
The above argument can be made more rigorous 
by considering the isomorphism between $\langle g^m \rangle$ and 
$\mathbb Z/s\mathbb Z$, or by proving the claim 
directly (I am an idiot, and this was a bad approach). 

\newpage 
\section{Normal subgroups}
\begin{defn}[Normal subgroup]
    \label{defn:normal_subgroup}
    A subgroup $H \leq G$ is normal, written $H \trianglelefteq G$, if:
    $$ 
        gH = Hg, \: \forall g \in G
    $$ 
    with the equivalent characterization:
    $$ 
        H \trianglelefteq G \Leftrightarrow gHg^{-1} = H, \: \forall g \in G
    $$ 
\end{defn}

A few things to tick off immediately; 
\begin{enumerate}
    \item Every subgroup of an abelian group is normal. 
    \item \{e\} and $G$ are always normal. 
    \item The alternating group $A_n \trianglelefteq S_n$. 
    \item Subgroups of index $2$ are always normal. 
\end{enumerate}

If you consider the map (conjugation)
$$ 
    x \mapsto gxg^{-1},
$$ 
which constitutes an automorphism of $G$ we can say 
the following. 

Normal subgroups are precisely those subgroups which are 
fixed under every conjugation. 

\begin{defn}[Quotient Group]
    \label{defn:quotient_group}
    If $H \trianglelefteq G$, define: 
    $$ 
        G/H := \{gH : g \in G\}
    $$ 
    with an operation we'll call multiplication by: 
    $$ 
        (gH)(kH) = (gk)H
    $$ 
\end{defn}

\begin{thm}
    The quotient group with multiplication, as defined above, 
    is a well-defined group. 
\end{thm}

\newpage 
\begin{defn}
    We define the \textbf{canonical projection} as 
    $$ 
        \pi : G \to G/H, \: \pi(g) = gH.
    $$ 
\end{defn}
Notably, the projection $\pi$ constitutes 
a homomorphism with kernel $\ker \pi = H$.
Notice that this claim of the kernel being $H$ itself 
should be immediately obvious as $hH$ for some 
$h \in H$ is itself $H$ as it's closed. 

\newpage 
\section{Exercises for Week 1}
\begin{prob}[29, p.48]
    Show that if $G$ is a finite group with identity $e$ and 
    an even number of elements, then there is 
    $a \neq e$ in the group such that $a * a = e$.
\end{prob}
\begin{proof}
    We are being asked to show that in a finite group with even 
    order, there exists some element other than the identity which 
    is it's own inverse. 

    Recall uniqueness of inverses and that for $g \in G$, 
    $(g^{-1})^{-1} = g$. Thus for every element there exists one 
    and only one element which is it's inverse. $e^{-1} = e$ is 
    covered. In $G \setminus \{e\}$ there are an odd number of 
    elements. Each of these have exactly one unique inverse meaning 
    that there is one element "left" over which then necessarily 
    has itself as it's inverse. 
\end{proof}

\begin{prob}[33, p.49]
    Let $G$ be an abelian group and let $c^n = c * c * \cdots * c$ 
    for $n$ factors c, where $c \in G$ and $n \in \mathbb Z^+$. 
    Give a proof by mathematical induction that $(a*b)^n = a^n * b^n$ 
    for all $a,b \in G$. 
\end{prob}
\begin{proof}
    Let $a,b \in G$. The base case where $n = 1$ is trivial so 
    we proceed to our assumption that for any $k \in \mathbb Z^+$ we 
    have $(a*b)^k = a^k * b^k$. 

    Let $n = k + 1$. 
    \begin{align*}
        (a*b)^n &= (a*b)^{k+1} \\ 
                &= (a*b)^k * (a*b) \\  
                &= a^k * b^k * (a*b) & \text{by hypothesis} \\ 
                &= a^k * (b^k * a) * b & \text{by associativity} \\ 
                &= a^k * (a * b^k) * b & \text{by commutativity} \\ 
                &= (a^k * a) * (b^k * b) \\ 
                &= (a^{k+1}) * (b^{k+1}) \\ 
                &= a^n * b^n
    \end{align*}
     
    By the principle of mathematical induction we have that 
    $$  
        (a*b)^n = a^n * b^n, \: \forall n \in \mathbb Z^+
    $$ 
\end{proof}

\newpage 
\begin{prob}[35, p. 49]
    Show that if $(a*b)^2 = a^2 * b^2$ for $a,b$ in a group, then 
    $a*b = b*a$.  
\end{prob}
\begin{proof}
    \begin{align*}
        (a*b)^2 &= (a*b)*(a*b) \\ 
        a^2 * b^2 &= (a*a) * (b*b) \\ 
        \Rightarrow (a*b) * (a*b) &= (a*a) * (b*b) \\ 
                    a * (b*a) * b &= a*(a*b)*b
    \end{align*}
    
    By one application each of left- and right cancellation we have 
    $$ 
        a*b = b*a
    $$ 
\end{proof}

\begin{prob}[41, p. 49]
    Let $G$ be a group and fix $g \in G$. Show that the map 
    $i_g$, s.t. $i_g(x) = gxg^{-1}$ for $x \in G$. Show that 
    $i_g$ is an automorphism on $G$.
\end{prob}
\begin{proof}
    Let $i_g : G \to G$ be defined as above. We need to show that 
    it satisfies all the properties of an isomorphism. 

    It is clear, through left- and right cancellation, that 
    $i_g$ is injective. 

    Notice also that for any $x \in G$ we have the pre-image 
    $g^{-1}xg$ since $i_g(g^{-1}xg) = gg^{-1}xgg^{-1} = x$, hence 
    $i_g$ is surjective. 

    Lastly we verify the homomorphism property: 
    \begin{align*}
        i_g(xy) &= gxyg^{-1} \\ 
                &= gxeyg^{-1} \\ 
                &= gxg^{-1}gyg^{-1} \\ 
                &= i_g(x) i_g(y)
    \end{align*}
    completing the proof. 
\end{proof}

\newpage 
\section{Problems for week of Jan 26}
\begin{prob}[Problem 44 Fraleigh's p.67]
    Let $G = \langle a \rangle$ be a cyclic group isomorphic 
    to a group $G'$. 
    If $\phi : G \to G'$ is an isomorphism, show that 
    for every $x \in G$, $\phi(x)$ is uniquely determined by $\phi(a)$.
    That is, if $\phi$ and $\psi$ constitutes isomorphisms from 
    $G \to G'$ with $\phi(a) = \psi(a)$, then $\phi(x) = \psi(x)$ 
    for every $x \in G$.
\end{prob}
\begin{proof}
    Let $G$ be generated by $a$, isomorphic to $G'$ via 
    $\phi$ and $\psi$, satisfying  
    \[
        \phi(a) = \psi(a)
    \]
    Let $x \in G$. Then $x = a^n$ for some $n \in \mathbb Z$, as 
    $G = \langle a \rangle$. 
    Observe the following:
    \begin{align*}
        \phi(x) &= \phi(a^n) \\ 
                &= \phi(a)^n \\ 
                &= \psi(a)^n \\ 
                &= \psi(a^n) = \psi(x)
    \end{align*}
    This completes the proof.
\end{proof}

\begin{prob}[Problem 6 Fraleigh's p.83]
    Compute $|\sigma|$ for 
    \[
        \sigma = 
        \begin{pmatrix}
            1 & 2 & 3 & 4 & 5 & 6 \\ 
            3 & 1 & 4 & 5 & 6 & 2
        \end{pmatrix}
    \]
\end{prob}
\begin{proof}[Solution]
    Observe that $\sigma$ is the permutation which sends 
    \begin{align*}
        1 &\to 3 \\ 
        2 &\to 1 \\ 
        & \ \vdots \\ 
        6 &\to 2
    \end{align*}
    so $\sigma^2$ is 
    \begin{align*}
        1 &\to 4 \\ 
        2 &\to 3 \\ 
          & \ \vdots \\ 
        6 &\to 1
    \end{align*}
    We want to find $n \in \mathbb N$ such that 
    $\sigma^n = $ the identity permutation. 
    It's easy to see that performing the permutation $6$ times 
    gives $1 \to 1$ which we would require. We also see that 
    $2 \to 2$, $3 \to 3$ and so on. 

    We could also write the permutation in compact form and see 
    that $\sigma$ is a cycle of length $6$, so the order is $6$. 
\end{proof}

\newpage 
\begin{prob}[Problem 7 Fraleigh's p.83]
    Compute $|\tau|$ for 
    \[
        \tau = 
        \begin{pmatrix}
            1 & 2 & 3 & 4 & 5 & 6 \\ 
            2 & 4 & 1 & 3 & 6 & 5 
        \end{pmatrix}
    \]
\end{prob}
\begin{proof}[Solution]
    Note that $\tau$ in compact notation is 
    \[
        \tau = (1 \ 2 \ 4 \ 3)(5 \ 6)
    \]
    so $\tau$ is the composition of a $2$- and $4$-cycle. 
    \[
        |\tau| = \text{lcm}(4, 2) = 4
    \]
\end{proof}
\begin{prob}[Problem 49 Fraleigh's p.86]
    If $A$ is a set, then a subgroup $H \leq S_A$ is transitive 
    on $A$ if for each $a,b \in A$ there exists $\sigma \in H$ 
    such that $\sigma(a)=b$. Show that if $A$ is a nonempty finite 
    set, then there exists a finite cylic subgroup $H$ of $S_A$ 
    with $|H| = |A|$ that is transitive on $A$.
\end{prob}
\begin{proof}
    Let $A = \{a_1, a_2, \ldots, a_n\}$ with $S_A$ being the 
    permutations on $A$. Let $\sigma \in S_A$ be the permutation which 
    sends $a_i \to a_{i+1}$ (take $i$ modulo $n$). It is clear that 
    $\sigma^k$ is the permutation which sends $a_i$ to $a_{i+k}$ and 
    furthermore $|\sigma| = n = |A|$. It is not hard to see 
    that $\langle\sigma\rangle$ forms a subgroup 
    where each permutation shifts 
    all elements of $A$. 
    
    It is clear that for any $a_i, a_j \in A$ we can find 
    a permutation which sends one to the other. Suppose, without 
    loss of generality, $i < j$. Then 
    $\sigma^{j-i}(a_i) = a_j$, and the inverse of said permutation 
    sends $a_j$ to $a_i$.

    A bit inelegantly written, but I digress. 
\end{proof}
\begin{prob}[Problem 1 Fraleigh's p.94]
    Find the orbit of 
    \[
        \sigma = 
        \begin{pmatrix}
            1 & 2 & 3 & 4 & 5 & 6 \\ 
            5 & 1 & 3 & 6 & 2 & 4
        \end{pmatrix}
    \]
\end{prob}
\begin{proof}[Solution]
    We are looking at $\sigma \in S_6$ acting on $[6]$ (I assume)
    so we want to find 
    \[
        \left\{ 
            \text{Orb}_\sigma(n) = \{\sigma^k(x) : 
            k \in \mathbb Z\} 
            \mid n = 1, 2, \ldots, 6
        \right\}
    \]

    Let us rewrite $\sigma$ more compactly. 
    \[
        1 \to 5 \to 2 \to 1 
    \]
    \[
        3 \to 3 
    \]
    \[
        4 \to 6 \to 4
    \]
    so 
    \[
        \sigma = (1 \ 5 \ 2)(4 \ 6)
    \]
    
    Thus we have the following orbits: 
    \begin{align*}
        \text{Orb}_\sigma(1) = \text{Orb}_\sigma(2) = 
        \text{Orb}_\sigma(5) = \{1,2,5\} \\ 
        \text{Orb}_\sigma(4) = \text{Orb}_\sigma(6) = \{4,6\} \\
        \text{Orb}_\sigma(3) = \{3\} \\ 
    \end{align*}
\end{proof}

\begin{prob}[Problem 7 Fraleigh's p.94]
    Compute: 
    \[
        (1 \ 4 \ 5)(7 \ 8)(2 \ 5 \ 7)
    \]
\end{prob}
\begin{proof}[Solution]
    \begin{align*}
        (1 \ 4 \ 5)(7 \ 8)(2 \ 5 \ 7)
        &= (1 \ 4 \ 5) \circ ((7 \ 8)(2 \ 5 \ 7)) \\ 
        &= (1 \ 4 \ 5) (2 \ 5 \ 8 \ 7) \\ 
        &= (2 \ 1 \ 4 \ 5 \ 8 \ 7)
    \end{align*}
\end{proof}
\begin{prob}[Problem 29 Fraleigh's p.96]
    Show that for every subgroup $H$ of $S_n$ for $n \geq 2$, 
    either all permutations in $H$ are even or exactly half are. 
\end{prob}
\begin{proof}
    Recall that an even permutation is a permutation which can be 
    expressed as an even number of transpositions (swaps of 
    $2$ elements). 

    Let $\text{sgn}(\sigma) = 1$ if $\sigma$ is even, 
    $\text{sgn}(\sigma) = -1$ otherwise for all $\sigma \in S_n$.  
    $\text{sgn}:S_n \to \{1,-1\}$ is a homomorphism.  
    Restrict this map to $H$. Then, the image: 
    \[
        \text{sgn}[H] = 
        \begin{cases}
            \{+1\} \\ 
            \{+1,-1\}
        \end{cases}
    \]
    In the first case it is clear that every element of $H$ must 
    be even. 

    In the other case where $\text{sgn}$ is surjective, the 
    kernel 
    \[
        \ker(\text{sgn}[H]) = H \cap A_n,
    \]
    the set of even permutations in $H$. 
    By the first isomorphism theorem, 
    \[
        H/(H\cap A_n) \cong \{+1, -1\}
    \]
    so 
    \[
        [H:H\cap A_n] = 2
    \]
    hence 
    \[
        |H \cap A_n| = \frac{1}{2} H
    \]
    i.e. half of $H$ is even. 
\end{proof}

\newpage 
\section{External- and Internal Direct Products}
\begin{defn}
    Let $H,K \leq G$, 
    \begin{align*}
        G' = H \times K = \{(h,k) \bigm| h \in H, \ k \in K\} \\ 
        G = H \oplus K = \{hk \bigm| h \in H, \ k \in K\}
    \end{align*}
    The internal direct product $H \oplus K$ presupposes that 
    $H, K \unlhd G$, and $H \cap K = \{e\}$.
\end{defn}
We really want to be able to say that $H \times K \cong H \oplus K$.

\begin{ex}
    Consider the dihedral group $D_3 = \{1,x,x^2, y, xy, x^y\}$. 
    Let $H = \langle x \rangle$ and $K = \langle y \rangle$.
    It's easy to see that 
    $H \cap K = \{e\} = \{1\}$, and $H \unlhd D_3$. 
    However $xK = \{x, xy\}$, but $Kx = \{x,yx\} = \{x, x^2y\}$. 
    Hence 
    \[
        D_3 \neq H \oplus K
    \]
\end{ex}

\begin{lem}
    $H,K \unlhd G$ and $H \cap K = \{e\}$. 
    \begin{enumerate}
        \item if $ab = a'b'$ where $a,a' \in H$ and $b,b' \in K$ 
            then $a = a'$ and $b = b'$.
        \item if $a \in H$ and $b \in K $ then $ab = ba$.
    \end{enumerate}
\end{lem}

\begin{thm}
    Assuming $H,K$ normal subgroups of $G$ with trivial intersection, 
    then 
    \[
        H \times K \cong H \oplus K
    \]
\end{thm}
\begin{proof}
Define a map
\[
\varphi : H \times K \to HK, \qquad \varphi(h,k) = hk.
\]

This map is well-defined since \(h \in H\), \(k \in K\) implies \(hk \in HK\).

Define a second map
\[
\psi : HK \to H \times K
\]
by sending \(g \in HK\) to the unique pair \((h,k)\) such that \(g = hk\).
Such a decomposition exists since \(HK = G\), and is unique because
\(H \cap K = \{e\}\).

We now verify that these maps are inverse to one another. For all
\((h,k) \in H \times K\),
\[
(\psi \circ \varphi)(h,k) = \psi(hk) = (h,k),
\]
and for all \(g \in HK\),
\[
(\varphi \circ \psi)(g) = g.
\]
Thus \(\varphi\) is bijective with inverse \(\psi\).

Finally, \(\varphi\) is a homomorphism, since for \((h_1,k_1),(h_2,k_2) \in H \times K\),
\begin{align*}
    \varphi((h_1,k_1)(h_2,k_2))
    &= \varphi(h_1h_2,k_1k_2) \\
    &= h_1h_2k_1k_2 \\
    &= h_1k_1h_2k_2 \\
    &= \varphi(h_1,k_1)\varphi(h_2,k_2),
\end{align*}
where we use that elements of \(H\) and \(K\) commute.

Therefore \(\varphi\) is an isomorphism, and \(HK \cong H \times K\).
\end{proof}

\begin{thm}
    If $|x|, |y| < \infty$ and $(x,y) \in G \times H$ then 
    $|(x,y)| = \text{lcm}(|x|,|y|)$.
\end{thm}

\newpage 
\section{Problems from week of 8 Feb}
\begin{prob}[Problem 47 Fraleigh's p.113]
    Let $G$ be an abelian group. Let $H \subseteq G$ consisting of the 
    identity and all elements of $H$ with order $2$. Show that 
    $H \leq G$.
\end{prob}
\begin{proof}
    Let $H = \{e\} \cup \{g \in G \bigm| |g| = 2\} \subseteq G$. 
    Let $x,y \in H$, and for the sake of simplicity assume they 
    are nontrivial, i.e. $x\neq e \neq y$. 
    To show that $H$ constitutes a subgroup of $G$ 
    it suffices to show that $xy^{-1} \in H$ (we already 
    know it is nonempty).

    \begin{align*}
        (xy^{-1})^2 &= xy^{-1}xy^{-1} \\ 
                    &= xxy^{-1}y^{-1} &(G \ \text{abelian}) \\ 
                    &= x^2(y^2)^{-1} \\ 
                    &= e(e) \\ 
                    &= e
    \end{align*}
    Hence $|xy^{-1}| = 2$, so $xy^{-1} \in H$.
    Thus $H$ is a subgroup of $G$.
\end{proof}

\begin{prob}[Problem 52 Fraleigh's p.113]
    Show that a finie abelian group is not cyclic if and only 
    if it contains a subgroup isomorphic to $\mathbb Z_p 
    \times \mathbb Z_p$ for some prime $p$.
\end{prob}
\begin{proof}
    Let $G$ be a finite abelian group. 

    We begin by proving the right-to-left implication. 
    Assume then that there is some $H \leq G$ with 
    $H \cong \mathbb Z_p \times \mathbb Z_p$. 
    Suppose, for a contradiction, 
    that $G = \langle g \rangle$ (cyclic) for some $g \in G$.
    Let $\phi : H \to \mathbb Z_p \times \mathbb Z_p$ be an 
    isomorphism. 

    Recall that as $G$ is cyclic, generated by $g$, we have that  
    there is some $k \in \mathbb Z$, for every $h \in H$, 
    such that $h = k\cdot g$. As $G$ is finite, $k \cdot g = 
    (|G| - k) \cdot g$, so we can safely assume $k \geq 0$ without 
    any loss of generality. 

    Since $\phi$ is bijective we know that, for every  
    $y \in \mathbb Z_p \times \mathbb Z_p$, there is some 
    $h \in H$ such that 
    \[
        \phi(h) = y
    \]
    Using the homomorphism property, and the 
    fact that $h = kg, \ k \in \mathbb N$, we get that: 
    \begin{align*}
        y &= \phi(h) \\ 
          &= \phi(k\cdot g) \\ 
          &= \phi\left(\displaystyle\sum_{i=1}^{k}g\right) \\ 
          &= \displaystyle\sum_{i=1}^{k} \phi(g) \\ 
          &= k \cdot \phi(g)
    \end{align*}
    As $y$ was chosen arbitrarily we now see that 
    $\mathbb Z_p \times \mathbb Z_p = \langle \phi(g) \rangle$, i.e. 
    cyclic. Being cyclic, together with the fact that 
    $|\mathbb Z_p \times \mathbb Z_p| = p^2$, it is the 
    case that there must be some element with order 
    $p^2$, but no such element exists so $\mathbb Z_p \times 
    \mathbb Z_p$ is not cyclic. We have arrived at a contradiction 
    so our assumption that $G$ is cyclic must be false. 

    Now we prove the other direction. 
    We now assume that we have a finite abelian group $G$ which 
    is not cyclic. As $G$ is finite we know that $G$ is finitely 
    generated. Thus by the fundamental theorem of finitely generated
    abelian groups we know that $G$ is isomorphic to a direct product 
    of cyclic groups of the form 
    \[
        \mathbb Z_{(p_1)^{\gamma_1}}
        \times 
        \mathbb Z_{(p_2)^{\gamma_2}}
        \times \cdots \times 
        \mathbb Z_{(p_n)^{\gamma_n}} 
        \times 
        \mathbb Z 
        \times \mathbb Z
        \times \cdots \times 
        \mathbb Z
    \]
    As $G$ is finite, the direct product does not include 
    $\mathbb Z$. Furthermore as $G$ is not cyclic we know that 
    \[
        G \not\cong \mathbb Z_{p^\gamma}
    \]
    Hence $G$ must be the direct product of at least $2$ 
    $\mathbb Z_{(p_i)^{\gamma_i}}$. As $\gamma_i \in \mathbb N 
    \setminus \{0\}$ and distinct from $\gamma_j$ for $j \neq i$ 
    there must be some $\mathbb Z_{(p_i)^{\gamma_l}}$ with 
    $\gamma_l \geq 2$. Notice that this group necessarily 
    has 
    \[
        H = \{0\} \times \cdots \times \{0\} \times 
        \mathbb Z_{p_i} \times \{0\} \times \cdots 
        \times \{0\} \times \mathbb Z_{p_i} \times \{0\} \times 
        \cdots \times \{0\}
    \]
    as a subgroup, which itself is isomorphic to 
    $\mathbb Z_{p_i} \times \mathbb Z_{p_i}$. 
    Then $\phi^{-1}[H]$ is a subgroup of $G$, which is 
    isomorphic to $\mathbb Z_{p_i} \times \mathbb Z_{p_i}$.
\end{proof}

\begin{prob}[Problem 46 Fraleigh's p.135]
    Let a group $G$ be generated by $\{a_i \bigm| i \in I\}$, 
    where $I$ is some indexing set and $a_i \in G$ for all $i \in I$.
    Let $\phi,\mu$ be two homomorphisms from $G$ into a group $G'$, 
    such that $\phi(a_i) = \mu(a_i)$ for every $i \in I$. 
    Prove that $\phi = \mu$.
\end{prob}
\begin{proof}
    Let $G = \{a_i \bigm| i\in I\}$ with homomorphisms 
    $\phi, \mu : G \to G'$ which coincide on $a_i$ for all $i \in I$.
    Let $x \in G$. 
    \begin{align*}
        \phi(x) &= \phi\left(\prod_{i \in I} a_i^{k_i}\right), \: 
        k_i \in \mathbb Z \\ 
                &= \prod_{i\in I} \phi(a_i^{k_i}) \\ 
                &= \prod_{i\in I} 
                \underbrace{\phi(a_i)\cdots 
                \phi(a_i)}_{k_i \; \text{times}} \\ 
                &= \prod_{i\in I} \underbrace{\mu(a_i)\cdots 
                \mu(a_i)}_{k_i \; \text{times}} \\ 
                &= \prod_{i\in I} \mu(a_i^{k_i}) \\ 
                &= \mu\left(\prod_{i \in I}a_i^{k_i}\right) 
                = \mu(x)
    \end{align*}
    $x$ was chosen arbitrarily hence $\phi$ and $\mu$ coincide 
    on all $x \in G$. In other words, $\phi = \mu$.
\end{proof}

\begin{prob}[52 Fraleigh's p.135]
    Let $\phi : G \to G'$ be a homomorphism with kernel $H$ and 
    let $a \in G$. Prove the set equality $\{x \in G \bigm| 
    \phi(x) = \phi(a)\} = Ha$.
\end{prob}
\begin{proof}
    We begin by proving that $Ha \subseteq \{x \in G \bigm| 
    \phi(x) = \phi(a)\}$.

    Suppose $ha \in Ha$. 
    Then $h \in H$ so 
    \[
    \phi(ha) = \underbrace{\phi(h)}_{\in H} \phi(a) 
        = e' \phi(a) = \phi(a)
    \]
    Thus $ha \in \{x \in G \bigm| \phi(x) = \phi(a)\}$. As $x$ 
    was arbitrarily chosen, we conclude that 
    \[ 
        Ha\subseteq \{x \in G \bigm| \phi(x) 
        = \phi(a)\}
    \]
    Let $x \in \{x \in G \bigm| \phi(x) = \phi(a)\}$. 
    Then 
    \begin{align*}
        \phi(x) &= \phi(a) \\ 
        \phi(x) \phi(a)^{-1} &= e' \\ 
        \phi(x a^{-1}) &= e'
    \end{align*}
    hence $xa^{-1} \in H$ which implies $x \in Ha$.
    Once again $x$ was arbitrarily chosen so 
    \[
        \{x \in G \bigm| \phi(x) = \phi(a)\} \subseteq Ha
    \]
    
    Summing up we conclude that 
    \[
        Ha = \{x \in G \bigm| \phi(x) = \phi(a)\}
    \]
\end{proof}

\newpage
\section{Problems week of 16 Feb, Quotient grps, simple grps}
\begin{prob}[Problem 3 Fraleigh's p.142]
    Find the order of the given factor group: 

    \[
        (\mathbb Z_4 \times \mathbb Z_2) / \langle (2,1)\rangle
    \]
\end{prob}
\begin{proof}[Solution]
We first compute the subgroup generated by $(2,1)$:
\[
\langle (2,1) \rangle = \{(0,0),(2,1)\},
\]
since $(2,1)+(2,1)=(0,0)$.

Because $\mathbb Z_4 \times \mathbb Z_2$ is abelian, every subgroup is normal,
so the quotient group is well-defined.

The elements of the quotient
\[
(\mathbb Z_4 \times \mathbb Z_2)/\langle(2,1)\rangle
\]
are the cosets
\[
(x,y)+\langle(2,1)\rangle
= \{(x,y),(x+2,y+1)\},
\quad (x,y)\in \mathbb Z_4\times\mathbb Z_2.
\]

Each coset has exactly $2$ elements. Since
\[
|\mathbb Z_4 \times \mathbb Z_2|=8,
\]
the number of distinct cosets is
\[
\frac{8}{2}=4.
\]

Hence the quotient group has order $4$.
\end{proof}
\begin{prob}[Problem 34 Fraleigh's p.143]
    Show that if a finite group $G$ has exactly one subgroup $H$ 
    of a given order, then $H$ is a normal subgroup of $G$.
\end{prob}
\begin{proof}
    Suppose $G$ is a finite group with only one subgroup $H$ of some 
    order $k \in \mathbb N$. 
    Recall that for the conjugacy of $H$ is still a subgroup, 
    furthermore
    \[
        |gHg^{-1}| = |H| = k, \: g \in G
    \]
    However $H$ is the only subgroup with order $k$ so 
    \[
        gHg^{-1} = H, \: \forall g \in G
    \]
    which implies that $H$ is normal.

\end{proof}
\begin{prob}[Problem 35 Fraleigh's p.143]
    Show that if $H$ and $N$ are subgroups of a group $G$, and 
    $N$ is normal in $G$, then $H \cap N$ is normal in $H$. 
\end{prob}
\begin{proof}
    Let $G$ be a group with $H \leq G$ and $N \trianglelefteq G$.
    We take it for granted that $H \cap N$ is indeed a group 
    (this is not hard to show). 

    Fix some arbitrary $h \in H$.
    We must show that $h(H\cap N)h^{-1} \subseteq H \cap N$. 
    
    Let $x \in H \cap N$. As $x, h \in H$ we ceirtanly have 
    $hxh^{-1} \in H$. As $x \in N$, $N$ normal, we have that 
    for any $h \in H \subseteq G$, $hxh^{-1} \in N$. 

    Hence 
    \[
        hxh^{-1} \in H \cap N
    \]
\end{proof}

\begin{prob}[Problem 3 Fraleigh's p.151]
    Classify the groups according to the fundamental theorem of 
    finitely generated abelian groups.
    \[
        (\mathbb Z_2 \times \mathbb Z_4)/\langle(1,2)\rangle
    \]
    \[
        (\mathbb Z_4 \times \mathbb Z_8)/\langle(1,2)\rangle
    \]
    \[ 
        (\mathbb Z_4 \times \mathbb Z_4 \times \mathbb Z_8)/
        \langle (1,2,4) \rangle
    \]
\end{prob}
\begin{proof}[Solution]
    We begin with 
    $(\mathbb Z_2 \times \mathbb Z_4)/\langle(1,2)\rangle$. 

    \begin{align*}
        \langle (1,2) \rangle &= \{(0,0), (1,2)\} \\ 
        (0,0) \{(0,0), (1,2)\} &= \{(0,0), (1,2)\} = A \\ 
        (1,0) \{(0,0), (1,2)\} &= \{(1,0), (0,2)\} = B \\ 
        (0,1) \{(0,0), (1,2)\} &= \{(0,1), (1,3)\} = C \\ 
        (1,1) \{(0,0), (1,2)\} &= \{(1,1), (0,3)\} = D \\ 
        (0,2) \{(0,0), (1,2)\} &= \{(1,0), (0,2)\} = B \\ 
        (1,2) \{(0,0), (1,2)\} &= \{(1,2), (0,0)\} = A \\ 
        (0,3) \{(0,0), (1,2)\} &= \{(0,3), (1,1)\} = D \\ 
        (1,3) \{(0,0), (1,2)\} &= \{(1,3), (0,1)\} = C \\ 
    \end{align*}
    We get an abelian group with $4$ distinct elements.
    Crucially, we can see that the coset of $(0,1)$ has order 
    $4$ and is thus a generator. 
    We conclude that $(\mathbb Z_2 \times \mathbb Z_4)/
    \langle (1,2)\rangle \cong \mathbb Z_4$.

    Now we look at $(\mathbb Z_4 \times \mathbb Z_8)/
    \langle(1,2)\rangle$.

    Notice that 
    \[
        (\mathbb Z_4 \times \mathbb Z_8) = \langle a,b 
        \bigm| 4a = 0, \: 8b = 0, \: ab = ba\rangle
    \]
    so quotienting with $\langle(1,2)\rangle$ only imposes 
    the extra relation that 
    \[
        a + 2b = 0 \Leftrightarrow a = -2b
    \]
    then $\text{ord}(1,2) = \text{lcm}(4,4) = 4$ so 
    $\langle (1,2)\rangle$ has $4$ elements.  
    $|\mathbb Z_4 \times \mathbb Z_8| = 32$ so the quotient group has 
    \[
        \frac{32}{4} = 8
    \]
    elements. 
    Up to isomorphism there exist $3$ abelian groups of order 
    $8$.
    The relation imposed by $\langle (1,2)\rangle$ changes nothing 
    so we require $8b = 0$. Clearly 
    \[
        (\mathbb Z_4 \times \mathbb Z_8)/\langle (1,2)\rangle 
        \cong \mathbb Z_8
    \]

    Lastly we look at 
    \[
        (\mathbb Z_4 \times \mathbb Z_4 \times \mathbb Z_8)/
        \langle (1,2,4) \rangle
    \]
    Notice that 
    \[
        \mathbb Z_4 \times \mathbb Z_4 \times \mathbb Z_8 
        = \{a,b,c \bigm| 4a = 0, \: 4b = 0, \: 8c = 0, \; 
        ab = ba, \; ac = ca, \; bc = ca\}
    \]
    with 
    \[
        \langle (1,2,4) \rangle
    \]
    imposing $a + 2b + 4c = 0$ which then means that 
    $a = -2b -4c$. The order of $\mathbb Z_4 \times 
    \mathbb Z_4 \times \mathbb Z_8$ is $128$. The order
    of the subgroup is $\text{lcm}(4, 2, 2)$ which is $4$. 
    Thus the order of the quotient group is 
    \[
        \frac{128}{4} = 32 = 2^5
    \]
    There are precisely $7$ abelian groups unique up to isomorphism.
    $b$ has order $4$ and $c$ has $8$ so we conclude that 
    the quotient group is isomorphic to 
    \[
        \mathbb Z_4 \times \mathbb Z_8 
    \]
\end{proof}
\begin{prob}[Problem 34 Fraleigh's p.153]
    Let $G$ be finite, containing a nontrivial subgroup of index $2$
    in $G$. Show that $G$ is not simple. 
\end{prob}
\begin{proof}
    There exists a subgroup $H \leq G$ with 
    $2$ left cosets. Since $H$ is always a left coset of $H$ there 
    must be one more left coset $gH$ for any $g \not\in H$. 
    Since the number of right cosets is the same we 
    have right cosets $H$ and $Hg$. There are only $2$ cosets 
    so the non-$H$ cosets must be the same 
    \[
        gH = Hg
    \]
    Thus left and right cosets coincide. Therefore 
    \[
        H \trianglelefteq G
    \]
    $H$ is nontrivial, and crucially $H \neq G$ (since the index 
    is $2$). Hence $G$ has a nontrivial proper subgroup. 
    In other words, $G$ is not simple. 
\end{proof}

\newpage 
\section{Group Actions}
\begin{defn}
    Let $X$ be a set, and $G$ a group. A (left) action of $G$ on $X$ 
    is a map 
    \[
        G \times X \to X 
    \]
    \[
        (g,x) \mapsto gx
    \]
    such that $ex = x$ and $g_1(g_2x) = (g_1g_2)x$ for all 
    $x \in X$ and $g_1,g_2 \in G$.
\end{defn}
The right action is similarily defined, and in the case where 
$G$ is abelian the actions coincide. 

\begin{ex}
    The following constitute group actions when 
    considering matrix-vector multiplication. 

    \begin{enumerate}
        \item $GL_2(\mathbb R)$ and its subgroups on $\mathbb R^2$. 
        \item same for $GL_n(\mathbb R)$ on $\mathbb R^n$.
    \end{enumerate}

    A very natural group action is $D_n$ acting on the regular polygon
    with $n$ sides, as $D_n$ \emph{is} the symmetries of 
    the regular $n$-gon.

    Furthermore $S_n$ (which is often thought of as the 
    bijections $f: [n] \to [n]$) can act on 
    said $[n]$. 

    Another group action which we have also seen many times 
    is conjugation where $G$ acts on itself. 
\end{ex}

\begin{defn} 
    \label{defn:faithful_group_action}
    Let $G$ be a group and $X$ a set. 

    We say that a group action $G \times X \to X$ is \emph{faithful} 
    if for any $g \in G, \; g\neq e$ implies there is some 
    $x \in X$ such that $g x \neq x$. 

    Alternatively it can be defined to be a group action for which 
    any $g_1,g_2 \in G$ with $g_1 \neq g_2$ implies that there 
    exists $x \in X$ such that $g_1 x \neq g_2 x$. 
 \end{defn}
 \begin{prop}
     Consider a group action $G \times X \to X$. 
     It defines a group homomorphism 
     $\phi : G \to \text{Sym}(X)$ for which 
     $\ker \phi = \{e\}$ if and only if the group action is 
     faithful. 
 \end{prop}
\newpage 
 \begin{prob}[Problem 11 Fraleigh's p.160]
     Let $X$ be a $G$-set. Show that $G$ acts faithfully on $X$ 
     if and only if no two distinct elements of $G$ have the same 
     action on each element $X$. 

     (I.e. prove the equivalent definition posed in 
     \cref{defn:faithful_group_action}) 
 \end{prob}
 \begin{proof}
    Let $G$ be a group, $X$ a $G$-set.  

    Assume first that the action of $G$ on $X$ is faithful. 
    Then $e \neq g \in G$ implies the existence of some 
    $x \in X$ for which 
    \[
        gx \neq x
    \]
    Suppose, for a contradiction, that there exist some 
    distinct $g_1,g_2 \in G$ which coincide on every $x \in X$. 
    Then 
    \begin{align*}
        g_1 x &= g_2 x \\ 
        g_2^{-1}g_1x &= x
    \end{align*}
    In other words there exists $g = g_2^{-1}g_1 \in G$ 
    for which every $x \in X$ is a fixed-point, contradicting 
    the assumption that $G$ acts faithfully on $X$. Thus $G$ acting 
    faithfully on $X$ must imply that any distinct pair of elements 
    of $G$ have some $x \in X$ for which they do not coincide. 

    Assume now that $G$ acts on $X$ such that for every 
    distinct $g_1,g_2 \in G$ for which there exists $x \in X$ 
    such that $g_1 x \neq g_2 x$. Let $g_2 = e \neq g_1$. 
    They are clearly distinct and furthermore 
    \[
        g_1 x \neq g_2x = x
    \]
    This holds for any $g_1 \in G$ so $G$ acts faithfully on $X$.  
 \end{proof}

 \begin{defn}[Orbit]
     Let $G$ be a group acting on a $G$-set $X$. Let $x \in X$.  

     The orbit of $x$, often denoted $\text{Orb}(x)$ or $Gx$ 
     is the set 
     \[
        \text{Orb}(x) := \left\{gx \bigm| g \in G\right\}
     \]
 \end{defn} 
 \begin{defn}[Stabilizer]
     Let $G$ be a group acting on a $G$-set $X$. Let $x \in X$. 

     The stabilizer of $x$, often denoted $\text{Stab}_G(x)$ 
     or $G_x$ is the set 
     \[
         \left\{g \in G \bigm| gx = x\right\}
     \]
 \end{defn}
 \newpage 
 \begin{thm}[Orbit-Stabilizer]
     For a finite group $G$ acting on a set $X$, the following 
     holds (conditions are equivalent). 

     \[
         |\text{Orb}(x)| = [G : \text{Stab}_G(x)]
     \]
     \[
         |G| = |\text{Orb}(x)| \cdot |\text{Stab}_G(x)|
     \]
 \end{thm}
 \begin{prob}
     Let $\{X_i \bigm| i \in I\}$ be a disjoint collection of 
     sets. Let each $X_i$ be a $G$-set for the same group $G$. 
    
     a) Show that $\bigcup_{i\in I}X_i$ can be viewed in a 
     natural way as a $G$-set, the \textbf{union} of the $G$-sets 
     $X_i$. 

     b) Show that every $G$-set $X$ is the union of its orbits. 
 \end{prob}
 \begin{proof}[Proof of (a)]
     Let $X=\bigcup_{i\in I}X_i$. Since the union is disjoint, each
     $x\in X$ lies in a unique $X_i$. Define
     \[
         g\cdot x := g\cdot_{X_i} x
     \]
     where the right-hand side denotes the given $G$-action on $X_i$.

     This is well-defined because the union is disjoint.
     For $x\in X_i$ we have
     \[
         e\cdot x = x
     \]
     since $X_i$ is a $G$-set, and
     \[
         (gh)\cdot x
         = (gh)\cdot_{X_i} x
         = g\cdot_{X_i}(h\cdot_{X_i}x)
         = g\cdot(h\cdot x).
     \]
     Thus the action axioms hold, and $X$ becomes a $G$-set.
 \end{proof}
 \begin{proof}[Proof of (b)]
     For $x\in X$, define the orbit
     \[
         G\cdot x = \{ g\cdot x \mid g\in G\}.
     \]
     Since $x=e\cdot x$, every element lies in its orbit,
     so
     \[
         X = \bigcup_{x\in X} G\cdot x.
     \]

     If $G\cdot x \cap G\cdot y \neq \emptyset$, then
     $g\cdot x = h\cdot y$ for some $g,h\in G$.
     Then
     \[
         x = g^{-1}h\cdot y,
     \]
     so $x\in G\cdot y$, which implies
     $G\cdot x = G\cdot y$.
     Thus distinct orbits are disjoint, and $X$ is the
     disjoint union of its orbits.
 \end{proof}
\newpage 
 \begin{thm}[Burnside's Formula]
     Let $G$ be a finite group and $X$ a finite $G$-set. If 
     $r$ is the number of orbits in $X$ under $G$, then 
     \[
        r \cdot |G| = \displaystyle\sum_{g \in G}|X_g|
     \]
     where $X_g$ are the \emph{fixed points} of $g$, namely 
     all $x \in X$ satisfying $gx = x$. 
 \end{thm}
 \begin{prob}[Problem 1 Fraileigh's p.164]
    Find the number of orbits in $\{1,\ldots, 8\}$ under the 
    cyclic subgroup $\langle (1 \ 3 \ 5 \ 6) \rangle \leq S_8$.
 \end{prob}
 \begin{proof}[Solution]

    Both $[8]$ and the given subgroup of $S_8$ are finite so we 
    use Burnside's formula. Denote 
    $\langle \sigma \rangle := \langle 1 \ 3 \ 5 \ 6\rangle$. 
    We know that $|\sigma| = 4$ so we must simply calculate 
    the sum of sizes of fixed points of each permutation in 
    $\langle \sigma \rangle$. Obviously $|X_e| = |X| = 8$.   
    Next up $\sigma$ has $2,4,7,8$ as fixed points so 
    $|X_\sigma| = 4$. Next up $\sigma^2 = (1 \ 5)(3 \ 6)$ and has 
    the same fixed points. $\sigma^3 = (1 \ 6 \ 5 \ 3)$ with 
    the same fixed points once again. Thus 
    \[
        r = 8 + 4 + 4 + 4  / |\sigma| = 5
    \]
 \end{proof}

 \newpage 
 \section{Rings, Fields \& Integral domains}
 \begin{defn}[Ring]
     A ring $R$ is a set together with two binary operations 
     $+$ and $\cdot$, written $(R, +, \cdot)$, satisfying 
     \begin{enumerate}
        \item $(R,+)$ is an abelian group (identity denoted $0$), 
        \item $(R,\cdot)$ has closure and associativity, 
        \item with distributive laws, i.e. 
            \[
                a \cdot (b + c) = ab + ac, \ (b+c) \cdot a = 
                ba + ca. 
            \]
     \end{enumerate}
     If $R$ has a multiplicative identity $1$, we say 
     $R$ is a ring with identity, or $R$ is a ring with unity 
     or $R$ is unital. 

     If $ab = ba$ for all $a,b \in R$, then $R$ is a commutative ring.
 \end{defn}
 So what, informally, a ring is, is a structure where you can 
 add, multiply and subtract, but not necessarily divide. 

 Rings have a bunch of immediately nice properties such as 
 \begin{itemize}
    \item $0 \in R$ is unique (follows from $(R,+)$ 
        being abelian grp),
    \item $-a$ is unique (follows from the same logic),
    \item $a \cdot 0 = 0 = 0 \cdot a$ (very nice), 
    \item $|R| \geq 2$, then $1 \neq 0$, 
    \item $-(a+b) = (-a) + (-b)$, 
    \item $-(-a) = a$, 
    \item $-(ab) = (-a)b = a(-b)$, 
    \item $ab = (-a)(-b)$.
 \end{itemize}
 
 \begin{defn}[Unit]
     Let $(R,+,\cdot)$ be a ring with identity, and let $u \in R$. 
     $u$ is called a unit if $u$ has a multiplicative inverse 
     $u^{-1}$ such that 
     \[
         u \cdot u^{-1} = u^{-1} \cdot u = 1
     \]
     We denote the set of units 
     \[
         \{u \in R \bigm| u \ \text{unit}\} =: R^x
     \]
     (This set is a group under the multiplication of $R$)
 \end{defn}

 \newpage 
 \begin{defn}[Zero divisor]
    Let $(R,+,\cdot)$ be a ring. A zero divisor is a non-zero 
    element $a \in R$ such that there exists non-zero 
    $b \in R$ satisfying 
    \begin{align*}
        ab &= 0 & \text{left zero divisor}
        ba &= 0 & \text{right zero divisor}
    \end{align*}
    Where these coincide in commutative rings. 
 \end{defn}
 \begin{defn}[Field]
     Let $(R,+,\cdot)$ be a ring. 

     $R$ is a \textbf{field} if 
     $(R\setminus \{0\}, \cdot)$ is an abelian group. 
 \end{defn}
 \begin{defn}
     Let $(R,+,\cdot)$ be a ring.  

     $R$ is a \textbf{division ring} if $(R \setminus \{0\}, \cdot)$ 
     is a group.
 \end{defn}

 Crucially, in both a divison ring and a field we know that 
 $\exists 1 \in R$, the multiplicative identity, meaning 
 a field/division ring is unital. 

The prototypical example of a commutative unital ring 
is $(\mathbb Z, +, \cdot)$. Importantly, $\mathbb Z$ is not 
a field. So a big area of focus will be to look at larger 
more powerful structures. 

\begin{defn}[Integral Domain]
    Let $(R,+,\cdot)$ be a ring. 

    $R$ is an \textbf{integral domain} if it is a commutative, 
    unital ring. 
\end{defn}
\begin{prop}
    A ring $R$ is an integral domain if and only if 
    $R$ has no zero-divisors and $|R| \geq 2$. 
\end{prop}
The nice property we can gleam from integral domains is that 
$a\cdot b = 0$ if and only if either $a$ or $b$ is zero.  

Usually, when we say integral domain, or field, we will be refering 
to said structure when it has $2$ or more elements. 
The reason for this is that a ring or field with 
$1$ element is not worth dwelling on. 

\begin{prop}
    A field is an integral domain. 
\end{prop}
\begin{proof}[Proof sketch]
    If $ab = 0$ and $a \neq 0$, then $b = a^{-1}(ab) = a^{-1}0 = 0$. 
\end{proof}

\begin{thm}
    Every finite integral domain is a field.
\end{thm}
\begin{proof}
    Let $D$ be a finite integral domain. 

    $(D\setminus \{0\},\cdot)$ is closed, because $D$ has no 
    zero divisors. Furthermore it is associative and $1 \in D 
    \setminus \{0\}$ by definition. The last thing we need 
    is inverses. 

    Let $r \in D, \: r \neq 0$. 
    Consider $r, r^2, r^3, \ldots r^k, \ldots$ 

    As $D$ is finite there must be repeted powers in this list. 
    Let $r^i = r^j, \: i < j$ wlog. Then 
    \[
        \underbrace{r\cdot r\cdots r}_{i} \cdot 1 = 
        \underbrace{r\cdot r \cdots r}_{j}
    \]
    Use left cancellation repeatedly to get 
    \[
        r^{j-i} = 1
    \]
    Let $s = r^{j-i-1}$. Then $rs = sr = 1$, so $s = r^{-1}$.
\end{proof}
\begin{thm}
    Every finite division ring is a field. 
\end{thm}

\newpage 
\section{Some problems pertaining to Rings, Fields, etc.}
\begin{prob}[Problem 3 \& 5 Fraleigh's p.174]
    Compute the products
    \[
        (11)(-4) \in \mathbb Z_{15}
    \]
    \[
        (2,3)(3,5) \in \mathbb Z_5 \times \mathbb Z_9
    \]
\end{prob}
\begin{proof}[Solution]
    \begin{align*}
        11 \cdot -4 &= -(11 \cdot 4) \\ 
                    &= -(44) \\ 
                    &\equiv 1 \mod 15
    \end{align*}

    \begin{align*}
        (2,3)(3,5) &= (2\cdot 3, 3 \cdot 5) \\ 
                   &= (6, 15) \\ 
                   &\equiv (1,6)
    \end{align*}
\end{proof}
\begin{prob}[Problem 11 Fraleigh's p.175]
    Determine whether the following structure is a ring, 
    then if it is a commutative ring, whether it is unital, 
    and whether it is a field. 

    \[
        \left\{ a + b \sqrt 2 \bigm| a,b \in \mathbb Z \right\}
    \]
\end{prob}
\begin{proof}[Solution]
    Let's denote this structure as $\mathbb Z(\sqrt 2)$.

    It is quite clear that $(\mathbb Z(\sqrt 2), +)$ forms an 
    abelian group as the integers is an abelian group 
    and the identity here is $0 + 0\sqrt 2 \in \mathbb Z(\sqrt 2)$. 
    Furthermore $(a+b\sqrt 2) + (c+d\sqrt 2) = 
    (a+c) + (b+d)\sqrt 2 \in \mathbb Z(\sqrt 2)$, so we're good. 

    The $\mathbb Z \setminus \{0\}$ with multiplication is closed, 
    but not every element has an inverse (in fact only $1$ and $-1$  
    do), so we'd expect something similar here. 

    Suppose $a+b\sqrt 2, c+d\sqrt 2 \in \mathbb Z(\sqrt 2)
    \setminus \{0\}$. Then 
    \begin{align*}
        (a+b\sqrt 2)(c + d\sqrt 2) &= ac + cb\sqrt2 + ad\sqrt 2 
        +2bd \\ 
        &= (ac + 2bd) + (ad + cb)\sqrt 2 \in \mathbb Z(\sqrt 2) 
        \setminus \{0\}
    \end{align*}
    So the multiplicative structure is closed with identity, but 
    there does not exist inverses for every element as for instance
    $2 \in \mathbb Z(\sqrt 2)^x$, but $2^{-1}$ does not exist. 
    It is also not hard to see that distributive laws work fine. 

    Thus we conclude that $\mathbb Z(\sqrt 2)$ is a 
    commutative unital ring. 
\end{proof}
\begin{prob}[Problem 20 Fraleigh's p.175]
    Consider the matrix ring $M_2(\mathbb Z_2)$. 

    a. Find the order (number of elements) of the ring.

    b. List all units in the ring.
\end{prob}
\begin{proof}[Solution of (a)]
    Recall that 
    \[
        M_2(\mathbb Z_2) 
        =
        \left\{
        \begin{pmatrix}
            a & b \\ 
            c & d
        \end{pmatrix}
        \Biggm| 
        a,b,c,d \in \mathbb Z_2
        \right\}
    \]  
    There are $4$ entries in each matrix in the ring, all of which 
    can take on one of two values from $\mathbb Z_4$. Thus 
    there are $2^4 = 16$ elements. 
\end{proof}
\begin{prob}[Problem 37 Fraleigh's p.176]
    Show that if $U$ is the collection of all units in a ring  
    $(R,+,\cdot)$ with unity, then $(U,\cdot)$ is a group. 
\end{prob}
\begin{proof}
    We conisder $U$ the collection of the units of a ring 
    $R$. $1 \in R$ (as $R$ has unity), 
    $(R,\cdot)$ has associativity by definition and 
    the set of units $U$ is precisely those $u \in R$ such that 
    \[
        u \cdot u^{-1} = u^{-1} \cdot u = 1
    \]
    so it remains to show that $U$ is closed. 
    Let $u,v \in U$. Then, as said, there exists inverses 
    $u^{-1},v^{-1} \in U$. Then  
    \begin{align*}
        (uv)(v^{-1}u^{-1}) &= u(vv^{-1})u^{-1} \\ 
                           &= u(1)u^{-1} \\ 
                           &= uu^{-1} \\ 
                           &= 1
    \end{align*}
    so $(uv)^{-1} = (v^{-1}u^{-1}) \in U$. Thus the product 
    of two units is itself a unit. 
    We have shown $(U,\cdot)$ to have all the properties of 
    a group.
\end{proof}

\begin{prob}[Problem 46 Fraleigh's p.176]
    An element $a$ of a ring $R$ is nilpotent if $a^n = 0$ for some 
    $n \in \mathbb Z^+$. Show that if $a$ and $b$ are nilpotent 
    in some ring commutative ring $R$, then $a+b \in R$ is 
    also nilpotent. 
\end{prob}
\begin{proof}
    Suppose $a,b \in R$ (commutative) such that 
    $a^n = 0$ and $b^m = 0$ for some $n,m \in \mathbb Z^+$.

    Let $k \in \mathbb Z^+$. Then
    \begin{align*}
        (a+b)^k &= \displaystyle\sum_{i=0}^k \binom{k}{i}a^{k-i}b^i \\
                &= a^k + ka^{k-1}b + \cdots + kab^{k-1} + b^k
    \end{align*}
    in particular, if $k := n + m$ then we observe the following: 
    \begin{align*}
        (a+b)^{n+m} &=  
        \displaystyle\sum_{k=0}^{n+m}\binom{n+m}{k}a^{n+m-k}b^k \\ 
        &= 
        \displaystyle\sum_{k=m + 1}^{n+m}\binom{n+m}{k}a^{n+m-k}b^k +
        \displaystyle\sum_{k=0}^{m}\binom{n+m}{k}a^{n+m-k}b^k \\ 
        &= \underbrace{\binom{n+m}{m+1}a^{n+m-m-1}b^{m+1}}_{
        \cdots \; a^{n-1}\cdot0\cdot b = 0} + 
        \ldots + 
        \underbrace{b^{n+m}}_{(b^n)\cdot 0 = 0} \\ 
        &+ \underbrace{a^{n+m}}_{0 \cdot a^m = 0} + \cdots + 
        \underbrace{\binom{n+m}{m}a^{n+m-m}b^{m}}_{\cdots \; 
        0 \cdot 0 = 0} \\ 
        &= 0
    \end{align*}
    since every term is $0$. 
\end{proof}

\begin{prob}[Problem 1 Fraleigh's p.182]
    Find all solutions of $x^3 - 2x^2 -3x = 0$ in $\mathbb Z_{12}$.
\end{prob}
\begin{proof}[Solution]
    \begin{align*}
        f(x) = x^3 - 2x^2 - 3x &= 0 \\ 
        x(x^2 - 2x - 3) &= 0 \\ 
        x(x-3)(x+1)     &= 0 
    \end{align*}
    Since we are in a ring with zero-divisors we are really asking 
    for which $x \in \mathbb Z_{12}$ the following holds 
    \[
        12 \bigm| x(x-3)(x+1)
    \]
    We begin by solving $\mod 3$ where 
    \[
        f(x) = x^2(x+1) \in \mathbb Z_3[x]
    \]
    with solutions $0,2 \in \mathbb Z_3$. $f(x)$ reduces to 
    \[
        f(x) = x(x-1)^2 \in \mathbb Z_4[x]
    \]
    with solutions $0,1,2 \in \mathbb Z_4$. 
    We get the solutions 
    \[
        \{0,2,5,6,8,9\} \subset \mathbb Z_{12}
    \]
\end{proof}
\begin{prob}[Problem 11 Fraleigh's p.182]
    Let $R$ be a commutative unital ring with characteristic 
    $4$. Compute and simplify $(a+b)^4$ for $a,b \in R$. 
\end{prob}
\begin{proof}[Solution]
    \begin{align*}
        (a+b)^4 &= a^4 + \binom{4}{1}a^3b + \binom{4}{2}a^2b^2 + 
        \binom{4}{3}ab^3 + b^4 \\ 
                &= a^4 + 4a^3b + 6a^2b^2 + 4ab^3 + b^4
    \end{align*} 
    As $4 \equiv 0 \mod 4$ and $6 \equiv 2 \mod 4$ we get 
    $(a+b)^4 \equiv a^4 + 2a^2b^2 + b^4$ in integral domain $R$ 
    with $\text{char}(R) = 4$. 
\end{proof}
\begin{prob}[Problem 23 Fraleigh's p.183]
    Show that a division ring has exactly two idempotent elements. 
\end{prob}
\begin{proof}
    Recall that we define division rings as non-trivial so 
    $|R| \geq 2$ crucially. The candidates for idempotent elements 
    are $0,1 \in R$ as $0 \cdot 0 = 0$ as $0\cdot a = 0 \; 
    \forall a \in R$. Then since $1 \cdot a = a \; \forall a \in R$ 
    we conclude that $1 \cdot 1 = 1$. Now we need to show that 
    there do not exist any other elements which satisfy this 
    property. Assume, for a contradiction, that there exists 
    $a \in R \setminus \{0,1\}$ such that 
    \[
        a^2 = a
    \]
    As $R$ is a division ring we know there exists $a^{-1} \in R$ 
    such that 
    \begin{align*}
        a^2a^{-1} &= aa^{-1} \\ 
        a(aa^{-1}) &= 1 \\ 
        a &= 1
    \end{align*}
    contradiction our assumption that $a \neq 1$. 
\end{proof}
\begin{prob}[Problem 29 Fraleigh's p.183]
    Show that the characteristic of an integral domain $D$ is either 
    $0$ or a prime $p$. 
\end{prob}
\begin{proof}
    Let $D$ be an integral domain, i.e. a commutative ring 
    with unity and no zero-divisors. 
    Assume, for a contradiction, that 
    $\text{char}(D) = k \in \mathbb Z^+$ (not prime). 
    Then $\text{char}(D) = k = nm, \: n,m \in \mathbb Z^+$.
    Then we can conclude that 
    \[
        k\cdot 1 = \underbrace{1+1+\cdots+1+1}_{k=nm \ \text{times}} 
        = 0
    \]
    but then 
    \[
        (n\cdot 1)(m\cdot 1) = 0
    \]
    so $n\cdot 1 = 0$ or $m\cdot1=0$ so $k$ is not the characteristic
    of $D$. 
\end{proof}

\newpage 
\section{Fermat's and Euler's theorems, Quotient Fields}
\begin{thm}[Fermat's Little Theorem]
    If $p$ is prime and $\gcd(a,p) = 1$, then 
    \[
        a^{p-1} \equiv 1 \mod p
    \]
    Equivalently 
    \[
        a^p \equiv a \mod p
    \]
\end{thm}
\begin{proof}[Proof idea]
    The key fact is that 
    \[
        (\mathbb Z/p\mathbb Z)^\times
    \]
    the nonzero elements mod $p$, form a group under multiplication.
    This group has $p-1$ elements (finite) and cyclic. By 
    Lagrange's Theorem, if $G$ is finite and $g\in G$ then 
    \[
        g^{|G|} = e
    \]
    Applying this to $G = (\mathbb Z/p\mathbb Z)^\times$ we get 
    \[
        a^{p-1} \equiv 1
    \]
\end{proof}

\begin{thm}{Euler's Theorem}
    If $\gcd(a,n)=1$, then 
    \[
        a^{\phi(n)} \equiv 1 \mod n,
    \]
    where $\phi(n)$ is Euler's totient function.
\end{thm}
\begin{proof}[Proof idea]
    We consider 
    \[
        (\mathbb Z/n\mathbb Z)^\times
    \]
    the group of units mod $n$. 
    It has exactly $\phi(n)$ elements. 
    Again by Lagrange: 
    \[
        a^{\phi(n)} \equiv 1 
    \]
\end{proof}
So Fermat's Little Theorem is just Euler's Theorem for $n$ prime. 

\subsection{Quotient Fields}
In $\mathbb Z$, you generally cannot divide by $2$. 
But in $\mathbb Q$, you can. 
The questions we want to answer is: 

"Can we build a field from a ring?"

As it turns out the answer is yes (conditions may apply...).

\newpage 
Recall that an integral domain is 
\begin{itemize}
    \item a commutative ring, 
    \item with identity, 
    \item with no zero divisors
\end{itemize}
Suppose $D$ is an integral domain. We construct 
$\operatorname{Frac}(D)$ like rational numbers.  
 
\begin{defn}[Quotient Field]
    The \textbf{field of fractions} or \textbf{quotient field}
    of an integral domain $D$ is the smallest field $K$ 
    (often denoted $\operatorname{Frac}(D)$)
    containing $D$ as a subring, constructed by forming formal 
    fractions $\frac{a}{b}$ where $a,b \in K \; (b \neq 0)$. 
    
    The operations on $K$ are defined as 
    \[
        \frac{a}{b} + \frac{c}{d} = \frac{ad + bd}{bd}
    \]
    and 
    \[
        \frac{a}{b}\cdot\frac{c}{d} = \frac{ac}{bd}
    \]

    We equip $K$ with an equivalence relation $\sim$ where we 
    say 
    \[
        \frac{a}{b} \sim \frac{c}{d}
    \]
    if $ad=bc$.
\end{defn}
As we see we get a field $\operatorname{Frac}(D)$ and 
we get an injective embedding
\[
    D \hookrightarrow{} \operatorname{Frac}(D)
\]
via 
\[
    a \mapsto \frac{a}{1}
\]


\begin{prob}[Problem 20.5 Fraleigh's p.189]
    Use Fermat's theorem to find the remainder of 
    $37^{49}$ when it is divided by $7$. 
\end{prob}
\begin{proof}[Solution]
    $\gcd(37,7) = 1$ and $p = 7$ is ceirtainly prime so 
    Fermat's Little Theorem applies. 

    We can rewrite 
    \[
        37^{49} = 37^{7^2} \equiv 2^{7^2}
    \]
    From FLT we have $2^{6} \equiv 1$. 
    \[
        2^{7^2} = 2^{6 \cdot 8 + 1} \equiv 1^8 \cdot 2 \equiv 2
        \mod 7
    \]
\end{proof}
\begin{prob}[Problem 20.12 Fraleigh's p.189]
    Describe all solutions to 
    \[
        22x \equiv 5 \mod 15
    \]
\end{prob}
\begin{proof}[Solution]
    The greatest common divisor $\gcd(22,15) = 1$ which is a 
    divisor of $5$ so there does exist solutions (in fact, as 
    $\gcd(22,15) = 1$ there exist exactly one solution). We can 
    simplify to get 
    \[
        7x \equiv 5 \mod 15
    \]
    As the greates common divisor is $1$ we know that $7$ is a 
    unit in $\mathbb Z/15\mathbb Z$. We want to find $7^{-1}$. 
    Notice that $7 \cdot 13 = 91 \equiv 1$. 
    \begin{align*}
        x \equiv 13 \cdot 5 \mod 15 \\ 
        13 \cdot 5 = 65 \\ 
        65 \equiv 5 \mod 15
    \end{align*}
    Thus 
    \[
        x \equiv 5 \mod 15
    \]
\end{proof}
\begin{prob}[Problem 20.15 Fraleigh's p.189]
    As in the task above for 
    \[
        39x \equiv 125 \mod 9
    \]
\end{prob}
\begin{proof}[Solution]
    Notice that 
    \[
        39 \equiv 3 \mod 9
    \]
    and 
    \[
        125 \equiv -1 \equiv 8 \mod 9
    \]
    so we can transform the problem into 
    \[
        3x \equiv 8 \mod 9
    \]
    Notice that $\gcd(3,9) = 3$, but $3$ is not a divisor of 
    $8$ thus the solution DNE. 
\end{proof}
\begin{prob}[Problem 20.27 Fraleigh's p.190]
    Show that $1$ and $p-1$ are the only elements of the 
    field $\mathbb Z_p$ that are their own multiplicative inverse. 
    [Hint: Consider the equation $x^2-1 = 0$].
\end{prob}
\begin{proof}
    Let $p$ be prime and let $x \in \mathbb Z_p$. 
    Notice that $x = x^{-1}$ iff 
    \[
        x^2 \equiv 1 \mod p \Leftrightarrow x^2 - 1 \equiv 0 \mod p
    \]
    It is not hard to see that $1$ and $p-1$ are solutions to 
    these equations, but we have to show that they are the only ones.
    Notice that $x^2 - 1$ can be rewritten so we get 
    \[
        (x-1)(x+1) \equiv 0 \mod p
    \]
    Recall that since $\mathbb Z_p$ is a field (no zero divisors) 
    it is the case that 
    $x-1 \equiv 0$ or $x + 1 \equiv 0$. I.e. 
    \[
        x \equiv 1 \mod p, \ x \equiv - 1 \mod p.
    \]
    Therefore
    \[
        x = 
        \begin{cases}
            1 \\ p-1
        \end{cases}
    \]
\end{proof}
\begin{prob}[Problem 21.1 Fraleigh's p.196]
    Describe the field $F$ of quotients of the integral subdomain
    \[
        D = \left\{n+mi \bigm| n,m \in \mathbb Z\right\}
    \]
    of $\mathbb C$. "Describe" means give the elements of 
    $\mathbb C$ that make up the field of quotients of $D$ 
    in $\mathbb C$.
\end{prob}
\begin{proof}[Solution]
    Recall that 
    \[
        \operatorname{Frac}(D) = 
        \left\{ \frac{a}{b} \biggm| a,b \in D \; (b \neq 0) \right\}
    \]
    $a,b \in D$ means that $a,b$ can be written as 
    \[
        a = n+mi, \ b = k+li, \ n,m,k,l \in \mathbb Z
    \]
    (note: $k+li \neq 0 \Rightarrow k \neq 0 \neq li$)
    so 
    \[
        \operatorname{Frac}(D) 
    \]
    are elements of the form 
    \begin{align*}
        \frac{n+mi}{k+li} &= \frac{n+mi}{k+li}\cdot\frac{k-li}{k-li}\\
        &= \frac{(n+mi)(k-li)}{k^2 + l^2} \\ 
        &= \frac{nk+kmi-nli + ml}{k^2 + l^2} \\ 
        &= \frac{nk+ml + (km-nl)i}{k^2 + l^2} \\ 
        &= \frac{nk+ml}{k^2 + l^2} + \frac{km-nl}{k^2+l^2}i 
    \end{align*}
    Now we know $\operatorname{Frac}(D) \subseteq \mathbb Q(i)$, but 
    is this subset relation strict?
    Let $a = r/s, \ b = t/s$. 
    Then 
    \[
        a+bi = \frac{r}{s} + \frac{t}{s}i 
    \]
    and since 
    \[
        r+ti \in \mathbb Z(i) = D, \ \text{and} \ 
        s \in \mathbb Z \subset \mathbb Z(i) = D
    \]
    we conclude that $\mathbb Q(i) \subseteq D$. 
    I.e. 
    \[
        D = \operatorname{Frac}(\mathbb Z(i)) = \mathbb Q(i) 
    \]
\end{proof}

\begin{prob}[Problem 21.12 Fraleigh's p.197]
    Let $R$ be a nonzero commutative ring, and let $T$ be 
    a nonempty subset of $R$ closed under multiplication and 
    containing neither $0$ nor divisors of $0$. 
    Starting with 
    $R \times T$ and otherwise exactly following the construction 
    in this section, we can show that the ring $R$ can be enlarged 
    to a partial ring of quotients $Q(R,T)$. Think about this for 
    $15$ minutes or so; look back over the construction and see 
    why things still work. In particular, show the following: 

    a. $Q(R,T)$ has unity even if $R$ does not. 

    b. In $Q(R,T)$, every nonzero element of $T$ is a unit. 
\end{prob}
\begin{proof}[Proof of (a)]
    We let $R$ be some nonzero commutative ring 
    and take a closed subset $T$ not containing $0$ or 
    any divisors of it and construct 
    $Q(R,T)$ to be 
    \[
        \frac{r}{t}, \; r \in R \ \text{and} \ t \in T
    \]
    Let $x \in Q(R,T)$ then 
    \[
        x = \frac{r}{t}
    \]
    for some $r \in R, \; t \in T$. As $T \subseteq R$ we 
    know that 
    \[
        \frac{t}{t} \in Q(R,T)
    \]
    Notice that 
    \[
        \frac{t}{t} \cdot \frac{r}{t} 
        = \frac{tr}{t^2} = 
        \frac{r}{t} \cdot \frac{t}{t}
    \]
    but 
    \[
        \frac{tr}{t^2} = \frac{r}{t}
    \]
    as 
    \[
        t^2 r = trt
    \]
    so $t/t = 1_{Q(R,T)}$. 
\end{proof}
\begin{proof}
    Suppose $t,s \in T$, nonzero, with  
    \[
        \frac{t}{s} \in Q(R,T)
    \]
    We claim this arbitrary nonzero element is invertible.
    Consider 
    \[
        \frac{s}{t}
    \]
    Observe that 
    \[
        \frac{t}{s}\frac{s}{t} = \frac{t}{t} = \frac{s}{t}\frac{t}{s}
    \]
    and as we showed in (a), and following from $st = ts$, we 
    get 
    \[
        1 = \frac{s}{t}\frac{t}{s}
    \]
\end{proof}
\begin{prob}[Problem 22.2 Fraleigh's p.207]
    Find the sum and product of the given polynomial in the 
    polynomial ring: 
    \[
        f(x) = x + 1, \: g(x) = 3x^2 + 2x + 3, \: \text{in} 
        \ \mathbb Z_2[x].
    \]
\end{prob}
\begin{proof}[Solution]
    \begin{align*}
        f(x)+g(x) &= 3x^2 + 3x + 4 \\ 
                  &\equiv x^2 + x \mod 2
    \end{align*}
    \begin{align*}
        f(x)g(x) &= (x+1)(3x^2 + 2x + 3) \\ 
                 &\equiv (x+1)(x^2 + 1) \mod 2 \\ 
                 &= x^3 + x^2 + x + 1   
    \end{align*}
\end{proof}

\begin{prob}[Problem 22.5 Fraleigh's p.207]
    How many polynomials are there of degree $\leq 3$ in 
    $\mathbb Z_2[x]$? (Include $0$)
\end{prob}
\begin{proof}[Solution]
    The polynomials in $\mathbb Z_2$ are precisely 
    \[
        ax^3 + bx^2 + cx + d, \ a,b,c,d \in \mathbb Z_2
    \]
    So there are $4$ independent choices of coefficients which 
    each can take on one of two values. Thus there are 
    \[
        2^4 = 16
    \]
    possible polynomials in $\mathbb Z_2$.
\end{proof}
\begin{prob}[Problem 22.12/13 Fraleigh's p.207]
    Find all zeros of the indicated finite field of 
    the given polynomial in said field. 
    \[
        x^2+1,\ \text{in} \ \mathbb Z_2
    \]
    and 
    \[
        x^3 + 2x +2, \ \text{in} \ \mathbb Z_7
    \]
\end{prob}
\begin{proof}[Solution]
    \begin{align*}
        x^2 + 1 &= 0 \\ 
        x^2 &= -1 \equiv 1 \mod 2 
    \end{align*}
    so the only zero is $1 \in \mathbb Z_2$.

    Next we work with 
    \[
        x^3 + 2x + 2 = 0
    \]
    Let us begin by trying some values. Clearly $0$ does not work. 
    If we plug in $1$ we get 
    \[
        1 + 2 + 2 = 5 \neq 0
    \]
    Next $2$ gives us 
    \[
        8 + 4 + 2 = 14 \equiv 0
    \]
    so $2$ is a solution.
    \[
        27 + 6 + 2 = 35 \equiv 0
    \]
    \[
        64 + 8 + 2 = 74 \equiv 4 \neq 0  
    \]
    and so on to find the rest of the zeros. 
\end{proof}
\begin{prob}[Problem 27 Fraleigh's p.208]
    Let $F$ be a characteristic zero field and let $D$ be the 
    formal polynomial differentiation map, so that 
    \[
        D\left(\displaystyle\sum_{k=0}^{n}a_kx^k\right)
        = 
        \displaystyle\sum_{k=1}^{n}ka_kx^{k-1}
    \]
    
    a) Show that $D : F[x] \to F[x]$ is a group homomorphism 
    of $\langle F[x], +\rangle$ onto itself. Is $D$ a ring 
    homomorphism?

    b) Find the kernel of $D$. 

    c) Find the image of $F[x]$ under $D$.
\end{prob}
\begin{proof}[Proof of (a)]
    Let $f,g \in F[x]$, i.e. 
    \[
        f(x) := \displaystyle\sum_{k=0}^{n}a_kx^k, \ 
        g(x) := \displaystyle\sum_{k=0}^{n}b_kx^k
    \]
    Then 
    \begin{align*}
        D(f+g) 
        &= D\left(\left[\displaystyle\sum_{k=0}^{n}a_kx^k\right] 
        + \left[\displaystyle\sum_{k=0}^{n}b_kx^k\right]\right) \\ 
        &= D\left(\displaystyle\sum_{k=0}^{n}(a_k+b_k)x^k\right) \\ 
        &= \displaystyle\sum_{k=1}^{n}k(a_k+b_k)x^{k-1} \\ 
        &= \displaystyle\sum_{k=1}^{n}k(a_k)x^{k-1} + 
        \displaystyle\sum_{k=1}^{n}k(b_k)x^{k-1} \\ 
        &= D(f) + D(g)
    \end{align*}
    It is not a ring homomorphism. Consider 
    \[
        x(x^2 + 1)
    \]
    Then 
    \[
        D(x(x^2 + 1)) = D(x^3 + x) = 3x^2 + 1
    \]
    but 
    \[
        D(x)D(x^2 + 1) = 2x
    \]
\end{proof}
\begin{proof}[Solution of (b)]
    The preimage of the zero polynomial under differentiation is 
    simply all the constant polynomials i.e. 
    \[
        c \in F
    \]
\end{proof}
\begin{proof}[Solution of (c)]
    Differentiation is surjective so the image of $F[x]$ is 
    simply $F[x]$. Take any $f \in F[x]$, i.e. 
    $f = \sum_{k=0}^{n}a_kx^k$. Then 
    \[
        D^{-1}(f) = 
        c + \displaystyle\sum_{k=0}^{n}\frac{a_k}{k+1}x^{k+1}
    \]
    for any $c \in F$. So $f \in F[x]$ has precisely $|F|$ 
    elements in the preimage.
\end{proof}

\newpage 
\section{Polynomial Rings}
\begin{defn}
    Let $R$ be a ring. 

    A polynomial over $R$ is 
    \[
        \displaystyle\sum_{k=0}^{n}a_kx^k
    \]
    where $a_k \in R$ and $n \geq 0$.

    The set of all such polynomials is denoted $R[x]$, 
    and is called the polynomial ring over $R$.

    
\end{defn}

Two polynomials $\sum_{k=0}^{n}a_kx^k$ and $\sum_{k=0}^{n}b_kx^k$
are equal iff $a_k = b_k, \: \forall k \in [n]$. 

With addition, they are very nicely behaved: 
\[
    \displaystyle\sum_{k=0}^{n}a_kx^k + 
    \displaystyle\sum_{k=0}^{n}b_kx^k = 
    \displaystyle\sum_{k=0}^{n}(a_k+b_k)x^k
\]
while multiplication looks like: 
\[
    \displaystyle\sum_{k=0}^{n}a_kx^k \cdot 
    \displaystyle\sum_{k=0}^{n}b_kx^k = 
    \displaystyle\sum_k\left(
    \displaystyle\sum_{i+j=k}a_ib_j
    \right)x^k
\]

Notably, if $R$ is a ring then $R[x]$ is a ring. Furthermore 
commutativity, identity and being and integral domain 
carry over. Note that not all properties transfer, e.g. 
being a field does not always carry over. 

\subsection{Degree of a Polynomial}

\begin{defn}
Let 
\[
f(x) = a_n x^n + \dots + a_0
\]
with $a_n \neq 0$. The \textbf{degree} of $f$ is
\[
\deg(f) = n .
\]

The degree of the zero polynomial is defined to be $-\infty$.
\end{defn}

If $R$ is an integral domain then

\[
\deg(fg) = \deg(f) + \deg(g).
\]

\subsection{Units in Polynomial Ring}

If $R$ is an integral domain, the only units in $R[x]$ are 
the constant polynomials whose coefficients are units in $R$.

For example, the only units in $\mathbb{Z}[x]$ are $\pm 1$.

\subsection{Evaluation Homomorphisms}

For $c \in R$ define
\[
\varphi_c : R[x] \to R
\]
by
\[
\varphi_c(f(x)) = f(c).
\]

This map is a ring homomorphism and is called the 
\textbf{evaluation homomorphism}.

\subsection{Roots}

\begin{defn}
Let $f(x) \in R[x]$ and $c \in R$.
We say $c$ is a \textbf{root} of $f$ if
\[
f(c) = 0 .
\]
\end{defn}

\begin{thm}[Factor Theorem]
If $c$ is a root of $f(x)$ in $R$, then
\[
(x-c) \mid f(x).
\]
\end{thm}

If $R$ is an integral domain and $\deg(f)=n$, then $f$ has at
most $n$ roots in $R$.

\section{Polynomial Factorisation}

Polynomial arithmetic behaves especially well over fields.

\begin{thm}[Division Algorithm]
Let $F$ be a field and $f,g \in F[x]$ with $g \neq 0$.
Then there exist unique polynomials $q,r \in F[x]$ such that
\[
f(x) = q(x)g(x) + r(x)
\]
where either $r=0$ or
\[
\deg(r) < \deg(g).
\]
\end{thm}

This allows us to compute greatest common divisors of polynomials
using the Euclidean algorithm.

\begin{defn}
A polynomial $f(x) \in F[x]$ is \textbf{irreducible} if

\begin{itemize}
\item $\deg(f) \ge 1$
\item whenever $f = gh$ with $g,h \in F[x]$, one of $g,h$
is a unit.
\end{itemize}
\end{defn}

\begin{thm}
If $F$ is a field then $F[x]$ is a unique factorization domain.
In particular, every polynomial can be written as a product of
irreducible polynomials, uniquely up to order and multiplication
by units.
\end{thm}

\subsection{Useful Irreducibility Criterion}

If $f(x) \in F[x]$ has degree $2$ or $3$, then

\[
f \text{ is irreducible}
\quad \Longleftrightarrow \quad
f \text{ has no root in } F.
\]

Thus to test irreducibility of such polynomials it suffices
to check whether they have roots.

\newpage 
\section{Problems from week of March 22}
\begin{prob}[Problem 23.9 Fraleigh's p.218]
    The polynomial $x^4 + 4$ can be factored into linear 
    factors in $\mathbb Z_5[x]$. Find this factorization.
\end{prob}
\begin{proof}[Solution]
    We work with 
    \[
        f(x) = x^4 + 4 = (x^2 + 2x + 2)(x^2 - 2x + 2)
    \]
    We see that 
    \[
        x^2 + 2x + 2 \equiv x^2 + 2x - 3 = (x+3)(x-1) \equiv 
        (x+3)(x+4) 
    \]
    and 
    \[
        x^2 - 2x + 2 \equiv x^2 + 3x + 2 = (x+2)(x+1)
    \]
    so 
    \[
        f(x) \equiv (x+4)(x+3)(x+2)(x+1) 
    \]
\end{proof}
\begin{prob}[Problem 23.14 Fraleigh's p.218]
    Show that $f(x) = x^2 + 8x - 2$ is irreducible over 
    $\mathbb Q$. Is $f(x)$ irreducible over $\mathbb R$?
\end{prob}
\begin{proof}
    We consider $f(x) \in \mathbb Q[x]$.
    \[
        f(x) = x^2 + 8x - 2 
    \]
    Since $f(x)$ is degree $2$ it is irreducible in $\mathbb Q$ 
    if and only if 
    it has no root in $\mathbb Q$. 
    \begin{align*}
        f(x) &= 0 \\ 
        x^2 + 8x - 2 &= 0 \\ 
        (x+4-3\sqrt{2})(x+4+4\sqrt 2) &= 0
    \end{align*}
    $\sqrt 2 \not \in \mathbb Q[x]$ so $f(x)$ is irreducible in 
    $\mathbb Q$. 

    As $\sqrt 2 \in \mathbb R$ and $f(x)$ is degree $2$ we conclude 
    that $f(x)$ is reducible in $\mathbb R[x]$. 
\end{proof}
\begin{prob}[Problems 18 \& 19 Fraleigh's p.219]
    Determine whether the given polynomials satisfy the Eisenstein
    criterion for irreducibility over $\mathbb Q$. 

    \[
        f(x) = x^2-12
    \]
    \[
        g(x) = 8x^3 + 6x^2 -9x + 24
    \]
\end{prob}
\begin{proof}[Solution]
    Recall that for some polynomial 
    \[
        \displaystyle\sum_{k=0}^{n}a_kx^k
    \]
    If there exists a prime $p$ which divdes $a_0,\ldots,a_{n-1}$
    and doesn't divide $a_n$ and $p^2$ doesn't divide $a_0$, 
    then it is irreducible over the rationals. 

    For 
    \[
        f(x) = x^2 - 12
    \]
    we have $a_0 = -12$, $a_1 = a_n = 1$. 
    No prime number divides $1$ so that condition is fulfilled. 
    Then we see that $3 | 12$, but $3^2$ does not divide $-12$.
    We conclude that $f(x)$ is irreducible in $\mathbb Q[x]$.

    Next look at 
    \[
        g(x)
    \]
    Observe that for $p = 3$ we have 
    \begin{align*}
        p &\bigm| 24 \\ 
        p &\bigm| -9 \\ 
        p &\bigm| 6 \\ 
    \end{align*}
    but $p$ does not divide $8$ and $p^2 = 9$ does not divide 
    $24$ so $g(x)$ is irreducible over the rationals.
\end{proof}
\begin{prob}[Problem 23.34 Fraleigh's p.219]
    Show that for a prime $p$, the polynomial 
    $x^p + a$ in $\mathbb Z_p[x]$ is not irreducible for any 
    $a \in \mathbb Z_p$
\end{prob}
\begin{proof}
    Let $p$ be prime and 
    \[
        x^p + a \in \mathbb Z_p[x]
    \]
    By Fermat's Little Theorem $a = a^p$. 
    Furthermore observe that 
    \[
        x^p + a = x^p + a^p = (x+a)^p
    \]
    So $x^p + a$ is reducible to 
    \[
        \prod_{[p]}(x+a) \in \mathbb Z_p[x]
    \]
 \end{proof}

 \begin{prob}[Problem 23.35 Fraleigh's p.219]
     If $F$ is a field and $a \neq 0$ is a zero of $f(x) = 
     a_0 + a_1x+\cdots x_nx^n$ in $F[x]$, show that 
     $1/a$ is a zero of $a_n + a_{n-1}x + \cdots + 
     a_0x^n$.
 \end{prob}
\begin{proof}["Thinking out loud" proof]
    Let $F$ be a field and let
    \[
        f(x) := \sum_{i=0}^{n} a_i x^i
    \]
    be a polynomial in $F[x]$. Assume $a \neq 0$ is a root of $f$, 
    i.e.
    \[
        f(a) = \sum_{i=0}^{n} a_i a^i = 0 .
    \]

    Consider the expression
    \[
        \frac{f(x)}{x^n}
        =
        \sum_{i=0}^{n} a_i x^{i-n}
        =
        a_0 x^{-n} + a_1 x^{1-n} + \cdots + a_{n-1} x^{-1} + a_n .
    \]
    This is not a polynomial in $F[x]$, but it suggests reversing the
    powers of $x$.

    To obtain a genuine polynomial, define
    \[
        g(x) := x^n f\!\left(\frac{1}{x}\right).
    \]
    Then
    \[
        \begin{aligned}
            g(x)
    &= x^n \sum_{i=0}^{n} a_i \left(\frac{1}{x}\right)^i \\
    &= \sum_{i=0}^{n} a_i x^{n-i} \\
    &= a_0 x^n + a_1 x^{n-1} + \cdots + a_{n-1}x + a_n .
        \end{aligned}
    \]
    Thus $g(x) \in F[x]$ and it is the polynomial obtained by 
    reversing the
    coefficients of $f$.

    Now evaluate $g$ at $1/a$:
    \[
        \begin{aligned}
            g\!\left(\frac{1}{a}\right)
    &= \left(\frac{1}{a}\right)^n f(a) \\
    &= \frac{f(a)}{a^n} \\
    &= 0 .
        \end{aligned}
    \]
    Hence $1/a$ is a root of $g$ whenever $a \neq 0$ is a root of $f$.
\end{proof}

\newpage 
\section{Homomorphisms of rings, Prime/maximal ideals}
\begin{defn}
    A \textbf{ring homomorphism} 
    \[
        \phi : R \to S
    \]
    between rings satisfies 
    \begin{align*}
        \phi(a+b) &= \phi(a) + \phi(b) \\ 
        \phi(ab) &= \phi(a)\phi(b)
    \end{align*}
    and typically 
    \[
        \phi(1_R) = 1_S
    \]
\end{defn}

\begin{thm}
    Let $R,S$ be rings and $\phi : R \to S$ a homomorphism. 
    Then $\ker \phi$ is an ideal of $R$.
\end{thm}
\begin{proof}
    Let $a,b \in \ker \phi$.

    \[
        \phi(a-b) = 0
    \]
    so closed under subtraction. 
    For any $r\in R$ we have 
    \[
        \phi(ra) = \phi(r)0 = 0
    \]
    so $ra \in \ker \phi$.

    Therefore kernels behave like normal subgroups.
\end{proof}

\begin{thm}[First Isomorphism Theorem for Rings]
    If 
    \[
        \phi : R \to S
    \]
    then 
    \[
        R/\ker \phi \cong \operatorname{Im} \phi
    \]
    meaning every homomorphic image of a ring is a quotient 
    by an ideal.
\end{thm}
\begin{ex}
    Take 
    \[
        \phi : \mathbb Z \to \mathbb Z_n
    \]
    as 
    \[
        \phi(k) = k \mod n
    \]
    The kernel is then 
    \[
        \ker \phi = n\mathbb Z
    \]
    so 
    \[
        \mathbb Z/n\mathbb Z \cong \mathbb Z_n
    \]
\end{ex}
\begin{defn}
    If $I$ is an ideal of a ring $R$ we define 
    \[
        R/I
    \]
    just like quotient groups, with operations 
    \begin{align*}
        (a+I) + (b + I) &= (a+b) + I \\ 
        (a+I)(b+I) &= ab + I 
    \end{align*}
\end{defn}

\subsection{Prime Ideals}
\begin{defn}
    An ideal $P \subset R$ is \textbf{prime} if 
    \[
        ab \in P \Rightarrow a \in P \ \text{or} \ b \in P
    \]
\end{defn}
\begin{ex}
    In $\mathbb Z$ all ideals are $(n)$ and the prime ideals are 
    $(p)$ where $p$ is a prime number. 
    If $a,b\in (p)$ then $p | ab$ so $p | a$ or $p | b$.
\end{ex}
\begin{thm}
    Suppose $P \subset R$ is an ideal.
    Then 
    \[
        P \ \text{prime} \Leftrightarrow R/P \ \text{is an integral 
        domain}
    \]
\end{thm}

\subsection{Maximal Ideals}
\begin{defn}
    An ideal $M \subset R$ is \textbf{maximal} if 
    $M \neq R$ and there is no ideal strictily between $M$ and 
    $R$, meaning 
    \[
        M \subset I \subset R
    \]
    implies $M = I$ or $R = I$.
\end{defn}
\begin{thm}
    Let $M \subset R$ be an ideal. 

    Then 
    \[
        M \ \text{maximal} \Leftrightarrow R/M \ \text{is a field}
    \]
\end{thm}

An important note is that maximal $\Rightarrow$ prime, but not 
necessarily the other way around, so being maximal is a strictly 
stronger condition. This is easy to see since a field 
is also an integral domain, but not vice versa.

\newpage 
\section{Problems problems and more problems}
\begin{prob}[Problem 26.12 Fraleigh's p.244]
    Give an example to show that a factor ring of an integral domain 
    may be a field.
\end{prob}
\begin{proof}[Solution]
    Recalling that an integral domain is a nonzero, commutative, 
    unital ring, we give the following example.

    Consider the integral domain $\mathbb Z$.
    Take some prime $p \in \mathbb Z$ and examine the factor ring 
    $\mathbb Z/\langle p\rangle$. It is known that 
    \[
        \mathbb Z/\langle p\rangle \cong \mathbb Z_p
    \]
    The integers modulo a prime is a field.
\end{proof}
\begin{prob}[Problem 26.13 Fraleigh's p.244]
    Give an example to show that a factor ring of an integral domain 
    may have $0$ divisors.
\end{prob}
\begin{proof}[Solution]
    Consider again the integers $\mathbb Z$. Take the ideal 
    generated by $6$ 
    \[
        \langle 6 \rangle = \{\ldots,-6,0,6,\ldots\}
    \]
    Notice that 
    \[
        \mathbb Z/\langle 6 \rangle \cong \mathbb Z_6
    \]
    and $2,3 \in \mathbb Z_6$, but 
    \[
        2\cdot 3 \equiv 0
    \]
\end{proof}
\begin{prob}[Problem 26.25 Fraleigh's p.244]
    Show that if $R$ is a ring with unity and $N$ is an ideal 
    of $R$ such that $N \neq R$, then $R/N$ is a ring with unity.
\end{prob}
\begin{proof}
    Let $R$ be a unital ring with ideal $N \neq R$.

    We study the quotient ring 
    \[
        R / N \neq \{N\}
    \]
    Let $(a + N) \in R / N$. Consider $(1+N) \in R/N$ 
    Then 
    \[
        (a+N)(1 + N) = (a\cdot1) + N = (1 \cdot a) + N 
        = (1 + N)(a + N)
    \]
\end{proof}
\begin{prob}[Problem 26.26 Fraleigh's p.245]
    Let $R$ be a commutative ring and let $a \in R$. 
    Show that $I_a = \{x \in R \bigm| ax = 0\}$ is an ideal of 
    $R$.
\end{prob}
\begin{proof}
    We consider, for some arbitrary element of a commutative ring $R$,
    the set of right (and correspondingly left) zero divisors of 
    $a$ 
    \[
        I_a = \{x \in R \bigm| ax = 0\}
    \]
    We know $I_a \neq \emptyset$ as $a\cdot 0 = 0$. 
    Let $b,c \in I_a$. Then $ab = ac = 0$, so
    \[
        a(b-c) = ab - ac = 0 - 0 = 0
    \]
    so $b-c \in I_a$. Therefore $(I_a,+) \leq (R,+)$. Let 
    $r \in R$. As stated $ab = 0$ and furthermore 
    \[
        a(rb) = a(br) = (ab)r = 0r = 0
    \]
    so $rb \in I_a$. Thus $I_a$ is an ideal of $R$.
\end{proof}
\begin{prob}[Problem 26.30 Fraleigh's p.245]
    An element $a$ of a ring $R$ is nilpotent if $a^n = 0$ for some 
    $n \in \mathbb Z^+$. Show that the collection of all nilpotent 
    elements in a commutative ring $R$ is an ideal, the nilradical 
    of $R$.
\end{prob}
\begin{proof}
    Let $R$ be a commutative ring and let $A \subseteq R$ be the 
    set 
    \[
        A = \{a \in R \bigm| \exists n \in \mathbb Z^+ : a^n = 0\}
    \]
    $A \neq \emptyset$ as $0^1 = 0$. Let $a,b \in A$. Then there 
    is some $m,n \in \mathbb Z^+$ such that 
    \[
        a^m = 0 \ \text{and} \ b^n = 0
    \]
    Let $k = m + n \in \mathbb Z^+$. 
    \begin{align*}
        (a-b)^k 
        &= \displaystyle\sum_{i=0}^{k}\binom{k}{i}a^i(-b)^{k-i}\\
        &= \displaystyle\sum_{i=0}^{m+n}\binom{m+n}{i}a^i(-b)^{k-i}\\
        &= \displaystyle\sum_{i=m+1}^{m+n}\binom{m+n}{i}a^i(-b)^{k-i}
        \\
        &+ \displaystyle\sum_{i=0}^{m}\binom{m+n}{i}a^{i}(-b)^{k-i}
    \end{align*}
    Notice that $\sum_{i=m+1}^{m+n}\binom{m+n}{i}a^i(-b)^{k-i}$ has 
    $i \geq m + 1$ for every term so $a^i = 0$ for all terms. 
    Hence every term is $0$. 
    For $\sum_{i=0}^{m}\binom{m+n}{i}a^i(-b)^{k-i}$ we have 
    $k-i \geq n$ so $(-b)^{k-i} = 0$ for every term. 
    Therefore 
    \[
        \displaystyle\sum_{i=m+1}^{m+n}\binom{m+n}{i}a^i(-b)^{k-i} 
        + \displaystyle\sum_{i=0}^{m}\binom{m+n}{i}a^i(-b)^{k-i} 
        = 0 + 0 = 0
    \]
    Therefore $(a-b) \in A$. 
    Now let $r \in R$ and observe that 
    \[
        (ra)^m = r^ma^m = r^m \cdot 0 = 0 
    \]
    so $ra \in A$. Thus $A$ constitutes an ideal of $R$.
\end{proof}

\begin{prob}[Problem 26.34 Fraleigh's p.245]
    Let $R$ be a commutative ring and $N$ an ideal of $R$. 
    Show that the set $\sqrt N$ of all $a \in R$, such that 
    $a^n \in N$ for some $n \geq \mathbb Z^+$, is an ideal of 
    $R$, the \textbf{radical ideal} of $N$.
\end{prob}
\begin{proof}
    Let $N$ be an ideal of some commutative ring $R$. 

    Let 
    \[
    \sqrt N = \left\{ a \in R \bigm| \exists n \in \mathbb Z^+ 
    \ \text{s.t.} \ a^n \in N\right\}
    \]
    First, observe that $\sqrt N \neq \emptyset$ as 
    $0 = 0^1 \in N$. Next, let $a,b \in \sqrt N$. We 
    show that $(a-b)^k \in N$ for some $k$, i.e. $a-b \in \sqrt N$. 
    By assumption there exists $m,n \in \mathbb Z^+$ so that 
    $a^m,b^n \in N$.  
    Let $k = m+n$. Then 
    \begin{align*}
        (a-b)^k 
        &= \displaystyle\sum_{i=0}^{k}\binom{k}{i}a^i(-b)^{k-i} \\ 
        &= \displaystyle\sum_{i=0}^{m+n}
        \binom{m+n}{i}a^i(-b)^{m+n-i}\\
        &=\displaystyle\sum_{i=m+1}^{m+n}
        \binom{m+n}{i}a^i(-b)^{m+n-i}\\
        &+ \displaystyle\sum_{i=0}^{m}\binom{m+n}{i}a^i(-b)^{m+n-i}
    \end{align*}
    The sum $\sum_{i=m+1}^{m+n}\binom{m+n}{i}a^i(-b)^{m+n-i}$ has 
    $i \geq m+1$ so each term is of the form $a^m(a \cdots a) 
    (-b)^{m+n-i}$. By assumption $a^m \in N$ and as 
    $R$ is commutative we can rearange terms and see that 
    since $N$ is an ideal of $R$ and $a\cdots a(-b)^{m+n-i} \in R$ 
    we have that each term is in $N$ by the multiplication property 
    of the ideal. As $N$ is a group under the addition, the sum 
    of these terms is in $N$. Apply this same argument to the 
    sum from $i=0$ to $m$, observing that $m+n - i \geq n$, to 
    find that $\sum_{i=0}^{m}\binom{m+n}{i}a^i(-b)^{m+n-i} \in N$. 
    Both sums are in $N$ so 
    \[
        (a-b)^k \in N
    \]
    i.e. $a-b \in \sqrt N$. 

    Lastly let $r \in R$ and see that 
    \[
        (ra)^m = r^ma^m \in N 
    \]
    as $N$ is an ideal. Thus $ra \in \sqrt N$.

    Therefore $\sqrt N$ is an ideal of $R$.
\end{proof}

\begin{prob}[Problem 27.5 Fraleigh's p.252]
    Find all $c \in \mathbb Z_3$ such that 
    \[
        \mathbb Z_3[x]/\langle x^2 + c \rangle 
    \]
    is a field.
\end{prob}
\begin{proof}
    Since $\mathbb Z_3$ is a field, the quotient
    \[
        \mathbb Z_3[x]/\langle x^2+c\rangle
    \]
    is a field if and only if $x^2+c$ is irreducible in 
    $\mathbb Z_3[x]$.

    The elements of $\mathbb Z_3$ are $0,1,2$. We check each case.

    If $c=0$, then $x^2$ is reducible since $x^2=x\cdot x$.

    If $c=1$, we check whether $x^2+1$ has a root in $\mathbb Z_3$.
    The squares in $\mathbb Z_3$ are
    \[
        0^2=0,\quad 1^2=1,\quad 2^2=1.
    \]
    Thus $x^2$ never equals $2$, so $x^2+1$ has no root and is 
    irreducible.

    If $c=2$, then
    \[
        x^2+2=x^2-1=(x-1)(x+1),
    \]
    so it is reducible.

    Therefore the quotient ring is a field only when $c=1$.
\end{proof}
\begin{prob}[Problem 27.15/16 Fraleigh's p.253]
    Find a maximal ideal of $\mathbb Z \times \mathbb Z$.

    Find a prime ideal of $\mathbb Z^2$ which is not maximal.
\end{prob}
\begin{proof}[Solution]
    We want to find the the proper ideal $M \subseteq \mathbb Z 
    \times \mathbb Z$ such that for any ideal $I$, 
    if $M \subseteq I \subseteq \mathbb Z \times \mathbb Z$ 
    then $I = \mathbb Z \times 
    \mathbb Z$ or $I = M$.

    The maximal ideals of $\mathbb Z^2$ is 
    $\langle p\rangle \times \mathbb Z$ and 
    $\mathbb Z \times \langle p\rangle$ where 
    $p$ is some prime. $\langle p \rangle^2$ does not work as the 
    quotient would be $\mathbb Z_p^2$. Both factors are nontrivial 
    so this is not a field. 
\end{proof}

\begin{prob}[Problem 18 Fraleigh's p.253]
    Is $\mathbb Q[x]/\langle x^2 - 5x + 6\rangle$ a field? Why?
\end{prob}
\begin{proof}
    Recall that the ideal generated by $x^2 - 5x + 6$ is 
    \[
        \langle x^2 -5x + 6\rangle = \{f(x) (x^2 - 5x + 6) : 
        f(x) \in \mathbb Q[x]\}
    \]
    We apply a theorem 
    \[
        \mathbb Q/\langle p \rangle \ \text{is a field} 
        \Leftrightarrow
        p \ \text{is irreducible over} \ \mathbb Q
    \]
    Notice that $x^2 - 5x + 6 = (x-2)(x-3)$ and 
    $x-2,x-3 \in \mathbb Q[x]$, so 
    \[
        \mathbb Q[x]/\langle x^2 - 5x + 6 \rangle
    \]
    is \textbf{not} a field.
\end{proof}
\begin{prob}[Problem 19 Fraleigh's p.253]
    Is $\mathbb Q[x]/\langle x^2 - 6x + 6\rangle$ a field? Why?
\end{prob}
\begin{proof}
    Notice that 
    \[
        x^2 - 6x + 6 = (x-(3-\sqrt 3))(x-(3+\sqrt 3))
    \]
    Notice that $\sqrt 3 \not\in \mathbb Q$, i.e. irreducible since 
    the polynomial is a quadratic, 
    so this quotient ring 
    is a field. ($\cong \mathbb Q(\sqrt 3)$)
\end{proof}

\newpage 
\section{Tackling some extension field problems early}
\begin{prob}[Problem 29.1/3/4 Fraleigh's p.272]
    For each following $\alpha$, show that the given number 
    $\alpha \in \mathbb C$ is algebraic over $\mathbb Q$ by finding 
    $f(x) \in \mathbb Q$ such that $f(\alpha) = 0$. 

    \[
        \alpha = 1+\sqrt 2
    \]
    \[
        \alpha = 1+i
    \]
    \[
        \alpha = \sqrt{1+\sqrt[3]{2}}
    \]
\end{prob}
\begin{proof}[Solutions]
    We begin by considering $\alpha = 1+\sqrt 2$. 
    Let $x = 1+\sqrt 2$. Then 
    \begin{align*}
        x &= 1+\sqrt 2 \\
        \sqrt 2 &= x - 1 \\ 
        2 &= (x-1)^2 \\ 
        x^2 -2x + 1 &= 2 \\ 
        x^2 -2x - 1 &= 0
    \end{align*}
    Thus $x^2 - 2x - 1 \in \mathbb Q$ is zero for 
    $x = \alpha = 1 + \sqrt 2$, hence $1 + \sqrt 2$ is algebraic over
    $\mathbb Q$.

    Next we look at $\alpha = 1+i$.
    Set $x = 1+i$.
    \begin{align*}
        i &= x - 1 \\ 
        -1 &= (x-1)^2 \\ 
        -1 &= x^2 - 2x + 1 \\
        x^2 -2x +2 &= 0
    \end{align*}
    $f(x) = x^2 - 2x + 2 \in \mathbb Q[x]$ and 
    $f(\alpha) = 0$ so $\alpha = 1 + i$ is algebraic over 
    $\mathbb Q$.

    Lastly let $\alpha = \sqrt{1+\sqrt[3]{2}}$.
    \begin{align*}
        x &= \sqrt{1+\sqrt[3]{2}} \\ 
        x^2 &= 1 + 2^{1/3} \\ 
        x^2 - 1 &= 2^{1/3} \\ 
        (x^2-1)^3 &= 2 \\ 
        (x^2-1)^3 - 2 &= 0
    \end{align*}
    As above ... $\alpha$ is algebraic over $\mathbb Q$.
\end{proof}

\begin{prob}[Problem 29.18 Fraleigh's p.272]
    a. Show that $x^2 + 1$ is irreducible in $\mathbb Z_3[x]$.

    b. Let $\alpha$ be a zero of $x^2 + 1$ in an extension field of 
    $\mathbb Z_3$. As in Example 29.19, give the multiplication and 
    additions tables for the nine elements of $\mathbb Z_3[\alpha]$, 
    written in the order $0,1,2,\alpha,2\alpha,3\alpha, 1 + \alpha, 
    1 + 2\alpha, 2 + \alpha$ and $2 + 2 \alpha$.
\end{prob}
\begin{proof}[Proof of (a)]
    Let $f(x) = x^2 + 1$. $f(x)$ is a classical introductory example 
    to motivate the complex numbers, having 
    $\alpha = \pm i \in \mathbb C\setminus{\mathbb R}$ as zero.  

    As $f$ is a degree $2$ polynomial with no zeros in $\mathbb Z_3$ 
    we conclude that it is irreducible in $\mathbb Z_3[x]$.
\end{proof}

\newpage 
\section{Trying an earlier exam}
\begin{prob}[Problem 1a, Exam 2023]
    Prove that the set 
    \[
        S = \{2^n \bigm| n \in \mathbb Z\}
    \]
    equipped with the usual operation of multiplication is a group.
\end{prob}
\begin{proof}
    It suffices to show that $(S,\cdot)$ is isomorphic to a known 
    group. Let $\phi : \mathbb Z \to S$ be defined by 
    \[
        \phi(n) = 2^n
    \]
    It is clear that $\phi$ is a bijection as the preimage of 
    $2^n$ is $n$ and $2^n = 2^m$ iff $n = m$. 

    Lastly we show $\phi$ to be a homomorphism. 
    \begin{align*}
        \phi(n+m) &= 2^{n+m} \\ 
                  &= 2^n \cdot 2^m \\ 
                  &= \phi(n)\phi(m)
    \end{align*}
    Thus $(S,\cdot)$ is isomorphic to $(\mathbb Z, +)$ 
    which is a known group.
\end{proof}

\begin{prob}[Problem 1b, Exam 2023]
    List all finite abelian groups of order $20$, up to isomorphism.
\end{prob}
\begin{proof}[Solution]
    The finite abelian groups of order $20$ are trivially 
    finitely generated so the fundamental theorem of finitely 
    generated abelian groups applies. 

    I.e. 
    \[
        G \cong \left(\prod_{i=1}^n 
        \mathbb Z_{p_i^{e_i}}\right) \times 
        \mathbb Z^k
    \]
    We are asked to find those of order $20$ (finite) so 
    \[
        G \cong \prod_{i=1}^n \mathbb Z_{p_i^{e_i}}
    \]
    where $\prod_{i=1}^n p_i^{e_i} = 20$. 

    $20 = 4\cdot5 = 2\cdot2\cdot 5$. Thus there exists exactly 
    two abelian groups of order $20$, namely 
    \[
        \mathbb Z_4 \times \mathbb Z_5
    \]
    and
    \[
        \mathbb Z_2 \times \mathbb Z_2 \times \mathbb Z_5
    \]
\end{proof}
\begin{prob}[Problem 1c, Exam 2023]
    Let $A_4$ be the alternating group on $4$ letters, that is 
    the subgroup of $S_4$ consisting of even permutations. 
    Find elements $\sigma_1,\sigma_2,\sigma_3 \in A_4$ such that 
    $\sigma_i$ has order $i$ (for $i = 1,2,3$). 
\end{prob}
\begin{proof}[Solution]
    $\sigma_1$ is the trivial permutation, i.e. the one which sends 
    \[
        1 \mapsto 1, 2 \mapsto 2, \ \text{etc.}
    \]
    Next a candidate for $\sigma_2$ is any permutation which is 
    the composition of two transpositions, e.g. 
    \[
        (1 \ 2)(3 \ 4)
    \]
    Notice that 
    \[
        (1 \ 2)(3 \ 4) \circ (1 \ 2)(3 \ 4) 
    \]
    sends 
    \[
        1 \to 2 \to 1, 2 \to 1 \to 2, 3 \to 4 \to 3, 4 \to 3 \to 4
    \]
    So $(1 \ 2)(3 \ 4)$ has order $2$. 

    An exampele of $\sigma_3$ is something like  
    \[
        (1 \ 2 \ 3) = (1 \ 3)(1 \ 2)
    \]
\end{proof}

\begin{prob}[Problem 1d, Exam 2023]
    Prove that $A_4$ is not abelian.
\end{prob}
\begin{proof}
    Observe that $(1 \ 2)(3 \ 4)$ and $(1 \ 3)(1 \ 2)$ do not 
    commute.

    \[
        (1 \ 2)(3 \ 4) \circ (1 \ 3)(1 \ 2)  
    \]
    sends 
    \begin{align*}
        1 \mapsto 2 \mapsto 1 \\ 
        2 \mapsto 3 \mapsto 4 \\ 
        3 \mapsto 1 \mapsto 2 \\ 
        4 \mapsto 4 \mapsto 3
    \end{align*}
    So 
    \[
        (1 \ 2)(3 \ 4) \circ (1 \ 3)(1 \ 2)  
        = (2 \ 4 \ 3)
    \]
    while 
    \[
        (1 \ 3)(1 \ 2) \circ (1 \ 2)(3 \ 4)
    \]
    sends 
    \begin{align*}
        3 \mapsto 4 \mapsto 4 \\
    \end{align*}
    So we can already see that 
    \[
        (1 \ 3)(1 \ 2) \circ (1 \ 2)(3 \ 4) 
        \neq 
        (1 \ 2)(3 \ 4) \circ (1 \ 3)(1 \ 2)
    \]
    Hence $A_4$ is not abelian.
\end{proof}

\newpage 
\begin{prob}[Problem 2a, Exam 2023]
    Determine the $0$-divisors in the ring $\mathbb Z_{10}$.
\end{prob}
\begin{proof}[Solution]
    $\mathbb Z_{10}$ is commutative so left- and right 
    divisors coincide. 

    \begin{align*}
        (2)(5) &= 10 \equiv 0 \\ 
        (4)(5) &= 20 \equiv 0 \\ 
        (6)(5) &= 30 \equiv 0 \\ 
        (8)(5) &= 40 \equiv 0 \\ 
    \end{align*}
    so the $0$-divisrors in $\mathbb Z_{10}$ are 
    $5$ with any power of $2$, which just so happens to be 
    the nonzero elements of $\mathbb Z_{10}$ which are not coprime to 
    $\mathbb Z_{10}$.
\end{proof}

\begin{prob}[Problem 2b, Exam 2023]
    Let $R$ be the ring $\mathbb Z \times \mathbb Z$, where 
    addition and multiplication is given by, for $(a,b),(c,d)\in 
    \mathbb Z \times \mathbb Z$, setting 
    \begin{align*}
        (a,b)+(c,d)&=(a+c,b+d) \\ 
        (a,b)(c,d)&=(ac,bd)
    \end{align*}
    Let 
    \[
        I = \{(a,0) | a \in \mathbb Z\}\subset R
    \]
    Prove that $I$ is an ideal of $R$.
\end{prob}
\begin{proof}
    We show that $I$ is a subgroup of $(R,+)$ with the 
    proper behavior w.r.t left-multiplication.

    First we begin by observing that 
    $(0,0) \in I \neq \emptyset$. 
    Let $(a,0),(b,0) \in I$. 
    Then we see that 
    \begin{align*}
        (a,0)-(b,0) &= (a,0)+(-b,0) \\ 
                    &= (a-b, 0 + 0) \\ 
                    &= (a-b,0) \in I
    \end{align*}
    So $I$ is a subgroup of the additive group of $R$. 

    Lastly let $(c,d) \in R$. 
    Then 
    \begin{align*}
        (c,d)(a,0) &= (ca,d0) \\ 
                   &= (ca,0) \in I
    \end{align*}
    Thus $I$ is an ideal of $R$.
\end{proof}
\begin{prob}[Problem 2c, Exam 2023]
    Prove that $I$ is a prime ideal and that $I$ is not 
    a maximal ideal. 
\end{prob}
\begin{proof}
    It is clear that $I \neq R$ as $(1,1) \in R \setminus I$. 

    Now suppose $(ab,0) \in I$. 
    Then 
    \begin{align*}
        (ab,0) &= (a,c)(b,d) 
    \end{align*}
    with $a,b,c,d \in \mathbb Z$ and $cd = 0$. 
    In $\mathbb Z$ there are no zero-divisors, thus 
    \[
        c = 0 \ \text{or} \ d = 0
    \]
    and hence $(a,c) \in I$ or $(b,d) \in I$.

    We show that $I$ is not maximal by giving some ideal $J$ so that 
    \[
        I \subseteq J \subseteq R
    \]
    with $I \neq J$ and $J \neq R$.

    Let $J$ be the set of $(a,b) \in R$ so that 
    $b$ is even. Formally 
    \[
        J = \{(a,b) \in R \bigm| b = 2n \ \text{for some} \ n 
        \in \mathbb Z\}
    \]
    It is trival to check that this is an ideal and 
    that 
    \[
        I \subsetneq J \subsetneq R
    \]
\end{proof}

\newpage
\begin{prob}[Problem 3a, Exam 2023]
    Let $p$ be a prime number, and let 
    $f(x) = x^p - x \in \mathbb Z_p[x]$.
    Determine (with justification) the zeros of 
    $f(x)$ in $\mathbb Z_p[x]$.
\end{prob}
\begin{proof}[Solution]
    The zeros are all numbers in $\mathbb Z_p$. 
    We can apply Fermat's little theorem 
    \[
        a^p \equiv a \mod p
    \]
    to get that 
    \[
        f(x) = x^p - x^p = x-x = 0
    \]
\end{proof}
\begin{prob}[Problem 3b, Exam 2023]
    Consider $x^2 + 1 \in \mathbb Z_3[x]$. Prove that the ring 
    $\mathbb Z_3[x]/\langle x^2 + 1\rangle$ is a field, and 
    determine its number of elements. 
\end{prob}
\begin{proof}
    We use the following theorem: 
    
    For some field $F$ and $f \in F[x]$ 
    \[
        F[x]/\langle f(x) \rangle
    \] 
    is a field if and only if $f(x)$ is irreducible in $F[x]$. 
    Since $f(x) = x^2 + 1$ is a degree $2$ polynomial we see 
    that $f$ is irreducible in $\mathbb Z_3[x]$ if and only if 
    $f$ has no zeros in $\mathbb Z_3[x]$. 

    It is known that $f(x)$ is irreducible since it is the motivating
    introductory example for the complex numbers with solution 
    $x = \pm i$. 

    We quickly check that $f(0) \neq 0$, $f(1) \neq 0$ and 
    $f(2) \neq 0$ to confirm this. 

    As $3$ is prime we see that $\mathbb Z_3$ is a field. 

    We conclude therefore that 
    \[
        \mathbb Z_3[x]/\langle x^2 + 1\rangle
    \]
    is a field.
\end{proof}

\newpage 
\section{Extension fields \& Vector spaces}
\begin{defn}
    A \textbf{field extension} of a field $F$ is a field 
    $K$ with $F \subseteq K$.

    It is common to write 
    \[
        K / F
    \]
\end{defn}
Some basic examples of the above include $\mathbb Q \subseteq 
\mathbb R$, $\mathbb Q \subseteq \mathbb C$. 

As we have seen earlier, 
\[
    \mathbb Z_3 \subseteq \mathbb Z_3[x]/\langle x^2 + 1\rangle 
    \cong \mathbb Z_3(i)
\]

\begin{defn}
    Given a field $F$ and an element $\alpha \not\in F$, define 
    \[
        F(\alpha)
    \]
    to be the smallest field containing both $F$ and $\alpha$.
\end{defn}
\begin{thm}
    If $\alpha$ satisfies a polynomial over $F$, then 
    \[
        F(\alpha) \cong F[x]/\langle f(x) \rangle
    \]
    where $f$ is the minimal polynomial.

    In other words, if $\alpha$ is algebraic over $F$ then this 
    holds.
\end{thm}
\begin{defn}
    The minimal polynomial of $\alpha$ over $F$ is  
    the monic polynomial of smallest degree in $F[x]$ such that 
    $\alpha$ is a zero of said polynomial.
\end{defn}
\begin{defn}
    The degree of an extension $[K : F]$ is the dimension of 
    $K$ as a vector space over $F$.
\end{defn}
\begin{ex}
    $[\mathbb Q(\sqrt 2) : \mathbb Q] = 2$ with basis 
    $\{1, \sqrt2\}$.
\end{ex}
\begin{thm}
    If $K/F$ is a field extension then $K$ is a vector space over 
    $F$.
\end{thm}
\begin{thm}
    If minimal polynomial has degree $n$ then 
    \[
        [F(\alpha) : F] = n
    \]
\end{thm}
\begin{thm}[Tower Law]
    If 
    \[
        F \subseteq K \subseteq L
    \]
    then 
    \[
        [L : F] = [L : K][K : F]
    \]
\end{thm}

\newpage 
\begin{prob}[Problem 4a, Exam 2023]
    Let $f(x) = x^5 - 8x^4 + 2x + 2 \in \mathbb Q[x]$. You may use 
    w/o proof that $f$ is irreducible. 

    Let $\alpha \in \overline{\mathbb Q}$ be a zero of 
    $f$. Compute $[\mathbb Q(\alpha) : \mathbb Q]$, and give 
    a basis for $\mathbb Q(\alpha)$ as a vector space over 
    $\mathbb Q$.
\end{prob}
\begin{proof}
    Since $f(x) \in \mathbb{Q}[x]$ is irreducible and $\alpha$ 
    is a root,
    $f$ is the minimal polynomial of $\alpha$ over $\mathbb{Q}$.

    Therefore,
    \[
        [\mathbb{Q}(\alpha) : \mathbb{Q}] = \deg f = 5.
    \]

    A basis for $\mathbb{Q}(\alpha)$ over $\mathbb{Q}$ is
    \[
        \{1,\alpha,\alpha^2,\alpha^3,\alpha^4\}.
    \]
\end{proof}
\begin{prob}[Problem 4b, Exam 2023]
    Let $\beta \in \mathbb Q(\alpha) \setminus \mathbb Q$. 
    Prove that $\mathbb Q(\beta) = \mathbb Q(\alpha)$.
\end{prob}
\begin{proof}
    $\beta \in \mathbb Q(\alpha)$ and $\beta \not \in \mathbb Q$, 
    by assumption. So 
    \[
        \mathbb Q \subseteq \mathbb Q(\beta) \subseteq 
        \mathbb Q(\alpha)
    \]
    We know that 
    \[
        [\mathbb Q(\alpha) : \mathbb Q] = 
        [\mathbb Q(\alpha) : \mathbb Q(\beta)] \cdot 
        [\mathbb Q(\beta) : \mathbb Q]
    \]
    $5$ is prime so one of the factors is $5$ and the other is $1$. 
    It can't be the case that 
    \[
        [\mathbb Q(\alpha) : \mathbb Q(\beta)] = 5
    \]
    since that would imply $\mathbb Q(\beta) = \mathbb Q$ which 
    is not the case, by assumption. Hence 
    $\mathbb Q(\alpha) = \mathbb Q(\beta)$.
\end{proof}

\newpage 
\section{Some problems relevant to Fraleigh's sections 29-33}
\begin{prob}
    Let $\alpha = \sqrt 2 + \sqrt 5$. Find the minimal 
    polynomial for $\alpha$ over $\mathbb Q$.
    Then compute degree and identify a basis.
\end{prob}
\begin{proof}[Solution]
    \begin{align*}
        x &= \sqrt 2 + \sqrt 5 \\ 
        x^2 &= 7 + 2\sqrt{10} \\ 
        x^2 - 7 &= 2\sqrt{10} \\ 
        \frac{x^2-7}{2} &= \sqrt{10} \\ 
        \frac{(x^2-7)(x^2-7)}{4} &= 10 \\ 
        \frac{x^4 -14x^2 + 49 - 40}{4} &= 0 \\ 
        \frac{x^4 - 14x^2 + 9}{4} &= 0 \\ 
        x^4 - 14x^2 + 9 &= 0 \\ 
    \end{align*}
    is the minimal polynomial of $\alpha$. 
    It is degree $r$ and thus a basis of 
    $Q(\alpha)$ over $\mathbb Q$ is 
    $\{1, \alpha, \alpha^2, \alpha^3\}$.
\end{proof}
\begin{prob}
    Give an example of a field extension which is finitely generated 
    and not finite.
\end{prob}
\begin{proof}[Solution]
    Consider $\mathbb C/\mathbb R$. $\mathbb C$ is a 
    $2$-dimensional vector space over $\mathbb R$, but 
    as a field $|\mathbb C| = \aleph_1$ and 
    \[
        |\mathbb C\setminus \mathbb R| = \aleph_1
    \]
\end{proof}
\begin{prob}
    Let $\alpha, \beta$ be algebraic over $F$. 
    Prove 
    \[
        [F(\alpha,\beta):F] \leq [F(\alpha):F][F(\beta):F]
    \]
    and explain when equality is achieved.
\end{prob}
\begin{proof}
    Suppose $\alpha,\beta$ algebraic over some field $F$.
    Then there exists minimal polynomials 
    \[
        f,g \in F[x]
    \]
    so that $f(\alpha) = 0$ and $g(\beta) = 0$. 
    With 
    \[
        F(\alpha) = 
        \left\{
        \displaystyle\sum_{i=0}^{\deg f - 1}c_i\alpha_i \biggm|
        c_i \in F
        \right\}
    \]
    and 
    \[
        F(\beta) = 
        \left\{
        \displaystyle\sum_{i=0}^{\deg g - 1}c_i\beta_i \biggm|
        c_i \in F
        \right\}
    \]
    Then 
    \[
        [F(\alpha):F][F(\beta):F] = (\deg f-1)(\deg g-1)
    \]
    It is clear that the two expressions are equal when 
    $g$ is irreducible in $F(\alpha)$, recalling 
    \[
        F \subseteq F(\alpha) \subseteq F(\alpha,\beta)
    \]
    A nice example is $\mathbb Q(\sqrt 2, \sqrt 5)$.

    $F(\alpha,\beta)$ is an extension field of $F(\alpha)$ so if 
    $\beta$ and $\alpha$ are independent in their minimal polynomials 
    then we get equality. 
\end{proof}

\begin{prob}[Problem 31.3/5/7 Fraleigh's p.291]
    Find the degree and basis for the given field extensions. 
    Be prepared to justify your answers. 

    3. $\mathbb Q(\sqrt 2, \sqrt 3, \sqrt{18})$ over $\mathbb Q$ 

    5. $\mathbb Q(\sqrt 2, \sqrt[3]2)$ over $\mathbb Q$ 

    7. $\mathbb Q(\sqrt 2, \sqrt 3)$ over $\mathbb Q$
\end{prob}
\begin{proof}[Solution]
    Notice that 
    \[
        \mathbb Q(\sqrt 2, \sqrt 3, \sqrt{18}) 
        = \mathbb Q(\sqrt 2, \sqrt 3)
    \]
    since $\sqrt{18} = \sqrt{9}\sqrt 2 = 3 \sqrt 2 \in \mathbb 
    Q(\sqrt 2)$.
    Now notice that 
    \[
        [\mathbb Q(\sqrt 2) : \mathbb Q] = 2 
    \]
    since $x^2 - 2$ is irreducible over $\mathbb Q$.
    Lastly we conclude that $\sqrt 3 \not \in \mathbb Q(\sqrt 2)$ 
    since 
    \begin{align*}
        \sqrt 3 &= a + b\sqrt 2 \\ 
        3 &= a^2 + 2ab\sqrt 2 + 2b^2
    \end{align*}
    Thus $2ab = 0$ so $a = 0$ or $b = 0$. Both cases work out 
    to show that $a,b \not\in \mathbb Q$. Therefore 
    \[
        [\mathbb Q(\sqrt 2, \sqrt 3) : \mathbb Q(\sqrt 2)] = 2
    \]
    By the tower law: 
    \[
        [\mathbb Q(\sqrt 2, \sqrt 3) : \mathbb Q] = 4
    \]
    with a standard basis being 
    $\{1,\sqrt 2, \sqrt 3, \sqrt 6\}$.

    Next up we examine $\mathbb Q(\sqrt 2, \sqrt[3] 2)$.

    \begin{align*}
        \sqrt[3]2 &= a + b\sqrt 2 \\ 
        2 &= (a+b\sqrt 2)(a^2 + 2b^2 + 2ab\sqrt 2) \\ 
        2 &= a^3 + 2ab^2 + 2a^2b\sqrt 2 + a^2b\sqrt 2 + 
        2b^3 \sqrt 2 + 2ab^2
    \end{align*}
    The left side is rational so we require $b = 0$, but then 
    $a = \sqrt[3] 2 \not\in \mathbb Q(\sqrt 2)$. Hence we conclude 
    that 
    \[
        [\mathbb Q(\sqrt 2, \sqrt[3]2) : \mathbb Q] = 6
    \]

    For problem 7 notice that we have already shown it for 
    problem 3.
\end{proof}
\begin{prob}[Problem 31.23 Fraleigh's]
    Show that if $E$ is a finite extension of a field 
    $F$ and $[E:F]$ is prime, then $E$ is a simple 
    extension of $F$ and, indeed $E = F(\alpha)$ for 
    every $\alpha \in E\setminus F$.
\end{prob}
\begin{proof}
    We assume $E$ is a finite extension of $F$ and in particular 
    \[
        [E:F] = p
    \]
    for some prime $p$. 
    Let $\alpha \in E\setminus F$ then $\alpha$ is algebraic 
    over $F$ so $\alpha$ has a minimal polynomial over 
    $F$ and 
    \[
        [F(\alpha) : F] = \deg m_\alpha(x)
    \]
    Now note that 
    \[
        F \subseteq F(\alpha) \subseteq E
    \]
    since $\alpha \in E$. By the tower law: 
    \[
        [E:F] = [E:F(\alpha)][F(\alpha):F] = p
    \]
    so one side of the product must be $1$. Since $\alpha \not\in F$
    it must be the case that 
    \[
        [F(\alpha) : F] > 1
    \]
    so $[F(\alpha):F] = p$, i.e. 
    \[
        E = F(\alpha)
    \]
\end{proof}
\begin{prob}[Problem 31.24 Fraleigh's p.292]
    Prove that $x^2 - 3$ is irreducible over 
    $\mathbb Q(\sqrt[3]2)$.
\end{prob}
\begin{proof}
    We can make the simple observation that 
    $\sqrt 3 \not\in \mathbb Q(\sqrt[3]2)$.
    Let $\alpha = \sqrt[3]2$. Then 
    \[
        \mathbb Q(\alpha) = \{a+b\alpha + c\alpha^2 : a,b,c \in 
        \mathbb Q\}
    \]
    So if $\sqrt 3 \in \mathbb Q(\alpha)$ then 
    \begin{align*}
        \sqrt 3 &= a + b\alpha + c\alpha^2
    \end{align*}
    Suppose this holds. Since $\mathbb Q(\sqrt 3) \subseteq 
    \mathbb Q(\alpha)$ then by the tower law 
    \[
        [\mathbb Q(\alpha)] = [\mathbb Q(\alpha) : \mathbb Q(\sqrt 3)]
        \cdot [\mathbb Q(\sqrt 3) : \mathbb Q]
    \]
    But then 
    \[
        3 = [\mathbb Q(\alpha) : \mathbb Q(\sqrt 3)] \cdot 2
    \]
    but $2$ does not divide $3$ so this is impossible.
\end{proof}
\begin{prob}[Problem 31.29 Fraleigh's p.292]
    Let $E$ be a finite extension of $F$, and let 
    $p(x) \in F[x]$ be irreducible over $F$ and have degree 
    that is not a divisor of $[E:F]$. Show that $p(x)$ has 
    no zeros in $E$.
\end{prob}
\begin{proof}
    Let $\alpha \in E$ be a hypothetical zero of $p(x)$.
    Since $p$ is irreducible in $F$ and $p(\alpha) = 0$, 
    $p$ must be the minimal polynomial of $\alpha$.

    Thus 
    \[
        [F(\alpha) : F] = \deg p(x) 
    \]
    Using the tower law 
    \[
        [E:F] = [E:F(\alpha)] [F(\alpha) : F]
    \]
    But then $\deg p$ divides $[E:F]$ which, by assumption, 
    is not the case. Hence we have a contradiction.
\end{proof}
\begin{prob}[Problem 31.30 Fraleigh's p.292]
    Let $E$ be an extension of $F$. Let $\alpha \in E$ 
    be algebraic of odd degree over $F$. Show that $\alpha^2$ 
    is algebraic of odd degree over $F$, and $F(\alpha) = 
    F(\alpha^2)$.
\end{prob}
\begin{proof}
    We fix some $\alpha \in E$ algebraic of odd degree over $F$.
    Then 
    \[
        [F(\alpha) : F] = 2n + 1, \ n \in \mathbb N 
    \]
    Crucially $\alpha \not \in F$, by assumption.
    Now suppose $\alpha^2 \in F$. 
    Then $\alpha^2 \not\in F(\alpha) \setminus F$ so 
    $[F(\alpha):F] = 2$ (since $x^2 - \alpha^2 \in F[x]$ would be a 
    minmal polynomial of $\alpha$), which contradicts our assumption. 
    Thus $\alpha^2 \not\in F$. 
    Now since $\alpha^2 \in F(\alpha)$ we see that 
    $\alpha$ satisfies 
    \[
        x^2 - \alpha^2 \in F(\alpha^2)[x]
    \]
    Hence $[F(\alpha) : F(\alpha^2)] \leq 2$. By the tower law 
    \[
        [F(\alpha) : F] = [F(\alpha) : F(\alpha^2)][F(\alpha^2):F]
    \]
    We see that then $[F(\alpha) : F(\alpha^2)]$ must be $1$, 
    hence 
    \[
        F(\alpha) = F(\alpha^2)
    \]
\end{proof}
\begin{prob}
    Let $\alpha = \sqrt 2 + \sqrt 3$.

    Show that $\alpha$ is algebraic over $\mathbb Q$.

    Show that $\mathbb Q(\alpha) = \mathbb Q(\sqrt 2, \sqrt 3)$.

    Deduce that $\mathbb Q(\sqrt 2, \sqrt 3)$ is a simple extension.
\end{prob}
\begin{proof}
    Clearly $\sqrt 2 + \sqrt 3 \not\in \mathbb Q$.
    Now it remains to find a polynomial in $\mathbb Q[x]$ for 
    which $\alpha$ is a zero.
    \begin{align*}
        x &= \sqrt 2 + \sqrt 3 \\ 
        x^2 &= 2 + 2\sqrt 2 \sqrt 3 + 3 \\ 
        x^2 - 5 &= 2 \sqrt 2 \sqrt 3 \\ 
        \frac{x^2 - 5}{2} &= \sqrt 6 \\ 
        \frac{(x^2 - 5)(x^2 - 5)}{4} &= 6 \\ 
        \frac{x^4 - 10x^2 + 25}{4} - 6 &= 0 \\ 
        x^4 - 10x^2 +1 &= 0
    \end{align*}
    and $x^4 - 10x^2 + 1 \in \mathbb Q[x]$ so 
    $\alpha$ is algebraic over $\mathbb Q$.

    Thus 
    \[
        [\mathbb Q(\alpha) : \mathbb Q] = [\mathbb Q(\sqrt 2, 
        \sqrt 3): \mathbb Q] = 4
    \]
    The standard basis for $\mathbb Q(\sqrt 2, \sqrt 3)$ is 
    $\{1,\sqrt 2, \sqrt 3, \sqrt 2 \sqrt 3\}$ which just so 
    happens to be the basis for $\mathbb Q(\alpha)$.
    We can also take some arbitrary element from either extension 
    field and show how to express it in terms of the other to show 
    that they are indeed equal.

    Since $\mathbb Q(\alpha)$ is simple, it follows from the above 
    observation that 
    \[
        \mathbb Q(\sqrt 2, \sqrt 3)
    \]
    is indeed simple.
\end{proof}

\section{Some problems from Section 33}
\begin{prob}[Problem 33.4 Fraleigh's]
    Find the number of primitive $8$th roots of unity 
    in $GF(9)$ (also known as $\mathbb F_9$).
\end{prob}
\begin{proof}
    The multiplicative group 
    $\mathbb F^x_9$ has order $8$ since $0$ is excluded.
    The multiplicative group of a finite field is 
    cyclic so 
    \[
        \mathbb F_9^x \cong \mathcal C_8
    \]
    The $8$th roots of unity are the elements $x \in \mathcal C_8$
    for which 
    \[
        x^8 = 1
    \]
    Since the whole group has order $8$, every nonzero element 
    satisfies this. A primitive $8$th root of unity 
    has order exactly $8$ and the Euler totient function 
    gives 
    \[
        \phi(8) = 4
    \]
\end{proof}
\begin{prob}[Problem 33.9 Fraleigh's p.305]
    Let $\overline{\mathbb Z}_2$ be the algebraic closure 
    of $\mathbb Z_2$, and let 
    $\alpha,\beta \in \overline{\mathbb Z}_2$ be zeros 
    of $x^3 + x^2 + 1$ and of $x^3 + x + 1$, respectively.
    Using the results of this section, show that 
    $\mathbb Z_2(\alpha) = \mathbb Z_2(\beta)$.
\end{prob}
\begin{proof}
    We begin by showing that the polynomials 
    $f(x) = x^3 + x^2 + 1$ and $g(x) = x^3 + x + 1$ are 
    irreducible over $\mathbb Z_2$.
    These are degree $3$ polynomials and $\mathbb Z_2$ is a field 
    so they are irreducible if and only if 
    they have no zeros in $\mathbb Z_2$.
    \[
        f(0) = 1 \ \text{and} \ f(1) = 1,
    \]
    and 
    \[
        g(0) = 1 \ \text{and} \ g(1) = 1.
    \]
    Thus $f,g$ are irreducible over $\mathbb Z_2$. 
    Hence 
    \[
        [\mathbb Z_2(\alpha) : \mathbb Z_2] 
        = 
        [\mathbb Z_2(\beta) : \mathbb Z_2] 
        = 3
    \]
    So both $\mathbb Z_2(\alpha)$ and $\mathbb Z_2(\beta)$ 
    are field with degree $2^3$. Since $2$ is prime we can 
    use the finite-field theorem and conclude that 
    these two fields are isomorphic. Furthermore 
    in an algebraic closure of some field $\mathbb F_p$, there 
    is exactly one subfield with $p^n$ elements, namely 
    the set of roots of 
    \[
        x^{p^n} - x.
    \]
    Thus 
    \[
        \mathbb Z_2(\alpha) = \mathbb Z_2(\beta).
    \]
\end{proof}

\begin{prob}[Problem 33.10 Fraleigh's p.305]
    Show that every irreducible polynomial in 
    $\mathbb Z_p[x]$ is a dvisor of 
    $x^{p^n} - x$ for some $n$.
\end{prob}
\begin{proof}
    Let $\alpha \in \overline{\mathbb Z}_p$ be the root of 
    some $f(x) \in \mathbb Z_p$, irreducible. 
    Then 
    \[
        [\mathbb Z_p(\alpha) : \mathbb Z_p] = \deg f
    \]
    So $\mathbb Z_p(\alpha)$ is a finite field with 
    $p^m$ elements.
    If a field has $p^m$ elements then 
    $a^{p^m} = a$ for all $a$ in said field.
    Equivalently every elements is a root of 
    \[
        x^{p^m} - x.
    \]
    Then 
    \[
        \alpha^{p^m} - \alpha = 0.
    \]
    If a polynomial $g$ is the minimal polynomial of $\alpha$ 
    over $\mathbb Z_p$, and $\alpha$ is also a root of another 
    polynomial $f$ then 
    \[
        g \bigm| f
    \]
\end{proof}
\begin{prob}[Problem 3a Exam 2016]
    What is an integral domain? Show that 
    $\mathbb Z[x]/(x^2 + 1)$ is an integral domain. 
    Is $\mathbb Z[x]/(x^2 - 1)$ an integral domain?
    Give reasons for your answer.
\end{prob}
\begin{proof}[Solution and Proof]
    An integral domain is a non-zero ring with multiplicative 
    identity ($1 \neq 0$) that has no zero divisors.

    For $f(x) \in F[x]$, $F[x]/\langle f(x) \rangle$ is 
    an integral domain iff $f(x)$ is irreducible in 
    $F[x]$.
    Since $x^2 + 1$ is a degree $2$ polynomial it is 
    irreducible if and only if it has no zeros in 
    $\mathbb Z$. $f(x) = x^2 + 1$ is famously a motivating example 
    for complex numbers with zeros $\pm i \in \mathbb C$. 
    A quick sanity check gives $f(0) = 1$, $f(\pm 1) = 2$ 
    and generally $f(n) = f(-n) < f(n + 1) = f(-(n + 1))$ so 
    it ceirtainly does not have any zeros in $\mathbb Z$ (can 
    be proven with induction or standard techniques if necessary).
    Thus $f$ is irreducible and so $\mathbb Z[x]/\langle f(x) 
    \rangle$ is an integral domain.

    Using this same procedure we see that 
    $x^2 - 1$ has zeros $\pm 1 \in \mathbb Z$ and is therefore 
    not irreducible so 
    \[
        \mathbb Z[x]/\langle x^2 - 1 \rangle
    \]
    is not an integral domain.
\end{proof}

\newpage 
\section{Automorphisms of Field Extensions \& 
Isomorphism Extension}
\begin{defn}[Field Automorphism]
    A \textbf{field automorphism} of $E$ is a bijection 
    \[
        \sigma : E \to E
    \]
    such that $\forall x,y \in E$: 
    \begin{enumerate}
        \item $\sigma (x+y) = \sigma(x) + \sigma(y)$ 
        \item $\sigma(xy) = \sigma(x)\sigma(y)$
        \item $\sigma(1) = 1$
    \end{enumerate}
\end{defn}
If $F \subseteq E$ is a field extension we write 
\[
    \operatorname{Aut}(E/F) := 
    \{\sigma : E \to E \bigm| \sigma 
        \ \text{is a field automorphism and} \ 
    \sigma|_F = \operatorname{id}_F\}.
\]

\begin{prop}
    A field automorphism $\sigma$ preserves algebraic 
    structure: 
    \begin{enumerate}
        \item $\sigma(0) = 0$ 
        \item $\sigma(-x) = -\sigma(x)$ 
        \item $\sigma(x^{-1}) = \sigma(x)^{-1}$ for 
            $x \neq 0$
    \end{enumerate}
\end{prop}
\begin{proof}[Proof Sketch]
    Straightforward from homomorphism 
    properties. For instance: 
    \[
        1 = \sigma(1) = \sigma(xx^{-1}) 
        = \sigma(x)\sigma(x^{-1})
    \]
    and so on.
\end{proof}

\begin{prop}
    $\operatorname{Aut}(E/F)$ is a group under composition.
\end{prop}

\begin{thm}
    Let $E/F$ be an extension, and let $\alpha \in E$ be 
    algebraic over $F$. Let $m_\alpha(x) \in F[x]$ be a 
    minimal polynomial. 

    Then for any $\sigma \in \operatorname{Aut}(E/F)$, 
    \[
        m_\alpha(\sigma(\alpha)) = 0.
    \]
\end{thm}
\begin{proof}
    Write $m_\alpha(x) = \sum_{i=0}^{n}a_ix^i, \ a_i \in F 
    \ (a_n = 1)$. 

    Since $\sigma$ fixes $F$, we have $\sigma(a_i) = a_i$. 
    Then:
    \[
        0 = \sigma(m_\alpha(\alpha)) 
        = \sigma\left(\sum a_i\alpha^i\right) 
        = \sum a_i\sigma(\alpha)^i = m_\alpha(\sigma(\alpha))
    \]
\end{proof}

\begin{prop}
    Let $E = F(\alpha)$, and let $\sigma \in \operatorname{Aut}(E/F)$.
    Then $\sigma$ is completely determined by $\sigma(\alpha)$.
\end{prop}

\begin{thm}
    Let $E = F(\alpha)$, where $\alpha$ is algebraic over 
    $F$ with minimal polynomial $m_\alpha(x)$. 

    Then 
    \[
        |\operatorname{Aut}(E/F)| \leq \deg m_\alpha.
    \]
\end{thm}
\begin{defn}
    A field embedding 
    \[
        \sigma : F \to E
    \]
    is an injective field homomorphism.
\end{defn}

\begin{thm}
    Let $F \subseteq E = F(\alpha)$, $\sigma : F \to K$ be 
    a field homomorphism, $m_\alpha(x)$ the minimal 
    polynomial of $\alpha$.
    Then there exists an extension 
    \[
        \widetilde\sigma : E \to K
    \]
    with $\widetilde\sigma(\alpha) = \beta$ if and only if 
    $\beta$ is a root of $\sigma(m_\alpha)(x)$.
\end{thm}

\begin{thm}[Isomorphism Extension Theorem]
    Let $\sigma : F \to K$ be a field isomorphism and 
    $\alpha$ algebraic over $F$ with minimal polynomial 
    $m_\alpha(x)$.

    Let $\beta \in \overline K$ be a root of 
    $\sigma(m_\alpha)(x)$. 

    Then $\sigma$ is extends to an isomorphism 
    \[
        \widetilde \sigma : F(\alpha) \to K(\beta)
    \]
    with 
    \[
        \widetilde \sigma(\alpha) = \beta.
    \]
\end{thm}
\begin{thm}[Automorphisms of Finite Fields]
    Let $E = \mathbb F_{p^n}$. Then 
    \[
        \operatorname{Aut}(E/\mathbb F_p) = \langle 
        \phi \rangle
    \]
    where 
    \[
        \phi(x) = x^p
    \]
    and 
    \[
        |\operatorname{Aut}(E/\mathbb F_p)| = n.
    \]
\end{thm}

\section{Exam Practice}
Henceforth these notes will be mostly revision, 
even though I haven't covered all topics yet. 
There will also probably be a lot of problems repeated.

\begin{prob}[1 Exam 2019]
    \begin{enumerate}[label=\alph*)]
        \item Is $\mathbb Z_{\geq 0}$ equipped with 
            addition a group? Justify your answer.

        \item What are the abelian groups of order $18$ up 
            to isomorphism?

        \item What are the cosets of the subgroup 
            $\langle (0,3)\rangle$ in $\mathbb Z_2 \times 
            \mathbb Z_9$? Describe the quotient 
            \[
                \frac{\mathbb Z_2 \times \mathbb Z_9}{
                    \langle(0,3)\rangle
                }
            \]
            as a product of finite cyclic groups.
    \end{enumerate}
\end{prob}
\begin{proof}[Proof of (a)]
    No it is not a group because there are additive inverses 
    for all elements except $0$. 

    A single counter-example is $1 \in \mathbb Z_{\geq 0}$ 
    which has no inverse.

    \[
        1 + x = 0 \Rightarrow x = -1 \not\in \mathbb Z_{\geq 0}.
    \]
    We can prove it rigorously as follows. 
    $0$ is obviously not the additive inverse of $1$. 
    Suppose $n \in \mathbb Z_{\geq 0}$ is not an inverse 
    of $1$. Assume, for a contradiction, that $n+1$ is 
    an inverse of $1$.
    Then 
    \begin{align*}
        1 + (n + 1) &= 0 \\ 
        n &= -2 
    \end{align*}
    From this we conclude that the inverse of $1$ is the 
    inverse of $2$, i.e 
    \[
        -1 = -2
    \]
    so $1 = 2$. Hence $n+1$ is not an inverse of $1$. 
    By induction no element of $\mathbb Z_{\geq 0}$ is an 
    inverse of $1$.
\end{proof}
\begin{proof}[Solution of (b)]
    We use the fundamental theorem of finitely generated 
    abelian groups. $18$ is finite so we can simply find 
    all ways of expressing $18$ as products of prime powers and 
    we'll be done.
    We see that 
    \[
        18 = 2 \cdot 9 = 2 \cdot 3^2 = 2 \cdot 3 \cdot 3
    \]
    so there are two (up to isomprhism) abelian groups of 
    order $18$, namely 
    \[
        \mathbb Z_2 \times \mathbb Z_9 \ \text{and} \ 
        \mathbb Z_2 \times \mathbb Z_3 \times \mathbb Z_3.
    \]
\end{proof}
\begin{proof}[Solution of (c)]
    $\langle (0,3) \rangle = \{(0,0), (0,3), (0,6)\}$. 
    The left and right cosets coincide since this group 
    is abelian and they are precisely
    \[
        \{x + y : y \in \langle (0,3)\rangle \}, \ 
        x \in \mathbb Z_2 \times \mathbb Z_9.
    \]
    Since for $(a,b)$, $b$ is taken mod $9$ only $b$ mod 3 
    matters. 
    So the left cosets are 
    \[
        (a,b) + \langle (0,3) \rangle 
    \]
    where $a \in \mathbb Z_2$ and $b \in \mathbb Z_3$.
    Denote $H := \langle (0,3) \rangle$.
    That gives us cosets
    \begin{align*}
        (0,0) + H, \ (0,1) + H, \ (0,2) + H \\ 
        (1,0) + H, \ (1,1) + H, \ (1,2) + H 
    \end{align*}
    i.e. $6$ cosets, which is what we expect 
    because the order of $\mathbb Z_2 \times \mathbb Z_9$ is 
    $18$ and $18/|H| = 6$. The quotient groups is abelian 
    and $6 = 2 \cdot 3$ so necessarily 
    \[
        \frac{\mathbb Z_2 \times \mathbb Z_9}{\langle (0,3)\rangle} 
        \cong \mathbb Z_2 \times \mathbb Z_3.
    \]
    We can also argue by observing the surjective homomorphism 
    $\phi : \mathbb Z_2 \times \mathbb Z_9 \to 
    \mathbb Z_2 \times \mathbb Z_3$, $\phi(a,b) = (a, b\mod 3)$ 
    has kernel 
    $\ker \phi = H$ so the result follows by the first isomorphism 
    theorem.
\end{proof}

\begin{prob}[2 Exam 2019]
    \begin{enumerate}[label=\alph*)]
        \item Let $\sigma = (2 \ 6 \ 1 \ 4 \ 3)$ and 
            $\tau = (5 \ 2 \ 1 \ 3)$ be permutations in $S_6$.
            Express the composition $\sigma\tau$ in disjoint 
            cycle notation.
            What is the order of $\sigma\tau$?
        \item Show that if $\sigma \in S_n$ is a cycle of odd length 
            $2k+1$, then $\sigma^2$ is also a cycle of length 
            $2k+1$.
        \item Let $G$ be a subgroup of the symmetric group 
            $S_n$ such that there exist permutations 
            $\sigma_2,\ldots,\sigma_n \in G$ satisfying 
            $\sigma_i(1) = i$ for $2 \leq i \leq n$. Show that 
            $G$ acts \textbf{transitively} on $\{1,\ldots,n\}$. In 
            other words, show that for each pair $1 \leq i,j \leq n$
            there exists a permutation $\sigma \in G$ such that 
            $\sigma(i) = j$.
    \end{enumerate}
\end{prob}
\begin{proof}[Solution of (a)]
    $\sigma$ is $2 \to 6 \to 1 \to 4 \to 3 \to 2$ while 
    $\tau$ is $5 \to 2 \to 1 \to 3$. 
    We work out $\sigma\tau$ like this: 

    \begin{align*}
        5 \to 2 \to 6 \Rightarrow 5 \to 6 \\ 
        2 \to 1 \to 4 \Rightarrow 2 \to 4 \\ 
        1 \to 3 \to 2 \Rightarrow 1 \to 2 \\ 
        3 \to 5 \to 5 \Rightarrow 3 \to 5 \\ 
        4 \to 4 \to 3 \Rightarrow 4 \to 3 \\ 
        6 \to 6 \to 1 \Rightarrow 6 \to 1
    \end{align*}
    so $\sigma\tau = (1 \ 2 \ 4 \ 3 \ 5 \ 6)$. 
    The order of an $n$-cycle is $n$ so $|\sigma\tau| = 6$.
\end{proof}
\begin{proof}[Proof of (b)]
    Suppose $\sigma$ is a cycle of length $2k+1$, i.e. 
    \[
        \sigma = (a_1 \ a_2 \ \cdots \ a_{2k+1}).
    \]
    Composing $\sigma^2$ sends $a_1 \to a_3$, $a_2 \to a_4$ and so 
    on. In particular $a_{i} \to a_{(i + 2)}$ where $i+2$ is 
    taken mod $2k+1$. Then all odd elements are taken 
    to odd elements, but the $2k+1$-th element is sent 
    to $a_2$ which is sent to $a_4$ and so on. Even elements 
    are sent to even elements until $a_{2k} \to a_1$ so 
    \[
        \sigma^2 = (a_1 \ a_3 \ \cdots \ a_{2k+1} \ a_2 \ 
        a_4 \ \cdots \ a_{2k})
    \]
    which has length $2k+1$.
\end{proof}
\begin{proof}[Proof of (c)]
    Since $G$ is a group, each permutation $\sigma_k$ has an 
    inverse $\sigma_k^{-1}$ which sends $k$ to $1$.
    Also $e \in G$ so for $1 \leq i,j \leq n$ we can send 
    $i$ to $j$ by sending $i$ to $1$ and then $1$ to $j$ so 
    \[
        (\sigma_j \circ \sigma_i^{-1})(i) = 
        \sigma_j(1) = j.
    \]
    This works well for the case where neither $i,j$ are 
    $1$. If $i$ is one then let $\sigma_i := e$ and likewise 
    for $j$.
\end{proof}
\begin{prob}[3 Exam 2019]
    \begin{enumerate}[label=\alph*)]
        \item Show that if $n$ is not prime, then $n \mathbb Z$ 
            is not a prime ideal of $\mathbb Z$.
        \item Show that the quotient 
            \[
                \frac{\mathbb Z_2[x]}{\langle x^3 + x + 1\rangle}
            \]
            is a field.
            How many elements does it contain?
        \item Find a polynomial $p(x) \in \mathbb Z_2[x]$ such that 
            $\deg(p) \leq 2$ and 
            \[
                x^5 + \langle x^3 + x + 1 \rangle
                = 
                p(x) + \langle x^3 + x + 1\rangle
            \]
            in $\frac{\mathbb Z_2[x]}{\langle x^3 + x + 1 \rangle}$.
    \end{enumerate}
\end{prob}
\begin{proof}
    Suppose $n$ is not prime. Then $n=ab$ for some integers
    $a,b$ with $1<a<n$ and $1<b<n$. Hence
    \[
        ab=n\in n\mathbb Z.
    \]
    However, $a\notin n\mathbb Z$ and $b\notin n\mathbb Z$, since
    neither
    $a$ nor $b$ is divisible by $n$. Thus there exist 
    $a,b\in\mathbb Z$
    such that $ab\in n\mathbb Z$ but $a,b\notin n\mathbb Z$.
    Therefore $n\mathbb Z$ is not a prime ideal of $\mathbb Z$.
\end{proof}
\begin{proof}[Proof of (b)]
    We will use the fact that the quotient (in this case) is a field 
    if and only if $f(x) = x^3 + x + 1$ is irreducible over 
    $\mathbb Z_2$. Notice that 
    \[
        f(0) = f(1) = 1 \neq 0
    \]
    and $f(x)$ is degree $3$ hence irreducible over 
    $\mathbb Z_2$. Thus the quotient is a field.
    There are $2^3 = 8$ elements.
\end{proof}
\begin{proof}[Solution of (c)]
    Since $x^3 + x + 1 = 0$ we have $x^3 = x + 1$. 
    Then $x^5 = x^2x^3 = x^2(x+1) = x^3 + x^2$. 
    Using the relation again we get 
    $x^5 = x^2 + x + 1$. So let $p(x) := x^2 + x + 1$. Then 
    $x^5 \langle x^3 + x + 1\rangle = p(x) + \langle x^3 + x 
    + 1\rangle$.
\end{proof}

\begin{prob}[1 Exam 2022, (Group Theory)]
    \medskip 
    \noindent
    \begin{enumerate}
        \item Let $G$ be a group of order $200$.  
            What are the sizes of Sylow $p$-subgroups of $G$? 
        \item Let $G = \mathbb Z_{12} \times \mathbb Z_8$. Find 
            the order of the element $(4,2) \in G$. 
            Determine the quotient group $G / \langle (4,2)\rangle$
            up to isomorphism.
        \item Let $p$ be a prime number. Suppose $G$ is a group 
            of order $p^n$ acting on a set $X$ such that 
            $p$ does not divide $|X|$. Show that the 
            action of $G$ must have a fixed point, i.e. 
            there exists an $x \in X$ such that $|\mathcal O| = 1$. 
    \end{enumerate}

    Note: Problem $1$ is not relevant to the current curriculum.
\end{prob}
\begin{proof}[Solution of (2)]
    $4$ has order $3$ in $\mathbb Z_{12}$ and $2$ has order 
    $4$ in $\mathbb Z_8$ so the order of $(4,2)$ in $\mathbb Z_{12} 
    \times \mathbb Z_8$ is $\operatorname{lcm}(3,4) = 12$.

    Thus there are $12$ elements in $\langle (4,2) \rangle$ which 
    we will use to characterize the quotient group.

    $G$ has $12 \cdot 8 = 96$ elements which by Lagrange's theorem 
    gives $96/12 = 8$ elements in the quotient group.
    Now since the quotient group is a finite abelian group 
    it must be isomorphic to $\mathbb Z_8$ or $\mathbb Z_2 \times 
    \mathbb Z_2$ or $\mathbb Z_2 \times \mathbb Z_2 \times 
    \mathbb Z_2$.

    We see quickly that $(0,1)$ has order $4$ in the quotient 
    so it can't be $\mathbb Z_2^3$. Furthermore $4(a,b) \equiv (a,b)$
    when taken mod $8$ so the maximum order of any element is 
    $4$ so we conclude that the quotient is (up to isomorphism) 
    \[
        \mathbb Z_4 \times \mathbb Z_2.
    \]
\end{proof}
\begin{proof}[Proof of (3)]
    We particion $X$ into disjoint orbits 
    $\mathcal O_1,\ldots, \mathcal O_j$. Assume, for contradiction,
    that $|\mathcal O_x| > 1$ for all $x \in X$.
    Since $|\mathcal O_x| = [G:G_x] = |G|/|G_x|$ we have that 
    $|\mathcal O_x| = p^{n_x}$ for some $n_x > 0$. Then 
    since 
    \[
        |X| = \sum_{i=1}^s |O_i|,
    \]
    we have that 
    \[
        |X| = \sum_{i=1}^s p^{n_i}.
    \]
    Then since $p$ divides every term we have 
    \[
        p \bigm| |X|
    \]
    which contradicts our assumption that $p$ does not divide 
    $|X|$.
\end{proof}

\begin{prob}[2 Exam 2022, Ring Theory]
    \medskip 
    \noindent 
    \begin{enumerate}
        \item Show that $\mathbb Z_{12}$ is not an integral domain. 
        \item Let $p$ be prime and $n > 1$. Find a maximal ideal 
            $M$ of $\mathbb Z_{p^n}$ and identify the 
            quotient $\mathbb Z_{p^n} / M$. 
        \item Show that if an ideal $I$ of the polynomial ring 
            $\mathbb Q[x]$ contains both $f(x) = x^3 - x^2 + x$ 
            and $g(x) = x^2 - 1$ then $I = \mathbb Q[x]$.
    \end{enumerate}
\end{prob}
\begin{proof}[Solution sketch of (1)]
    $\mathbb Z_{12}$ contains zero-divisors 
    $3 \cdot 4 = 12 \equiv_{12} 0$ and is therefore not an 
    integral domain.
\end{proof}
\begin{proof}[Proof attempt of (2)]
    We need to find an ideal $M$ so that 
    if $M \subseteq K \subseteq \mathbb Z_{p^n}$ then 
    $K = M$ or $K = \mathbb Z_{p^n}$.

    Let $M := \langle p \rangle$. Then $M$ is a maximal ideal.  

    $M$ absorbs any power of $p$ so $|\mathbb Z_{p^n}|/M| < p^n - 
    n$. Etc.   
\end{proof}
\begin{proof}[Proof of (3)]
    We notice that 
    \[
        \frac{x^3 - x^2 + x}{x^2 - 1} 
        = x - 1 - \frac{1}{x^2 - 1} =: r(x).
    \]
    Something, something use addition and multiplication with 
    these polynomials to get scalars, $x$ and such to be able 
    to make all polynomials over $\mathbb Q$.
\end{proof}
\begin{prob}[3, Exam 2022 (Finite Fields)]
    Consider $F = \mathbb Z_3$ and the irreducible polynomial 
    $f(x) = x^3 + 2x + 2 \in F[x]$.
    Let $E$ be $F[x]/\langle f(x) \rangle$.
    \begin{enumerate}
        \item What is the size of $E$?
        \item Let $\alpha \in \overline F$ denote a zero of 
            $f(x)$ so that $E \cong F(\alpha)$. Let 
            \[
                \sigma_3 : F(\alpha) \to F(\alpha)
            \]
            denote the Frobenius automorphism defined by 
            $\sigma_3(a) = a^3$ for all $a \in F(\alpha)$.
            Write $\sigma_3(\alpha)$ and $\sigma_3(\alpha)^2$ in the 
            basis $\{1,\alpha,\alpha^2\}$ for $F(\alpha)$ over 
            $F$.
        \item Assume $g(x) = x^3 + 2x + 1 \in F[x]$ is irreducible 
            over $F$ and let $\beta \in \overline F$ be a zero of 
            $g(x)$.

            Prove that $F(\alpha)$ and $F(\beta)$ are isomorphic 
            without giving an isomorphism. However explain why 
            the map $\psi_{\alpha,\beta} : F(\alpha) \to F(\beta)$
            which fixes $F$ and sends $\alpha$ to $\beta$ cannot 
            be made into a field isomorphism.
    \end{enumerate}

\end{prob}
\begin{proof}[Solution of (1)]
    $\deg f = 3$ so $|E| = 3^3 = 27$.
\end{proof}
\begin{proof}[Solution of (2)]  
    $f(\alpha) = \alpha^3 - 2 \alpha + 2 = 0$ so 
    $\alpha^3 = 2\alpha - 2 = \alpha + 1$.

    Then 
    \[
        \sigma_3(\alpha) = \alpha^3 = \alpha + 1.
    \]
    Next 
    \[
        \sigma_3^2(\alpha) = (\alpha + 1)^3 = \alpha^3 + 1 
        = \alpha + 2.
    \]
\end{proof}
\begin{proof}[Proof of (3)]
    Since both are fields of size $27 = 3^3$ they are isomorphic.  

    The map $\alpha \to \beta$ cannot be a field isomorphism because 
    their respective minimal polynomials do not coincide.
\end{proof}

\begin{prob}
    Let 
    \[
        f(x) = x^4 - 2 \in \mathbb Q[x].
    \]
    Find 
    \begin{enumerate}
        \item the roots of $f$, 
        \item the splitting field $K$, 
        \item the degree $[K:\mathbb Q]$, 
        \item what the automorphism in $\operatorname{Gal}(K/\mathbb 
            Q)$ can do to $\sqrt[4]2$ and $i$.
    \end{enumerate}
\end{prob}
\begin{proof}[Solution]
    \begin{align*}
        f(x) &= (x^2 - \sqrt 2)(x^2 + \sqrt 2) \\ 
             &= (x - \sqrt[4]2)(x + \sqrt[4]2)(x^2 + \sqrt 2)
    \end{align*}
    so we apply the quadratic formula to the latter term: 
    \begin{align*}
        x &= \frac{\pm \sqrt{-4\sqrt{2}}}{2} \\ 
          &= \pm \sqrt{-\sqrt{2}} \\ 
          &= \pm i \sqrt[4]{2}
    \end{align*}
    yielding 
    \[
        f(x) = (x + \sqrt[4]2)(x - \sqrt[4]2)(x + i\sqrt[4]2)
        (x - i\sqrt[4]2).
    \]
    which has roots 
    \[
        \pm \sqrt[4]2 \ \text{and} \ 
        \pm i \sqrt[4]2. 
    \]
    Thus the splitting field is 
    \[
        K = \mathbb Q(\sqrt[4]2, i).
    \]

    As we can see $f$ is irreducible over $\mathbb Q$ by 
    Eisenstein with $p = 2$ and in fact it is the minimal 
    polynomial of $\sqrt[4]2$ so 
    \[
        [\mathbb Q(\sqrt[4]2) : \mathbb Q] = \deg f = 4.  
    \]
    We can also see that $i \not \in \mathbb R \supset 
    \mathbb Q(\sqrt[4] 2)$ so $K$ has nontrivial degree over 
    $\mathbb Q(\sqrt[4] 2)$ and since the minimal polynomial is 
    $x^2 + 1$ we conclude that 
    \[
        [K : \mathbb Q(\sqrt[4]2)] = 2
    \]
    so 
    \[
        [K : \mathbb Q] = 8.
    \]

    Let $\alpha = \sqrt[4]2$. Then $K = \mathbb Q(\alpha, i)$ and 
    every automorphism in the Galois group is determined by 
    where it sends $\alpha$ and $i$.
    For the roots to satisfy their respective polynomials we have 
    \[
        \sigma(\alpha) \in \{\alpha, -\alpha, i\alpha, -i\alpha\}
    \]
    and 
    \[
        \sigma(i) \in \{i,-i\}.
    \]
\end{proof}

\begin{prob}
    Let 
    \[
        R = \mathbb F_3[x]/\langle x^2 + x + 2 \rangle. 
    \]

    Let 
    \[
        \alpha = x + \langle x^2 + x + 2 \rangle.
    \]

    \begin{enumerate}[label=\alph*)]
        \item Show that 
            $x^2 + x + 2$ is irreducible over 
            $\mathbb F_3$. Conclude that 
            $R$ is a field.

        \item How many elements does $R$ have? 
            Write them in the form 
            \[
                a + b\alpha, \ a,b \in\mathbb F_3. 
            \]
        \item Use the quotient relation to compute 
            \[
                \alpha^2, \alpha^3, \alpha^4. 
            \]

        \item Find the inverse of 
            \[
                1 + \alpha
            \]
            in $R$.
    \end{enumerate}
\end{prob}
\begin{proof}[Proof of (a)]
    Let $f(x) := x^2 + x + 2$. Since $\deg f \in \{2,3\}$ we 
    have that $f$ is irreducible if and only if 
    $f$ has no zeros in the ground field. 

    We quickly check that 
    \[
        f(0) = 2, \ f(1) \equiv 1, \ f(2) \equiv 2,  
    \]
    meaning $f$ has no zeros in $\mathbb F_3$ and is thus 
    irreducible. 

    Since $\mathbb F_3$ is a field we have that 
    the quotient is a field if and only if $f$ is irreducible 
    over $\mathbb F_3$. Thus we conclude that 
    $R$ is a field.
\end{proof}

\begin{proof}[Solution to (b)]
    We use the standard result to get 
    \[
        |R| = |\mathbb F_3|^{\deg f} = 9.
    \]

    Conclusively 
    \[
        R = \{0,1,2, \alpha, 1 + \alpha, 2 + \alpha, 
        2\alpha, 2\alpha + 1, 2\alpha + 2\}.
    \]
\end{proof}

\begin{proof}[Solution to (c)]
    We use
    \[
        \alpha^2 + \alpha + 2 = 0
    \]
    so 
    \[
        \alpha^2 = 2\alpha + 1
    \]
    and then 
    \[
        \alpha^3 = \alpha(2\alpha + 1) 
        = 2\alpha^2 + \alpha 
        = 2(2\alpha + 1) + \alpha 
        = 2\alpha + 2   
    \]
    and lastly 
    \[
        \alpha^4 = \alpha(2\alpha + 2) 
        = 2\alpha^2 + 2\alpha 
        = 2(2\alpha + 1) + 2\alpha
        = 4\alpha + 2 + 2\alpha 
        = 2. 
    \]
\end{proof}
\begin{proof}[Solution to (c)]
    We must find $a + b\alpha$ such that 
    \[
        (1 + \alpha)(a + b\alpha) = 1.
    \]
    Expanding we get 
    \begin{align*}
        a + b\alpha + a\alpha + b\alpha^2 &= 1 \\ 
        a + a\alpha + b\alpha + 2b\alpha + b &= 1 \\ 
        a + 3b &= a \\ 
               &\; \vdots
    \end{align*}
    We get $(1 + \alpha)^{-1} = \alpha$. 
\end{proof}

\begin{prob}
    Let 
    \[
        I = \langle x^2 + 1 \rangle \subseteq \mathbb Q[x].
    \]
    \begin{enumerate}[label=\alph*)]
        \item Show that $I$ is maximal. 
        \item Show that 
            \[
                \mathbb Q[x]/I \cong \mathbb Q(i).
            \]
        \item Find the inverse of 
            \[
                x + 1 + I
            \]
            in the quotient ring.
    \end{enumerate}
\end{prob}
\begin{proof}[Proof of (a)]
    $\mathbb Q[x]$ is a PID so every ideal is generated by 
    a polynomial.
    Since $\mathbb Q$ is a field, $\langle f(x) \rangle$ is maximal 
    if and only if $f$ is irreducible over $\mathbb Q$, 
    which $x^2 + 1$ obviously is. 
\end{proof} 

\begin{proof}[Proof of (b)]
    We will use the isomorphism theorem for rings. 

    Let $\psi : \mathbb Q[x] \to \mathbb Q(i)$ be 
    \[
        \psi(p) := p(i).
    \]
    Then very clearly $\langle x^2 + 1\rangle \subseteq 
    \ker \psi$. Now take any $p(x) \in \ker \psi$. 
    Then 
    \[
        p(x) = q(x)(x^2 + 1) + r(x),
    \]
    where 
    \[
        \deg r < 2.
    \]
    So $r(x) = ax + b$ for some $a,b \in \mathbb Q$. 
    Evaluate at $i$ to get 
    \[
        0 = p(i) = q(i)(i^2 + 1) + r(i) = ai + b,
    \]
    forcing $a = b = 0$. Thus 
    \[
        p(x) = q(x)(x^2 + 1) \in \langle x^2 + 1 \rangle.
    \]
    Hence $\ker \psi = \langle x^2 + 1 \rangle$.

    Lastly the map is surjective, i.e. $\operatorname{im} \psi = 
    \mathbb Q(i)$, 
    because take any $\gamma \in \mathbb Q(i)$. Then 
    \[
        \gamma = a + bi, \ a,b \in \mathbb Q.
    \]
    Notice then that we can find the polynomial 
    \[
        f(x) := a + bx \in \mathbb Q[x], 
    \]
    for which we have 
    \[
        \psi(f) = a + bi. 
    \]
    Thus by the first isomorphism theorem for rings, 
    \[
        \frac{\mathbb Q[x]}{\langle x^2 + 1\rangle} 
        \cong \mathbb Q(i).
    \]
\end{proof}

\begin{prob}[Problem 3 Exam 2021]
    \ 
    \medskip

    \begin{enumerate}[label=\alph*)]
        \item Determine if $\mathbb Z_2[x]/\langle x^4 + 1 \rangle$ 
            is an integral domain.
        \item Show that $f(x) = x^4 + x^3 + 1$ is 
            irreducible in $\mathbb Z_2[x]$. Explain why 
            $f(x)$ admits a zero $\theta$ in the field 
            $\mathbb Z_{16}$, seen as an extension of 
            $\mathbb Z_2$, with degree 
            $[\mathbb Z_{16} : \mathbb Z_2] = 4$. Prove that 
            $f(x)$ is the primitive polynomial in $\mathbb Z_2[x]$.
    \end{enumerate}
\end{prob}
\begin{proof}[Solution of (a)]
    We notice that, in $\mathbb Z_2[x]$, 
    \[
        x^4 + 1 = x^4 - 1 = (x^2 + 1)(x^2 - 1)
    \]
    and so is the product of two lower-degree (yet non-zero) 
    polynomials in $\mathbb Z_2[x]$, and is thus not 
    irreducible. From this we can very quickly 
    see that $\langle x^4 + 1 \rangle$ is not a 
    prime ideal, meaning $\mathbb Z_2[x]/\langle x^4 + 1 \rangle$
    is not an integral domain.
\end{proof}
\begin{proof}[Proof of (b)]
    We see that $f$ has no zeros in $\mathbb Z_2$ and that 
    any factorization into two 
    degree-$2$ polynomials 
    \[
        f(x) = (x^2 + ax + b)(x^2 + cx + d)
    \]
    requires $b=d=1$, but then $1 = a+c = 0$.
    Therefore $f$ is irreducible over the ground field.
\end{proof}

\begin{prob}
    Let 
    \[
        f(x) = x^3 + x + 1 \in \mathbb F_2[x],
    \]
    and let 
    \[
        K = \mathbb F_2[x]/\langle f \rangle.
    \]
    Let 
    \[
        \theta = x + \langle f \rangle \in K.
    \]
    \begin{enumerate}[label=\alph*)]
        \item Show that $f$ is irreducible over 
            $\mathbb F_2$. Conclude $K$ is a field.
        \item Use 
            \[
                f(\theta) = 0
            \]
            to express $\theta^3$
            in terms of $1,\theta, \theta^2$.
        \item Determine whether $\theta$ is a generator 
            of $K^\times$. 
        \item Write the factorization of $f$ into linear 
            factors over $K$ using Frobenius conjugates.
    \end{enumerate}
\end{prob}
\begin{proof}[Proof of (a)]
    Since $f$ is degree-$3$ we have irreducible if and only 
    if no zeros. We notice that 
    \[
        f(0) = f(1) = 1.
    \]
    Thus $f$ is irreducible over $\mathbb F_2$.

    Since $\mathbb F_2$ is a field we have that the quotient 
    \[
        \mathbb F_2/\langle f \rangle
    \]
    is a field if and only if $f$ is irreducible over 
    $\mathbb F_2$. Thus $K$ is a field with 
    \[
        |K| = 2^3 = 8.
    \]  
\end{proof}
\begin{proof}[Solution of (b)]
    We have 
    \[
        \theta^3 + \theta + 1 = 0
    \]
    meaning 
    \[
        \theta^3 = -\theta - 1 = \theta + 1.
    \]
\end{proof}
\begin{proof}[Solution of (c)]
    $\langle f \rangle$ is maximal so 
    \[
        \frac{\mathbb F_2}{\langle f \rangle}
    \]
    can be viewed as a field-extension of $\mathbb F_2$ of 
    degree $\deg f = 3$, with the 
    basis $\{1, \theta, \theta^2\}$. We know 
    \[
        |K| = 8
    \]
    so 
    \[
        K \cong \mathbb F_8.
    \]
    Now the question is whether 
    \[
        \langle \theta \rangle 
        \cong K^\times \cong \mathbb F_7.
    \]
    We compute the lowest $j$ such that
    \[
        \theta^j = 1.
    \]
    We see that 

    \begin{align*}
        \theta^7 
        &= \theta^2\theta^2\theta^3 \\ 
        &= \theta^4(\theta + 1) \\ 
        &= \theta^5 + \theta^4 \\ 
        &= \theta^2(\theta + 1) + \theta^4 \\ 
        &= \theta^3 + \theta + 2\theta^2 \\ 
        &= \theta^3 + \theta \\ 
        &= -1 \\ 
        &= 1. 
    \end{align*}
    Doing further computations for $\theta^j$ for $j < 7$ we see 
    that this is indeed the order of $\theta$. Thus 
    \[
        |\langle \theta \rangle| = 7 = |\mathbb F_7|
    \]
    and we conclude that 
    $\theta$ does indeed generate the multiplicative group.

\end{proof}

\begin{prob}
    Let 
    \[
        f(x) = x^4 + x + 1 \in \mathbb F_2[x],
    \]
    and define 
    \[
        K = \mathbb F_2[x]/\langle f \rangle.
    \]
    Let 
    \[
        \alpha = x + \langle f(x) \rangle \in K.
    \]

    \begin{enumerate}[label=\alph*)]
        \item Show that $f(x)$ is irreducible over 
            $\mathbb F_2$.
        \item Conclude that $K$ is a field. How many elements does 
            $K$ have? 
            Give a basis for $K$ as a vector space over 
            $\mathbb F_2$.
        \item Use $f(\alpha) = 0$ to compute 
            \[
                \alpha^4, \ \alpha^5, 
                \ \alpha^6, \ \alpha^7.
            \]
            Write each in the basis from $b)$.
        \item Determine whether $\alpha$ generates 
            $K^\times$.
    \end{enumerate}
\end{prob}
\begin{proof}[Proof of (a)]
    Suppose $f$ is reducible. Then it factors into two 
    degree-$2$ polynomials 
    \[
        f(x) = x^4 + x + 1 = (x^2 + ax + b)(x^2 + cx + d).
    \]
    Then we require $bd = 1$ so $b = d = 1$, and 
    \[
        adx + bcx = (ad + bc)x = x, 
    \]
    so 
    \[
        ad + bc = 1
    \]
    and since $b = d = 1$ we then have 
    \[
        a + c = 1.
    \]
    Then either $a=1$ or $c = 1$ while the other 
    is $0$. Suppose wlog $a = 1$ and $c = 0$. Then 
    \[
        x^4 + x + 1 = (x^2 + x + 1)(x^2 + 1) 
        = x^4 + x^2 + x^3 + x + x^2 + 1 
        = x^4 + x^3 + x + 1
    \]
    which is false. Our assumption was incorrect and 
    $f$ is irreducible. 
\end{proof}
\begin{proof}[Solution of (b)]
    As $f$ is irreducible we conclude 
    $\langle f \rangle$ is a maximal ideal of 
    $\mathbb F_2[x]$. Since $\mathbb F_2$ is a field 
    (and therefore a commutative ring), we have that 
    \[
        \frac{\mathbb F_2[x]}{\langle f \rangle} 
    \]
    is a field. 
    
    Since $\mathbb F_2$ is a finite field, and any 
    non-trivial quotient of the polynomial ring over a finite 
    field by an irreducible polynomial is a finite field 
    extension, we conclude 
    \[
        \frac{\mathbb F_2[x]}{\langle f \rangle} 
    \]
    must be a (finite) field extension.

    The standard result is that 
    \[
        \left| \frac{\mathbb F_2[x]}{\langle f \rangle} \right| 
        = |\mathbb F_2|^{\deg f} 
        = 16.
    \]
    The standard basis is then 
    \[
        1, \alpha, \alpha^2, \alpha^3
    \]
    since 
    \[
        \left[\frac{\mathbb F_2[x]}{\langle f \rangle} : \mathbb F_2
            \right] 
        = 4.
    \]
\end{proof}
\begin{proof}[Solution of (c)]
    We have that 
    \[
        \alpha^4 + \alpha + 1 = 0, 
    \]
    so 
    \[
        \alpha^4 = \alpha + 1.
    \]
    Then 
    \[
        \alpha^5 = \alpha(\alpha + 1) = \alpha^2 + \alpha,
    \]
    and 
    \[
        \alpha^6 = \alpha(\alpha^2 + \alpha) = \alpha^3 + \alpha^2. 
    \]
    Lastly 
    \[
        \alpha^7 = \alpha^4 + \alpha^3 
        = \alpha^3 + \alpha + 1.
    \]
\end{proof}

\begin{proof}[Proof of (d)]
    We have that 
    \[
        K := \frac{\mathbb F_2[x]}{\langle f \rangle} 
        \cong \mathbb F_{16},  
    \]
    since the finite fields are classified up to isomorphism. 
    Then the question is whether 
    \[
        \langle \alpha \rangle 
        = K^\times \cong \mathbb F_{16}^\times \cong 
        C_{15} \leq \mathbb F_{16}.
    \]
    Then if we find that $|\langle \alpha \rangle| = 15$ we 
    will be finished. 
    First we can check whether 
    $\alpha^{15} = 1$ as we would expect. 

    \begin{align*}
        \alpha^{15} 
        &= \alpha^8(\alpha^3 + \alpha + 1) \\ 
        &= \alpha^{11} + \alpha^9 + \alpha^8 \\ 
        &= (\alpha^5 \alpha^6) + \alpha^3\alpha^3 + (\alpha + 1)^2 \\ 
        &= (\alpha^2 + \alpha)(\alpha^3 + \alpha^2) + 
        \alpha^3 \alpha^3 + \alpha^2 + 1 \\ 
        &= \alpha^5 + \alpha^4 + \cdots 
    \end{align*}
\end{proof}

\begin{prob}
    Let $G = C_4 = \langle r \rangle$ act on the set 
    \[
        X = \{0,1\}^4
    \]
    of binary strings of length $4$ by cyclic rotation.
    Thus 
    \[
        r(a_1,a_2,a_3,a_4) = (a_4,a_1,a_2,a_3).
    \]
    In other words, two binary strings are considered 
    equivalent if one can be obtained from the other by 
    rotating the positions.

    \begin{enumerate}[label=\alph*)]
        \item Write down the elements of $G$ and describe how 
            each acts on $X$. 
        \item For each $g \in G$, compute the number of fixed 
            points 
            \[
                |X^g| = \{x \in X : gx = x\}.
            \]
        \item Use Burnside's formula to determine the number 
            of orbits. 
        \item Interpret the answer: how many 
            binary of necklaces of length $4$ are there, up to 
            rotation?
    \end{enumerate}
\end{prob}
\begin{proof}[Solution and proof to all]
    $G = \{e, r, r^2, r^3\}$.
    Suppose $(a_1,a_2,a_3,a_4) \in X$. Then 
    \[
        r(a_1,a_2,a_3,a_4) = (a_4,a_1,a_2,a_3)
    \]
    and 
    \[
        r^2(a_1,a_2,a_3,a_4) = r(a_4,a_1,a_2,a_3) 
        = (a_3,a_4,a_1,a_2)
    \]
    and 
    \[
        r^3(a_1,a_2,a_3,a_4) = (a_2,a_3,a_4,a_1)
    \]
    with $e$ acting on $X$ by doing nothing.

    $X^e = X$ as always. 

    $X^r = \{(a,a,a,a) : a \in \{0,1\}\} = \{(0,0,0,0),(1,1,1,1)\}$. 
    This is quickly seen since we require $a_1 = a_4$ and 
    $a_2 = a_1$, but then $a_2 = a_4$ and since $a_3 = a_2$ they 
    must all be equal.
    By a similar argument we get the same result for 
    $X^{r^3}$. 

    However, $X^{r^2}$ is different since it just swaps 
    entries $1$ and $2$ with $3$ and $4$, respectively. 

    Thus 
    \[
        X^{r^2} = \{(a,b,a,b) : a,b \in \{0,1\}\}
    \]
    which means $|X^{r^2}| = 2^2 = 4$.

    The number of orbits, by Burnside's lemma/formula is given by 
    \[
        |X/G| = \frac{1}{|G|}\displaystyle\sum_{g \in G}
        |X^g|.
    \]
    Meaning we get 
    \[
        |X/G| = \frac{1}{4} 
        (\underbrace{|X^e|}_{|X|} + \underbrace{|X^r|}_{2} + 
        \underbrace{|X^{r^2}|}_{4} + \underbrace{|X^{r^3}|}_{2})
        = \frac{1}{4}(16 + 2 + 4 + 2) = 24/4 = 6.
    \]

    Thus there are $6$ necklaces unique, up to rotation, when 
    you have two types of pearls, which is quick to verify mentally. 
\end{proof}

\begin{prob}
    Let $G = D_4$ be the symmetries of the square, 
    acting on the set 
    \[
        X = \{\text{colorings of the four vertices of a 
        square using red/blue}\}.
    \]
    We will say that two colorings are the same if 
    one can be obtained from the other by rotating 
    or reflecting the square. 

    \begin{enumerate}[label=\alph*)]
        \item List the elements of $D_4$, and describe their 
            action on the vertices. 
        \item Compute the number of fixed colorings 
            $|X^g|$ for each type of symmetry $g \in D_4$.
        \item Use Burnside's formula to compute 
            the number of orbits.
        \item Pick the coloring 
            \[
                x = (R,R,B,B)
            \]
            around the square. Find the orbit and stabilizer.
            Verify orbit-stabilizer: 
            \[
                |Gx| = \frac{|G|}{|G_x|}. 
            \]
        \item Explain in words how the orbit-stabilizer theorem 
            and Burnside's formula are related, but different.
    \end{enumerate}
\end{prob}
\begin{proof}[Proof and solution of all]
    $D_4$ is the set of rigid symmetries of the square. 
    We can rotate the square $r$ and reflect it $s$. 
    Rotating $4$ times is the same as doing something so 
    $|r| = 4$. Then also $s^2 = e$ and $srs = r^{-1}$.
    Then 
    \[
        D_4 = \{r,s \bigm| r^4 = e, \ s^2 = e, \ srs = r^{-1}\}
        = 
        \{e, r,r^2,r^3,s,rs, r^2s, r^3s\}.
    \]
    Rotating a square $(1,2,3,4)$ gives 
    \[
        r(1,2,3,4) = (4,1,2,3)
    \]
    and reflecting gives 
    \[
        s(1,2,3,4) = (2,1,4,3). 
    \]
    So we can think of these actions as the bijections 
    acting on the indices of the vertices
    $r : \mathbb Z_4 \to \mathbb Z_4$, 
    $r(i) := i + 1 \mod 4$, 
    \[
        s(i) = 
        \begin{cases}
            i + 1 \mod 4, & \text{if odd}\ i \\ 
            i - 1 \mod 4, & \text{if even}\ i
        \end{cases}.
    \]
    Then $r^n(i) = i + n \mod 4$ and 
    \[
        s^n(i)
        = 
        \begin{cases}
            s(i), & \text{if} \ n \ \text{odd} \\ 
            i , & \text{if} \ n \ \text{even}
        \end{cases}.
    \]

    The number of colorings for $e$ is $|X^e| = |X|$. 

    For $g = r$ we, by the same argument as in the necklace problem,
    \[
        |X^r| = 2. 
    \]

    For $g = r^3 = r^{-1}$ we get the same result. 

    For $r^2$ we get 
    \[
        X^{r^2} = \{(a,b,a,b) : a,b \in \{R,B\}\},
    \]
    so 
    \[
        |X^{r^2}| = 4.
    \]

    For $g = s$ we once again get 
    \[
        |X^s| = 4.
    \]

    For $g = rs$ we get 
    \[
        r \circ s(1,2,3,4) = r(2,1,4,3) = 
        (3,2,1,4)
    \]
    which is fixed only if $1 = 3$. Meaning we have $2$ choices 
    for indices $1$ and $3$ and $2$ for index $2$ and likewise for 
    index $4$. Thus $|X^{rs}| = 2^3 = 8$.

    For $g = r^2s$ we get 
    \[
        r^2 \circ s (1,2,3,4) = r(3,2,1,4) = (4,3,2,1). 
    \]
    Here we have $1 \mapsto 4$ and vice versa, likewise for 
    $2$ and $3$. Thus we get $|X^{r^2s}| = 4$. 

    Lastly for 
    $g = r^3s$ we have 
    \[
        r^3 s(1,2,3,4) = r(4,3,2,1) = (1,4,3,2)
    \]
    for which $1$ and $3$ are fixed and $2$ and $4$ are swapped. 
    Like with $g = rs$ we get 
    $|X^{r^3s}| = 8$.

    Then by Burnside's lemma, the number of orbits is 
    \[
        \#(G/X) = 
        \frac{1}{8} (16 + 2 + 2 + 4 + 4 + 8 + 4 + 8) 
        = \frac{48}{8} = 6.
    \]

    Let $x = (R,R,B,B)$. The orbit is 
    \[
        \{x, (B,R,R,B), (B,B,R,R), (R,B,B,R)\}
    \]
    with $|\operatorname{Orb}_G(x)| = 4$.
    Then the stabilizer is 
    \[
        \{e,s\}.
    \]
    Then we have 
    \[
        |G| = 8 = 2 \cdot 4 = |\operatorname{Orb}_G(x)| \cdot 
        |\operatorname{Stab}_G(x)|.
    \]

    Burnside's formula relates the number of orbits to 
    the number of elements fixed by elements in $G$ while 
    the orbit-stabilizer theorem relates 
    elements of $G$ which fix $x \in X$, what 
    elements can be reached by $x$ through the action of the group 
    and states that the product of those will be the size of the 
    group (when finite). The easiest example is a group acting on 
    itself and seeing that $e$ can reach every element in the 
    group ($|G|$ elements), but is only fixed by itself, 
    $|\{e\}| = 1$. 
\end{proof}

\begin{prob}
    Let 
    \[
        f(x) = x^3 + x + 1 \in \mathbb F_2[x],
    \]
    and let 
    \[
        K = \mathbb F[x]/\langle f(x) \rangle.
    \]
    Let 
    \[
        \alpha = x + \langle f(x) \rangle.
    \]

    \begin{enumerate}[label=\alph*)]
        \item Show that $f$ is irreducible over 
            $\mathbb F_2$. Conclude that $K$ is a field.
        \item Use 
            \[
                f(\alpha) = 0
            \]
            to compute 
            \[
                \alpha^3, \alpha^4, 
                \alpha^5, \alpha^6, \alpha^7.
            \]
            in the basis $\{1,\alpha,\alpha^2\}$.
        \item The Frobenius automorphism is
            \[
                \sigma_2 : K \to K, \ \sigma_2(a) = a^2.
            \]
            Compute 
            \[
                \sigma_2(\alpha), \ \sigma_2^2(\alpha), \ 
                \sigma_2^3(\alpha).
            \]
        \item Use Frobenius conjugates to factor 
            $f(x)$ into linear factors over $K$.
    \end{enumerate}
\end{prob}
\begin{proof}[Proof of (a)]
    Since $\deg f \in \{2,3\}$ we can use the fact that $f$ is 
    irreducible over the ground field if and only if 
    it has no zeros in said field. 
    Since 
    \[
        f(0) = 1 = f(1)
    \]
    we have that $f$ is irreducible over $\mathbb F_2$.

    As $\mathbb F_2[x]$ is a field and $\langle f \rangle$ is 
    a maximal ideal by $f$ being reducible we have that 
    \[
        \mathbb F_2[x]/\langle f \rangle
    \]
    is a field. 
\end{proof}
\begin{proof}[Solution of (b)]
    \begin{align*}
        f(\alpha) 
        &= 0 \\ 
        \alpha^3 + \alpha + 1 
        &= 0 \\ 
        \alpha^3 
        &= \alpha + 1. 
    \end{align*}
    Then 
    \begin{align*}
        \alpha^4 &= \alpha^2 + \alpha \\ 
        \alpha^5 &= \alpha^3 + \alpha^2 \\ 
                 &= \alpha^2 + \alpha + 1 \\ 
        \alpha^6 &= \alpha(\alpha^2 + \alpha + 1) \\ 
                 &= \alpha + 1 + \alpha^2 + \alpha \\ 
                 &= \alpha^2 + 1  \\ 
        \alpha^7 &= \alpha + 1 + \alpha \\ 
                 &= 1.
    \end{align*}
\end{proof}
\begin{proof}[Solution of (c)]
    \begin{align*}
        \sigma_2(\alpha) &= \alpha^2 \\ 
        \sigma_2^2(\alpha) &= \sigma_2(\alpha^2) \\ 
                           &= \alpha^4 \\ 
                           &= \alpha^2 + \alpha \\ 
        \sigma_2^3(\alpha) &= \sigma(\alpha^2 + \alpha) \\ 
                           &= (\alpha^2 + \alpha)(\alpha^2 + \alpha)
                           \\ 
                           &= \alpha^4 + 2\alpha^3 + \alpha^2 \\ 
                           &= \alpha^2 + \alpha + \alpha^2 \\ 
                           &= \alpha.
    \end{align*}
\end{proof}
\begin{proof}[Solution of (d)]
    Since $f$ has coefficients in $\mathbb F_2$, Frobenius 
    sends roots to roots. Therefore the roots of $K$ are 
    \[
        \alpha, \ \alpha^2, \ \alpha^4.
    \]
    Writing this in the basis we get roots 
    \[
        \alpha, \ \alpha^2, \ \alpha^2 + \alpha.
    \]
    Thus 
    \[
        f(x) = (x-\alpha)(x-\alpha^2)(x-\alpha^4) 
        = (x+\alpha)(x+\alpha^2)(x+\alpha^4).
    \]
\end{proof}

\begin{prob}
    Let 
    \[
        f(x) = x^3 - 5 \in \mathbb Q[x].
    \]
    Let 
    \[
        \alpha = \sqrt[3]5
    \]
    be the real cube root of $5$, and let 
    \[
        \zeta = \zeta_3 = e^{2i\pi/3}.
    \]
    \begin{enumerate}[label=\alph*)]
        \item Show that $f(x)$ is irreducible over the 
            rationals. 
        \item Find all roots of $f(x)$ in $\mathbb C$, and 
            describe the splitting field $K$ of $f$ over 
            $\mathbb Q$.
        \item Compute 
            \[
                [K:\mathbb Q].
            \]
        \item Show that 
            \[
                \operatorname{Gal}(K/\mathbb Q) \cong S_3.
            \]
        \item Describe the two automorphisms of $K$ that generate 
            the Galois group.
    \end{enumerate}
\end{prob}
\begin{proof}[Proof of (a)]
    Since $f$ is over $\mathbb Q$ with integer coefficients we can 
    use Eistenstein's criterion. Let $p = 5$. Then 
    $p$ divides $5$, but not the leading coefficient $1$ and 
    $p^2 = 25$ does not divide $5$. Thus $f$ is irreducible over 
    $\mathbb Q$.
\end{proof}
\begin{proof}[Solution of (b) and (c)]
    We have the obvious 
    \[
        f(\sqrt[3]5) = 0.
    \]
    Then we notice that 
    \[
        f(\zeta_3\sqrt[3]5) = 
        e^{2\pi}\cdot 5 - 5 
        = 0.
    \]
    Thus the roots in $\mathbb C$ are 
    \[
        \sqrt[3]5 \ \text{and} \ \zeta_3\sqrt[3]5.
    \]
    Thus the splitting field is 
    \[
        K = \mathbb Q(\sqrt[3]5, \zeta_3).
    \]
    Noticing that $f$ is actually the minimal polynomial of 
    $\sqrt[3]5$ we have 
    \[
        [\mathbb Q(\sqrt[3]5) : \mathbb Q] = 3. 
    \]
    Furthermore, generally for an $n$-th root of unity the minimal 
    polynomial is 
    \[
        \prod_{d|n} (x^{n/d})^\mu(d)
    \]
    where $\mu$ is the Möbius function. 
    So the minimal polynomial of $\zeta_3$ over $\mathbb Q(\sqrt[3]
    5)$ is 
    \[
        x^2 + x + 1
    \]  
    meaning, by the tower law,  
    \[
        [K:\mathbb Q] 
        = 2 \cdot 3 = 6.
    \]
\end{proof}
\begin{proof}[Solution of (d)]
    For the splitting field for a degree $n$ polynomial we have 
    \[
        \operatorname{Gal}(K/F) \leq S_n
    \]
    so 
    \[
        \operatorname{Gal}(K/\mathbb Q) \leq S_3. 
    \]
    Since $K/\mathbb Q$ is Galois we have 
    \[
        |\operatorname{Gal}(K/\mathbb Q)|
        = [K:\mathbb Q] = 6 = |S_3|. 
    \]
    The only subgroup of $S_3$ with $|S_3|$ elements is $S_3$ itself.
    Hence 
    \[
        \operatorname{Gal}(K/\mathbb Q)
        \cong S_3.
    \]
\end{proof}
\begin{proof}[Solution of (e)]
    Define $\sigma(\alpha) = \zeta\alpha$ and 
    $\sigma(\zeta) = \zeta$. Then 
    \[
        \sigma(\alpha) = \zeta\alpha
    \]
    and
    \[
        \sigma(\zeta\alpha) = \zeta^2\alpha
    \]
    and 
    \[
        \sigma(\zeta^2\alpha) = \alpha.
    \]
    So $\sigma$ permutes the three roots, and has 
    order $3$.

    Define $\tau(\alpha) = \alpha$, and 
    $\tau(\sigma) = \sigma^2$.
    Then 
    \[
        \tau(\alpha) = \alpha, 
    \]
    and 
    \[
        \tau(\alpha\sigma) = \alpha\sigma^2
    \]
    and 
    \[
        \tau(\alpha\sigma^2) = \alpha\sigma.
    \]
    Thus $\tau$ is complex conjugation and since $\alpha$ is fixed 
    it has order $2$.

    If we label $\alpha = 1, \alpha\zeta = 2, \alpha\zeta^2 = 3$ 
    we get 
    \[
        \sigma = (1 \ 2 \ 3), \ \tau = (2 \ 3)
    \]
    so a $3$-cycle and a transposition, which generate 
    $S_3$.
\end{proof}

\begin{prob}
    Let 
    \[
        f(x) = x^3 + x + 1 \in \mathbb F_2[x],
    \]
    and let 
    \[
        K = \frac{\mathbb F_2[x]}{\langle f(x) \rangle}.
    \]
    Write 
    \[
        \alpha = x + \langle f(x) \rangle \in K.
    \]

    \begin{enumerate}[label=\alph*)]
        \item Show that $f(x)$ is irreducible over 
            $\mathbb F_2$.
        \item Conclude that $K$ is a field. How many elements 
            does $K$ have? Give a basis for $K$ as a vector space 
            over $\mathbb F_2$.
        \item Use $f(\alpha) = 0$ to compute 
            \[
                \alpha^3, \ \alpha^4, \ \alpha^5, \ \alpha^6, 
                \ \alpha^7.
            \]
            Write each in the basis. 
        \item Show that $\alpha$ generates $K^\times$. 
        \item Let $\sigma : K \to K$, $\sigma(a) = a^2$. 
            Show that $\sigma$ is an automorphism of $K$ and 
            compute 
            \[
                \sigma(\alpha), \ \sigma^2(\alpha), 
                \ \sigma^3(\alpha).
            \]
        \item Determine the Galois group 
            \[
                \operatorname{Gal}(K/\mathbb F_2).
            \]
    \end{enumerate}
\end{prob}
\begin{proof}[Proof of (a)]
    Since $\deg f \in \{2,3\}$ we have that 
    $f$ is irreducible if and only if it 
    has no zeros in the ground-field 
    $\mathbb F_2$. 
    We quickly see that 
    \[
        f(0) = 1 = f(1)
    \]
    so $f$ has no zeros 
    in $\mathbb F_2$ and is 
    thus irreducible. 
\end{proof}
\begin{proof}[Solution of (b)]
    Since $\mathbb F_2[x]$ is a PID 
    we have that each ideal is generated 
    by some $f$, and every non-zero prime ideal 
    is a maximal ideal. Thus $\langle f \rangle$ is 
    a maximal ideal if and only if $f$ is 
    irreducible, which we have shown. 
    Since $\langle f \rangle$ is maximal we 
    then have that $K$ is a field. 

    We have $|K| = |\mathbb F_2|^{\deg f} = 8$.
    This is because since, for $g(x) \in \mathbb F_2[x]$, 
    we have $g(x) = f(x)q(x) + r(x)$ where 
    $\deg r < \deg f$. Therefore every 
    element in the quotient can be written 
    as a linear combination of 
    \[
        1, x, x^2 
    \]
    with scalars in $\mathbb F_2$. 
    So we have $2$ choices for $1$, 
    $2$ for $x$ and $2$ for $x^2$. 
    Hence $|K| = 2^3 = 8$.

    If $\alpha = x + \langle f(x) \rangle$, we have 
    the standard basis 
    $\{1, \alpha, \alpha^2\}$. 
\end{proof}
\begin{proof}[Solution of (c)]
    We have 
    \[
        \alpha^3 + \alpha + 1 = 0
    \]
    so rearranging, and recalling that 
    addition and subtraction are interchangable 
    in $\mathbb F_2$ we get 
    \[
        \alpha^3 = \alpha + 1. 
    \]
    Then 
    \[
        \alpha^4 = \alpha^2 + \alpha. 
    \]
    Next 
    \[
        \alpha^5 = \alpha^3 + \alpha^2 = \alpha^2 + \alpha + 1. 
    \]
    Next 
    \[
        \alpha^6 = \alpha^3 + \alpha^2 + \alpha 
        = \alpha^2 + 1. 
    \]
    Last 
    \[
        \alpha^7 = \alpha^3 + \alpha 
        = 1. 
    \]
\end{proof}

\begin{proof}[Proof of (d)]
    We have that $|\alpha| = 7$ by (c), so 
    $|\langle \alpha \rangle| = 7$.
    We have that $|K| = 8$ and since 
    the finite fields are characterized up 
    to isomorphism we have 
    \[
        K \cong \mathbb F_8
    \]
    which means that $K^\times \cong C_7 
    \leq \mathbb F_8$ (as groups).
    Since $\alpha$ generates 
    a subgroup of $\mathbb F_8$ which has 
    $7$ elements and the only subgroup of 
    $\mathbb F_8$ with $7$ elements is 
    $C_7$ we conclude that 
    \[
        \langle \alpha \rangle 
        \cong C_7 
        \cong K.
    \]
\end{proof}
\begin{proof}[e]
    We confirm the homomorphism properties first. 
    \begin{align*}
        \sigma(a+b) 
        &= (a+b)^2 \\ 
        &= a^2 + 2ab + b^2 \\ 
        &= a^2 + b^2 \\ 
        &= \sigma(a) + \sigma(b)
    \end{align*}
    since $2 \equiv 0$. 
    Then 
    \begin{align*}
        \sigma(ab) 
        &= (ab)^2 \\ 
        &= a^2b^2 \\ 
        &= \sigma(a)\sigma(b).
    \end{align*}

    Let $\phi : \mathbb F_8 \to \mathbb F_8$ defined 
    by $\phi(a) = a^2$. Then suppose 
    $\phi(a) = 0$. Then $a^2 = 0$ so 
    $a = 0$. This means the map is injective since 
    $\ker \phi = \{0\}$. Furthermore we se see that 
    \[
        0 \mapsto 0, \ 1 \mapsto 1, \ 2 \mapsto 4, \ldots
    \] 
    and we can confirm for ourselves in this way that 
    the map is indeed bijective. 
    Now let $\psi$ be the isomorphism from 
    $K$ to $\mathbb F_8$. Such a map exists since 
    we have established that $K \cong \mathbb F_8$. 
    Then $\sigma : K \to K = \psi \circ \phi \circ \psi^{-1}$, i.e. 
    a composition of isomorphisms and therefore an isormorphism 
    on $K$. Hence $\sigma$ is an automorphism of 
    $K$.

    Then 
    \begin{align*}
        \sigma(\alpha) 
        &= \alpha^2 \\ 
        \sigma^2(\alpha ) 
        &= \alpha^4 = \alpha^2 + \alpha \\ 
        \sigma^3(\alpha) &= 
        \alpha^8 = \alpha(\alpha^7) = \alpha. 
    \end{align*}
\end{proof}
\begin{proof}[Solution of (f)]
    We use the standard result that 
    for $\operatorname{Gal}
    (\mathbb F_{q^n}/\mathbb F_q)$, we have 
    that the galois group is generated by the 
    $s$-th power of the 
    Frobenius automorphism where $s$ is the 
    characteristic of the field. 
    Thus 
    \[
        \operatorname{Gal}(K/\mathbb F_2) 
        = \langle \sigma \rangle.
    \]
    As we saw we have 
    $|\sigma^2| = 3$ which is actually 
    inline with $[K : \mathbb F_2]$ even 
    though the extension is not automatically 
    Galois as it would be if we we're working over 
    $\mathbb Q$ for instance.

    In particular we have 
    \[
        \operatorname{Gal}(K/\mathbb F_2) 
        = \mathbb Z/3 \mathbb Z.
    \]
\end{proof}

\begin{prob}
    Let 
    \[
        \phi : \mathbb Z[x] \to \mathbb Z_6
    \]
    be the ring homomorphism defined by 
    \[
        \phi(f(x)) = f(2) \mod 6. 
    \]

    \begin{enumerate}[label=\alph*)]
        \item Show that $\phi$ is a ring homomorphism. 
        \item Find $\ker \phi$. 
            Try to describe it as an ideal generated by 
            simple elements of $\mathbb Z[x]$. 
        \item Identify 
            \[
                \mathbb Z[x] / \ker \phi. 
            \]
        \item Is $\ker \phi$ prime? 
        \item Is $\ker \phi$ maximal? 
        \item Replace $\mathbb Z_6$ with $\mathbb Z_5$. Define 
            \[
                \psi : \mathbb Z[x] \to \mathbb Z_5
            \]
            by 
            \[
                \psi(f(x)) = f(2) \mod 5. 
            \]
            Find the kernel, and decide whether it is 
            prime or maximal.
    \end{enumerate}
\end{prob}
\begin{proof}[Solution of (b)]
    $\ker \phi = \{f(x) \in \mathbb Z[x] : \phi(f(x)) = 0\}$. This set includes 
    $f(x) = 6n \cdot g(x), \ n \in \mathbb Z, \ g(x) \in \mathbb Z[x]$,
    $(x-2) \cdot g(x), \ g(x) \in \mathbb Z[x]$. 
    It is therefore not hard to see that 
    \[
        \ker \phi = \langle x-2, 6 \rangle, 
    \]
    since $x-2, 6 \in \ker \phi$ and for some 
    $f(x) \in \ker \phi$ we have that 
    $f(x) = 0$ for $x = 2$ or $f(x) \equiv 0$ 
    modulo $6$. 
\end{proof}
\begin{proof}[Solution of (c)]
    By the first isomorphism theorem we have 
    \[
        \mathbb Z[x] / \ker \phi \cong \operatorname{im} \phi
    \]
    and since $\phi$ is trivially surjective 
    (for $n \in \mathbb Z_6$ consider the plynomial 
    $f(x) = n$), we have that 
    \[
        \mathbb Z[x] / \ker \phi \cong \mathbb Z_6.
    \]
\end{proof}
\begin{proof}[Proof of (d) and (e)]
    Since $\mathbb Z[x]$ is a commutative ring 
    we have that $\mathbb Z[x]/\ker \phi$ is 
    an integral domain iff $\ker \phi$ is 
    prime and said quotient ring is 
    a field iff $\ker \phi$ is maximal.
    As we have identified the quotient ring 
    to be isomorphic to $\mathbb Z_6$ we can 
    answer no to both of these questions since 
    $2 \cdot 3 = 6 \equiv 0$ which disqualifies 
    the quotient ring from being an integral domain 
    which disqualifies it from being a field. 
    
    Thus $\ker \phi$ is neither maximal nor prime. 
\end{proof}
\begin{proof}[Proof of (f)]
    By the first isomorphism theorem and the same 
    arguments as earlier we have 
    \[
        \mathbb Z[x]/\ker \psi \cong \mathbb Z_5
    \]
    and $\mathbb Z_5$ is indeed a field 
    (since $\mathbb Z_q$ is a field when $q$ 
    is prime) and thus $\ker \psi$ is maximal 
    which implies that it is also prime.
\end{proof}

\begin{prob}
    Let 
    \[
        f(x) = x^3 - 2 \in \mathbb Q[x]. 
    \]
    Let 
    \[
        \alpha = \sqrt[3]2
    \]
    be the real cube root of $2$, 
    and let 
    \[
        \omega = e^{2\pi/3}
    \]
    be a primitive third root of unity. 
    \begin{enumerate}[label=\alph*)]
        \item Show that $f(x)$ is irreducible over 
            $\mathbb Q$.
        \item Find all roots of $f$ in 
            $\mathbb C$, and show that the splitting 
            field is 
            \[
                K = \mathbb Q(\alpha, \omega). 
            \]
        \item Compute 
            \[
                [\mathbb Q(\alpha) : \mathbb Q].
            \]
        \item Conclude that 
            \[
                [K : \mathbb Q] = 6.
            \]
        \item Show that every $\mathbb Q$-automorphism 
            of $K$ is determined by the images 
            of $\alpha$ and $\omega$.
        \item Construct two automorphisms $\sigma,\tau 
            \in \operatorname{Gal}(K/\mathbb Q)$ by 
            \[
                \sigma(\alpha) = \omega\alpha, \ \sigma(\omega) = \omega, 
            \]
            and 
            \[
                \tau(\alpha) = \alpha, \ \tau(\omega) = \omega^2.
            \]
            Show that these are valid automorphisms. 
        \item Compute 
            \[
                \sigma^3, \ \tau^2, \tau\sigma\tau^{-1}.
            \]
            Conclude that 
            \[
                \operatorname{Gal}(K/\mathbb Q)\cong S_3.
            \]
    \end{enumerate}
\end{prob}
\begin{proof}[Proof of (a)]
    $f$ is a polynomial with integer coefficients so 
    we can use Eisenstein's criterion. 
    We see that for $p = 2$ we have 
    \[
        p \mid a_0 = -2 
    \]
    and the other terms 
    which are $0$
    while 
    \[
        p \nmid a_3 = 1, 
    \]
    and 
    \[
        4 = p^2 \nmid a_0 = 2.
    \]
    Thus, by Eisenstein's criterion, $f$ is irreducible 
    over $\mathbb Q$.
\end{proof}
\begin{proof}[Proof of (b)]
    Over $\mathbb C$ we have 
    \[
        f(\alpha) = f(\sqrt[3]2) = (\sqrt[3]2)^3 - 2 = 0, 
    \]
    and 
    \[
        f(\omega\alpha) = f(\omega\sqrt[3]2) = 0.  
    \]
    These are the obvious ones. Next up we see that 
    \[
        f(\omega^2\alpha) = 0
    \]
    as well.
    Then, we have 
    \[
        K \subseteq \mathbb Q(\alpha, \omega),
    \]
    since $\alpha,\omega\alpha,\omega^2\alpha \in 
    \mathbb Q(\alpha,\omega)$.
    
    For the other direction we can see that 
    $\omega,\alpha$ generate $K$ when 
    adjoining to $\mathbb Q$ so 
    $E \subseteq K$. 
\end{proof}
\begin{proof}[Proof of (c)]
    We have that 
    $f$ is the minimal polynomial 
    of $\sqrt[3]2$ over $\mathbb Q$ so 
    \[
        [\mathbb Q(\alpha) : \mathbb Q] = 
        \deg f = 3. 
    \]
    Now since $\omega \in \mathbb C$ 
    we have $\omega \not\in \mathbb Q(\alpha)$ 
    because $\mathbb Q(\alpha)$. 
    We can see that this is the case because 
    \[
        \mathbb Q(\alpha) 
        = \{a + b\sqrt[3]2 \bigm| a,b \in \mathbb Q\}, 
    \]
    and $a,b,\sqrt[3]2 \in \mathbb R$. 
    Since $\mathbb C \supset 
    \mathbb R\supset \mathbb Q(\alpha)$ we conclude 
    that the claim is true.

    The minimal polynomial of an 
    $n$-th root of unity is 
    \[
        \prod_{d| n}(x^d - 1)^{\mu(n/d)}
    \]
    which, for prime $n$ works out nicely to 
    \[
        \sum_{0 \leq k \leq n-1} x^n. 
    \]
    In this case the primitive polynomial is 
    \[
        x^2 + x + 1
    \]
    and thus 
    \[
        [\mathbb Q(\alpha,\omega) : \mathbb Q(\alpha)] = 2. 
    \]
    Therefore, by the tower law, we have 
    \[
        [\mathbb K : \mathbb Q] = 2 \cdot 3 = 6.
    \]
\end{proof}
\begin{proof}[Proof of (d)]
    A $\mathbb Q$-automorphism of $K$ 
    is $\psi : K \to K$ such that 
    \[
        \psi[\mathbb Q] = \mathbb Q. 
    \]
    Thus, for every nonidentity $\mathbb Q$-automorphism 
    on $K$ we have 
    \[
        \psi(\alpha) = \beta
    \]
    where $\alpha,\beta$ are roots of $f$. 

    We have every root is sent to some root so 
    the images of $\omega$ must be fixed 
    or sent to $\omega^2$. If $\omega \mapsto 
    \alpha$, for instance we would have 
    $\psi(\omega\alpha) = \psi(\alpha^2) \neq 0$. 
    $\alpha$ can be sent to $\alpha\omega$. 
\end{proof}
\begin{proof}[Proof of (f)]
    We will show $\operatorname{Gal}(K/\mathbb Q) \cong 
    S_3$ first. 
    First since $K$ is algebraic over $\mathbb Q$ and 
    the splitting field of 
    a $\mathbb Q[x]$ polynomial and 
    $\mathbb Q$ is characteristic $0$ we get 
    that $K$ is a normal- (by the former) and 
    separable (by the latter) extension, since 
    every algebraic axtension of a characteristic $0$  
    field is separable.
    
    Thus $K/\mathbb Q$ is Galois and therefore 
    \[
        [K:\mathbb Q] = |\operatorname{Gal}(K/\mathbb Q)|. 
    \]
    Since $\operatorname{Gal}(K/\mathbb Q) \leq S_3$ (from $K$ being 
    splitting field over $\deg f = 3$ polynomial) we have 
    \[
        \operatorname{Gal}(K/\mathbb Q) \cong S_3.
    \]
\end{proof}

\begin{prob}
    Let 
    \[
        K = \mathbb Q(\sqrt 2, \sqrt 3).
    \]
    \begin{enumerate}[label=\alph*)]
        \item Prove carefully that 
            \[
                [\mathbb Q(\sqrt 2):\mathbb Q] = 2.
            \]
            Then prove that 
            \[
                \sqrt 3 \not\in \mathbb Q(\sqrt 2). 
            \]
            Conclude that 
            \[
                [K : \mathbb Q] = 4. 
            \]
        \item Show that 
            \[
                K
            \] 
            is the splitting field over $\mathbb Q$ of 
            \[
                f(x) = (x^2 - 2)(x^2 - 3).
            \]
            Conclude that 
            \[
                K/\mathbb Q
            \]
            is Galois.
        \item Show that every $\mathbb Q$-automorphism 
            of $K$ is determined by the images 
            of $\sqrt 2$, and $\sqrt 3$. 
            Show the possible images and conclude 
            that 
            \[
                \operatorname{Gal}(K/\mathbb Q) \leq 4.  
            \]
            Construct all $4$ automorphisms and 
            conclude that  
            \[
                \operatorname{Gal}(K/\mathbb Q)
                \cong C_2 \times C_2.
            \]
    \end{enumerate}
\end{prob}
\begin{proof}[Proof of (a)]
    We have that, for $f(x) = x^2 - 2 \in 
    \mathbb Q[x]$, 
    $\sqrt 2$ is a $0$. 
    This polynomial is monic, and furthermore 
    no lower degree nonzero polynomial can have 
    $\sqrt 2$ as a zero. If we suppose 
    that $f'$ is such a polynomial, then 
    \[
        f'(x) = x - a 
    \]
    such that $\sqrt 2 - a = 0$, meaning 
    $\sqrt 2 = a$, but then $f'(x)\not\in 
    \mathbb Q[x]$.
    Thus $f$ is the minimal polynomial of 
    $\sqrt 2$.
    Then 
    \[
        [\mathbb Q(\sqrt 2) : \mathbb Q] 
        = \deg f = 2.
    \]

    Suppose $\sqrt 3 \in \mathbb Q(\sqrt 2)$. 
    Then, since $\sqrt 3 \not \in \mathbb Q$, we 
    have $\sqrt 3 \in \mathbb Q(\sqrt 2) \setminus 
    \mathbb Q$.
    Then $\sqrt 3 = a + b\sqrt 2$ such that 
    $b \neq 0$. Then 
    \[
        3 = a^2 + 2ab\sqrt 2 + b^2. 
    \]
    If $a \neq 0$ we have 
    \[
        3 - a^2 - b^2 = 2ab\sqrt 2, 
    \]
    but the left-hand-side is rational and 
    the right-hand side is not, a contradiction. 
    Suppose $a = 0$. Then 
    \[
        3 = b^2 
    \]
    so $b = \sqrt 3 \not\in \mathbb Q$. 
    Thus $\sqrt 3 \not\in \mathbb Q(\sqrt 2)$.
    By a similar argument as earlier, the minimal 
    polynomial of $\sqrt 3$ over $\mathbb Q(\sqrt 2)$ is 
    $x^2 - 3$. 
    Thus, by the tower law, we have 
    \[
        [\mathbb Q(\sqrt 2, \sqrt 3) : \mathbb Q] = 2 \cdot 2 = 4. 
    \]
\end{proof}
\begin{proof}[Proof of (b)]
    Let $L$ be the splitting field of $f$. 
    It is trivial to see that 
    \[
        f(\pm\sqrt 2) = f(\pm\sqrt 3) = 0
    \]
    are the zeros of $f$ in $\mathbb C \supset 
    \overline{\mathbb Q}$.  
    Therefore $L \subseteq K$.
    The basis of $K$ is $\{1,\sqrt 2, \sqrt 3, \sqrt 6\}$, 
    and so the first $3$ are contained in $L$ and furthermore 
    \[
        \sqrt 6 = \frac{3\sqrt 2}{\sqrt 3} \in L.
    \]
    Therefore $K = L$. 

    $L = K$ is algebraic over $\mathbb Q$ as 
    it contains $\mathbb Q$ and elements which 
    are the zeros of $f \in \mathbb Q[x]$. 
    Furthermore since $L=K$ is the splitting field 
    of a polynomial in $\mathbb Q[x]$ it is a 
    normal extension. Lastly since 
    $\mathbb Q$ is characteristic $0$, 
    we have that all algebraic extensions are 
    separable. 

    All three properties give us that 
    $L=K$ is Galois as an extension of 
    $\mathbb Q$.
\end{proof}

\begin{proof}[Proof of (c)]
    The $\mathbb Q$-autmorphisms of $K$ are 
    those which fix $\mathbb Q$ and 
    preserve 
    \[
        f(a) = 0
    \]
    iff 
    \[
        f(\psi(a)) = 0. 
    \]
    We require 
    \[
        (x^2 - 2)(\sqrt 2) = 0
    \]
    so $\sqrt 2 \mapsto \pm 2$ since 
    \[
        (x^2 - 2)(a) = 0
    \]
    if $a \in \{\sqrt 2, -\sqrt 2\}$. 
    Similarly for $\sqrt 3 \mapsto \pm \sqrt 3$. 

    Then we have a maximum of $4$ possible 
    isomorphisms from $K \to K$ which 
    respect this. 


    There are two choices to make 
    with two options each. For each 
    valid $\mathbb Q$-automorphism we have 
    send each of $\sqrt 2$ and $\sqrt 3$ 
    to themselves or their additive inverse. 
    Thus $\operatorname{Gal}(K/\mathbb Q) \cong C_2 \times C_2$. 
\end{proof}

\begin{prob}
    Let $G$ be a finite group and let 
    $H \leq G$ be a subgroup of index 
    $n$. Let 
    \[
        X = G / H 
    \]
    be the set of left cosets of $H$.

    Define an action on $X$ by 
    \[
        g \cdot (aH) = (ga) H. 
    \]

    \begin{enumerate}[label=\alph*)]
        \item Prove carefully that this is a 
            group action. 
        \item Let $\phi : G \to S_x$ 
            be the homomorphism induced by 
            this action. 
            Prove that 
            \[
                \ker \phi = 
                \bigcap_{g\in G} gHg^{-1}.
            \]
            This subgroup is often called the 
            core of $H$ in $G$.
        \item Prove that $\ker \phi$ is the 
            largest normal subgroup of $G$ 
            contained in $H$.
        \item Assume now that 
            \[
                |G| = 21
            \]
            and that $H \leq G$ has 
            order $7$. 

            Use the action to prove that 
            $H \trianglelefteq G$. 
        \item Use Cauchy's theorem to prove that 
            every group of order $21$ has 
            a normal subgroup of order $7$.
    \end{enumerate}
\end{prob}
\begin{proof}[Proof of (a)]
    We must prove the following: 
    \begin{enumerate}
        \item if $aH = bH$, then 
            $(ga)H = (gb)H$,
        \item $e \cdot (aH) = aH$, 
        \item $g_1 \cdot (g_2 \cdot H) = (g_1g_2) \cdot H$.
    \end{enumerate}
    We begin with the first.

    Suppose $aH = bH$. We defined the 
    action to be $g \cdot (aH) = (ga)H$ so 
    \begin{align*}
        aH &= bH \\ 
        g(aH) &= g(bH) \\ 
        (ga)H &= (gb)H.
    \end{align*}
    
    Next the second property: 
    \begin{align*}
        e \cdot aH 
        &= (ea)H \\ 
        &= (a)H \\ 
        &= aH.
    \end{align*}

    The last property is also evident from 
    the definition 
    \[
        g_1 (g_2 H) \underset{\text{def}}{=} (g_1g_2)H.
    \]
    
\end{proof}

\begin{proof}[Proof of (b)]
    The homomorphism induced by $G$ 
    acting on $X$ is $\psi : G \to \operatorname{Sym}(X)$, 
    defined by 
    \[
        \psi(g)(aH) = g \cdot aH = (ga)H. 
    \]

    We want to find 
    \[
        \ker \phi = \{g \in G : \psi(g) = e\}.
    \]
    That is, we want to find all $g \in G$ such that 
    $\psi(g)(aH) = aH$ for every $aH \in X$.

    We remark that 
    \[
        aH = bH \Leftrightarrow b^{-1}a \in H, 
    \]
    and we need to find $g \in G$ such that 
    \[
        gxH = xH, \ \forall x \in G.
    \]
    Then we are looking for $g \in G$ such that 
    \[
        x^{-1}gx \in H
    \]
    for every $x \in G$. 

    This is equivalent to 
    \[
        g \in xHx^{-1} 
    \]
    for every $x \in G$. 

    Summing up all these equivalences we get 
    \[
        \ker \phi 
        = \bigcap_{g \in G} gHg^{-1}.
    \]

\end{proof}
\begin{proof}[Proof of (c)]
    The map $\psi$ is a group homomorphism. Then 
    $\ker \psi$ is a normal subgroup of the domain 
    $\ker \psi \trianglelefteq G$. 

    We have $\ker \psi = \bigcap_{g \in G} gHg^{-1}$ 
    from (b). For $g = e$ we have 
    \[
        \bigcap_{g\in G} \subseteq eHe^{-1} = H. 
    \]

    Let $N$ be some normal subgroup of $G$ such that 
    $N \subseteq H$. We must show that 
    $N \subseteq \ker \psi$. Since $N$ is invariant 
    under conjugation we have 
    \[
        N = gNg^{-1}
    \]
    for every $g \in G$.
    Since $N \subseteq H$, we have 
    \[
        N = gNg^{-1} \subseteq gHg^{-1}, \ \forall g \in G.
    \]
    Thus 
    \[
        N \subseteq \bigcap_{g \in G} gHg^{-1} 
        = \ker \psi.
    \]
\end{proof}
\begin{proof}[Proof of (d)]
    From Lagrange's theorem we have 
    \[
        |X| = \left|\frac{G}{H}\right|
        = \frac{|G|}{|H|} = 3. 
    \]
    The action induces the 
    homomorphism 
    $\psi : G \to S_3$. 

    Thus by the first isomorphism theorem 
    we have that 
    \[
        \frac{G}{\ker \psi} \cong \operatorname{im}
        \phi \leq S_3. 
    \]
    Thus 
    \[
        \left|\frac{G}{\ker \psi}\right| \leq 6. 
    \]
    We have $\ker \psi \subseteq H$, so 
    $|\ker \psi| \leq 7$. We must also have that 
    the order of the kernel divides $|G| = 21$. 
    We have $\ker \psi \geq 3$ by the constraint 
    of the quotient group having $\leq 6$ elements. 
    Then, $4,5,6$ do not divide $21$ so we must 
    have that $|\ker \psi| = 7$. By $(c)$, 
    $\ker \psi$ is the largest normal subgroup 
    of $G$ contained entirely in $H$, but 
    $|\ker \psi| = |H|$, so $H = \ker \psi$. 
    Thus $H$ is normal.
\end{proof}
\begin{proof}[Proof of (e)]
    Cauchy's theorem tells us that if 
    $G$ is a finite group and $p$ is prime 
    such that $p$ divides $|G|$. Then 
    there exists an element in $G$ with 
    order $p$. Let $g$ be that element. Then
    $\langle g \rangle \leq G$. Apply this for 
    $p = 7$ so that we get 
    $|\langle g \rangle| = 7$. Apply (d) 
    to see that the kernel of the homomorphism 
    induced by the group action on the set of 
    cosets of $\langle g \rangle$ to see that 
    $\langle g \rangle$ is normal. 
\end{proof}

\begin{prob}
    Let 
    \[
        R = \mathbb Z \times \mathbb Z
    \]
    with componentwise addition and multiplication. 
    \begin{enumerate}[label=\alph*)]
        \item Find all units in $R$. 
        \item Find all zero divisors in $R$. 
        \item Prove that every ideal $I \trianglelefteq R$ 
            has the form 
            \[
                I = m\mathbb Z \times n \mathbb Z
            \]   
            for some integers $m,n \geq 0$. 
        \item Let 
            \[
                J = 2\mathbb Z \times 3 \mathbb Z.
            \]
            Identify the quotient 
            \[
                \frac{R}{J}
            \]
            up to isomorphism. 
        \item Is $J$ prime, maximal? 
        \item Find all pairs $(m,n)$ such that 
            \[
                m\mathbb Z \times n \mathbb Z
            \]
            is a prime ideal of $R$.
    \end{enumerate}
\end{prob}
\begin{proof}[Solution of (a)]
    We want to find $(a,b) \in R$ for which there 
    exists $(d,c) \in R$ such that 
    \[
        (a,b)(d,c) = (1,1),
    \]
    meaning 
    \[
        ad = 1, \ \text{and} \ bc = 1. 
    \]
    Then $a^{-1} = d$ and $b^{-1} = c$ in 
    $\mathbb Z$. The only units in 
    $\mathbb Z$ are $1$ and $-1$ both having 
    themselves as inverses. Thus 
    the units are $(a,b) = (1,1), (a,b) = (1,-1), 
    (a,b) = (-1, 1), (a,b) = (-1,-1)$ with 
    themselves as the multiplicative inverse 
    $(a,b)^{-1} = (c,d) = (a,b)$. 
\end{proof}
\begin{proof}[Solution of (b)]
    We want to find nonzero 
    $(a,b)$ for which there exists 
    nonzero $(c,d)$ such that 
    \[
        (a,b)(c,d) = (0,0).
    \]
    Therefore $(0,0)$ is out of the question. 
    We then want to find $a,b,c,d \in \mathbb Z$ such 
    that 
    \[
        ab = 0 \ \text{and} \ cd = 0, 
    \]
    but we forbid $a = b = 0$ or $c = d = 0$ 
    similtaneously. 
    Thus we can choose e.g. $a, d \in \mathbb Z\setminus\{0\}$, 
    $b = c = 0$ and vice-versa. 

    Then the zero-divsors are all pairs 
    \[
        (a,b),(c,d) \ \text{s.t.} \ 
        a,d \in \mathbb Z \setminus \{0\} \ 
        \text{and} \ b = c = 0 \ \text{or} \ 
        a = d = 0 \ \text{and} \ 
        b,c \in \mathbb Z\setminus\{0\}.
    \]
\end{proof}
\begin{proof}[Proof of (c)]
    It is obvious from the problem statement that 
    we are talking about nontrivial proper ideals. 

    We will first prove, for arbitrary 
    $m,n \in \mathbb Z$, that 
    $m\mathbb Z \times n \mathbb Z$ is 
    an ideal (two-sided, though all ideals of 
    $R$ are two-sided since it is commutative) of 
    $R$. 

    $m\mathbb Z\times n\mathbb Z$ is the 
    direct product of the additive groups 
    $m\mathbb Z, n \mathbb Z$ and is therefore 
    itself an additive group. It is 
    a subgroup of $\mathbb Z \times \mathbb Z$. 

    Suppose $(a,b) \in m \mathbb Z \times n \mathbb Z$ 
    and that $(c,d) \in R$. Then $(a,b) = (m\alpha, n\beta)$ for 
    some $\alpha,\beta \in \mathbb Z$. 
    Then 
    \[
        (c,d)(a,b) = (c \cdot m \alpha,d \cdot n \beta) 
        = (m (c\alpha), n (d\beta)) \in m\mathbb Z \times n \mathbb Z, 
    \]
    and likewise for right-multiplication.
    Thus $m \mathbb Z \times n \mathbb Z$ is an ideal 
    of $R$ for arbitary $m,n$. 

    Suppose $I$ is an ideal of $R$. 
    Then $I$ is an additive subgroup of $R$ and 
    absorbs multiplication from $R$.

    The additive group $(R, +)$ is 
    $(\mathbb Z \times \mathbb Z, +)$, a (finitely generated) 
    abelian group. Thus every subgroup is a finitely generated 
    abelian group, meaning that for every subgroup $I \leq R$, we 
    have 
    \[
        I \cong \prod_{i=1}^k \mathbb Z_{p_i^{e_i}} \times 
        \prod_{i=1}^l \mathbb Z,
    \]
    by the fundamental theorem of finitely generated abelian 
    groups.
    From this it is clear that $I \cong m \mathbb Z \times n \mathbb Z$ 
    or $I \cong m \mathbb Z \times \mathbb Z$. 
\end{proof}
\begin{proof}[Solution of (d)]
    We recall that $\mathbb Z/n\mathbb Z \cong \mathbb Z_n$ and 
    thus immediately suspect 
    \[
        \frac{\mathbb Z\times \mathbb Z}{2\mathbb Z \times 3 \mathbb Z }
        \cong \mathbb Z_2 \times \mathbb Z_3.
    \]
    Define the homomorphism 
    $\phi : R \to \mathbb Z_2 \times \mathbb Z_3$ by 
    \[
        \phi(a,b) = (a \mod 2, b \mod 3). 
    \]
    Clearly $R$ is a homomorphism and 
    surjective. Thus $\operatorname{im}\phi = 
    \mathbb Z_2 \times \mathbb Z_3$. 
    Furthermore $\phi(a,b) = 0$ if and only if 
    $a \cong 0 \mod 2$ and $b \cong 0 \mod 3$, 
    meaning $a = 2x, b = 3y$ so 
    \[
        \ker \phi = 2\mathbb Z \times 3 \mathbb Z.
    \]
    By the first isomorphism theorem we have 
    \[
        \frac{\mathbb Z\times 
        \mathbb Z}{2\mathbb Z \times 3 \mathbb Z} = 
        \frac{R}{\ker \phi} \cong  
        \operatorname{im} \phi = 
        \mathbb Z_2 \times \mathbb Z_3.
    \]
\end{proof}
\begin{proof}[Proof of (e)]
    $R$ is a commutative ring so the 
    quotient ring is a field if and only if 
    the ideal is maximal. Similarly the 
    quotient is an integral domain if 
    and only if the quotient is a 
    prime ideal. 

    Thus we can see that the quotient is neither 
    since $(1,0)(0,1) = (0,0)$ in $\mathbb 
    Z_2 \times \mathbb Z_3$. 
\end{proof}
\begin{proof}[Proof of (e)]
    Only $(m,n) \in \{(1,0),(1,p),(0,1),(p,1)\}$ 
    where $p$ are prime give us prime ideals. 
    Then the maximal ideals are a subset of this 
    and clearly they are those $(m,n) \in \{(p,1),(1,p)\}$.
    Up to isomorphism these sets are cut in half.
\end{proof}
\begin{prob}
    Let $G$ be a finite abelian group with 
    $|G| = 72 = 2^3 \cdot 3^2$.
    \begin{enumerate}[label=\alph*)]
        \item List all abelian groups of order 
            $|G|$ up to ismorphism.
        \item Which of the groups are cyclic?
        \item Let 
            \[
                A = (\mathbb Z_2)^3 \times 
                (\mathbb Z_3)^2.
            \]
            How many elements of order exactly 
            $6$ exist?
        \item Let 
            \[
                B = \mathbb Z_4 \times 
                \mathbb Z_2 \times \mathbb Z_9.
            \]
            Find an element of order $36$ or 
            prove that no such elements exists. 
        \item Suppose $G$ is an abelian group 
            of order $72$ and contains an 
            element of order $72$. Prove 
            $G \cong \mathbb Z_{72}$.
            Suppose $G$ contains an element 
            of order $36$. Is $G$ cyclic or 
            can you find a counter-example?
    \end{enumerate}
\end{prob}
\begin{proof}[Solution of (a)]
    We use a special case of the fundamental 
    theorem of finitely generated abelian groups, 
    that being the classification theorem for 
    finite abelian groups. We have 
    \[
        G \cong \prod_{i}\mathbb Z_{p_i^{e_i}}
    \] 
    such that 
    \[
        \prod_{i}p_i^{e_i} = 72. 
    \]
    We must then find the unique 
    ways of doing this. 
    We have $72 = 2^3 \cdot 3^2$. 
    Then we get the following groups 
    \begin{align*}
        (\mathbb Z_2)^3 \times (\mathbb Z_3)^2 \\ 
        \mathbb Z_2 \times \mathbb Z_2^2 \times (\mathbb Z_3)^2 \\ 
        \mathbb Z_2^3 \times (\mathbb Z_3)^2 \\ 
        \mathbb Z_2^3 \times \mathbb Z_3^2 \\ 
        \mathbb Z \times \mathbb Z_2^2 \times \mathbb Z_3^2 \\ 
        (\mathbb Z_2)^3 \times \mathbb Z_3^2. 
    \end{align*}
\end{proof}
\begin{proof}[Solution of (b)]
    Each $\mathbb Z_n$ is cyclic and the direct 
    product of such groups is cyclic if and only if 
    their orders are relatively prime. 

    That leaves only 
    \begin{align*}
        \mathbb Z_8 \times \mathbb Z_9. 
    \end{align*}

    This group is isomorphic to 
    $\mathbb Z_{72}$ which can be shown 
    with the first isomorphism theorem 
    using the surjective $\phi : \mathbb Z \to 
    \mathbb Z_8 \times \mathbb Z_9$, 
    \[
        \phi(x) = (x \mod 8, x \mod 9),
    \]
    and identifying that $\ker \phi = \mathbb Z_{72}$.
\end{proof}

\begin{proof}[Proof of (c)]
    We notice that 
    $1$ has order $2$ in 
    $\mathbb Z_2$ and 
    $1$ and $2$ have order $3$ in 
    $\mathbb Z_3$. Then 
    in the direct product the order is 
    the least common multiple of the 
    entries. Thus 
    \[
        (a,b,c,d,e)
    \]
    has order $6$ when at least one 
    of $a,b,c$ is non-zero and at least 
    one of $d,e$ is non-zero. 
    Then we can make $2$ choices for 
    $a,b,c$ since one must be non-zero 
    and $1$ choice for $d,e$ for 
    the same reason. Thus there are 
    $2^2 \times 3 = 12$ such elements. 
\end{proof}
\begin{proof}[Proof of (d)]
    We have $1$ having 
    order $n$ in $\mathbb Z_n$. 
    Therefore $(1,1,1)$ has order 
    $\operatorname{lcm}(2,4,9) = 36$ in $B$. 
\end{proof}
\begin{proof}[Proof of (e)]
    We begin with the second claim. 
    $B$ as defined in (d) has an element of 
    order $36$ and is not cyclic. 

    It is evident that $\mathbb Z_{72}$ has an 
    element of order $72$, namely $1$. 

    If $|G| = 72$ and there exists 
    $g \in G$ such that $|g| = 72$ then 
    $
        \langle g \rangle
    $
    is a subgroup of $G$ with order $72$, 
    meaning 
    \[
        \langle g \rangle 
        \cong C_{72} \cong \mathbb Z_{72}. 
    \]
    The only subgroup of $G$ with order $|G|$, 
    is $G$ so $\langle g \rangle = G$. 
    Thus 
    \[
        G \cong \mathbb Z_{72}.
    \]
      
\end{proof}

\begin{prob}
    Let $E / \mathbb Q$ be a finite Galois extension such 
    that 
    \[
        \operatorname{Gal}(E/\mathbb Q) \cong S_5. 
    \]
    Let 
    \[
        G = \operatorname{Gal}(E/\mathbb Q).
    \]
    \begin{enumerate}
        \item Compute $[E:\mathbb Q]$.
        \item Prove that there exists an 
            intermediate field $K$ so that 
            \[
                [K : \mathbb Q] = 24. 
            \]
        \item Prove that there exists an intermediate 
            field $L$ such that  
            \[
                [L : \mathbb Q] = 10.
            \]
        \item Does there necessarily exist an intermediate field 
            $M$ such that 
            \[
                [M : \mathbb Q] = 7? 
            \]
        \item Let $A_5 \leq S_5$. Let $K = E^{A_5}$, the fixed 
            field of $A_5$.
            Compute 
            \[
                [K : \mathbb Q]
            \]
            and 
            \[
                [E:K].
            \]  
            Decide whether $K / \mathbb Q$ is Galois.
        \item Let 
            \[
                H = (\langle 1 \ 2 \ 3 \ 4 \ 5 )\rangle \leq S_5. 
            \]
            Let $K = E^H$. 
            Compute 
            \[
                [K : \mathbb Q]
            \]
            and 
            \[
                [E : K].
            \]
            Is $K / \mathbb Q$ necessarily Galois?
    \end{enumerate}
\end{prob}
\begin{proof}[Solution of (a)]
    The extension $E/\mathbb Q$ is Galois. 
    Then 
    \[
        |\operatorname{Gal}(E/\mathbb Q)| = 
        [E : \mathbb Q]. 
    \]
    Thus 
    \[
        [E : \mathbb Q] 
        = |S_5| = 120.
    \]
\end{proof}
\begin{proof}[Proof of (b)]
    We have $H:=\langle (1 \ 2 \ 3 \ 4 \ 5)\rangle \leq S_5$ and 
    $|H| = 5$. By the tower law: 
    \[
        [K : \mathbb Q] = 24
    \]
\end{proof}
\begin{proof}[Proof of (c)]
    We need a subgroup of order $12$. 
    Take $H := A_4$.

    Let $L = E^{A_4}$. Then 
    \[
        [L :\mathbb Q] = [S_5 : S_4] = 10.
    \]
\end{proof}
\begin{proof}[Proof of (d)]
    Suppose there exists 
    $M \subset E$ so that 
    $[M : \mathbb Q] = 7$. Then 
    $M / \mathbb Q$ is Galois and the 
    Galois group $H$ satisfies 
    $H \leq G$. Then 
    \[
        120 = |G| = [G : H] \cdot 7,
    \]
    meaning that $7$ would divide 
    $120$, which it does not. 
    Therefore there does not exist such an 
    intermediary field. 
\end{proof}

\begin{prob}
    Let 
    \[
        E = \mathbb Q(\sqrt 2, \sqrt 3, \sqrt 5).
    \]
    For $a,b,c \in \{0,1\}$, every element of $E$ 
    can be written uniquely as a $\mathbb Q$-linear 
    combination of $(\sqrt 2)^a, (\sqrt 3)^b, (\sqrt 5)^c$. 
    The basis is 
    \[
        \{1,\sqrt 2, \sqrt 3, \sqrt 5, \sqrt 6, \sqrt 10, \sqrt 15, \sqrt 30\}.
    \]
    Define automorphisms $\sigma_2,\sigma_3,\sigma_5\in 
    \operatorname{Gal}(E/\mathbb Q)$ by 
    \begin{align*}
        \sigma_2(\sqrt2)=-\sqrt2,\qquad \sigma_2(\sqrt3)=\sqrt3,\qquad \sigma_2(\sqrt5)=\sqrt5, \\ 
        \sigma_3(\sqrt2)=\sqrt2,\qquad \sigma_3(\sqrt3)=-\sqrt3,\qquad \sigma_3(\sqrt5)=\sqrt5, \\ 
        \sigma_5(\sqrt2)=\sqrt2,\qquad \sigma_5(\sqrt3)=\sqrt3,\qquad \sigma_5(\sqrt5)=-\sqrt5.
    \end{align*}

    \begin{enumerate}[label=\alph*)]
        \item Show that $[E:\mathbb Q] = 8$. 
            Then explain why 
            \[
                \operatorname{Gal}(E/\mathbb Q) 
                \cong C_2 \times \cong C_2 \times \cong C_2.
            \]
        \item Let 
            \[
                H = \langle \sigma_2 \rangle.
            \]
            Find the fixed field 
            \[
                E^H.
            \]
        \item Let 
            \[
                H = \langle \sigma_2, \sigma_3 \rangle. 
            \]
            Find $E^H$.
        \item Let 
            \[
                H = \langle \sigma_2 \sigma_3\rangle. 
            \]
            Find $E^H$.
        \item Let 
            \[
                H = \langle \sigma_2\sigma_3, \sigma_5 \rangle.
            \]
            Find $E^H$.
            Compute the degrees 
            \[
                [E:E^H], [E^H : \mathbb Q].
            \]
        \item Let $K = \mathbb Q(\sqrt 30)$. Find 
            $\operatorname{Gal}(E/K)$. 
        \item Let $K = \mathbb Q(\sqrt 2 + \sqrt 3)$. 
            Find $\operatorname{Gal}(E/K)$.
    \end{enumerate}
\end{prob}
\begin{proof}[Proof of (a)]
    The problem states that every element of $E$ can be uniquely written as a $\mathbb Q$-linear combination of the $8$-element basis 
    \[
        \{1, \sqrt{2}, \sqrt{3}, \sqrt{5}, \sqrt{6}, \sqrt{10}, \sqrt{15}, \sqrt{30}\}.
    \]
    Since the dimension of $E$ as a vector space over $\mathbb Q$ is equal to the cardinality of its basis, we immediately obtain $[E:\mathbb Q] = 8$.

    Because $E$ is the splitting field of the separable polynomial $(x^2-2)(x^2-3)(x^2-5)$ over $\mathbb Q$, the extension is Galois, which implies $|\operatorname{Gal}(E/\mathbb Q)| = [E:\mathbb Q] = 8$. 
    
    The given automorphisms $\sigma_2, \sigma_3, \sigma_5$ satisfy $\sigma_i^2 = \text{id}$ for each $i \in \{2, 3, 5\}$ because applying any of them twice negates the respective radical twice (e.g., $\sigma_2(\sigma_2(\sqrt{2})) = \sigma_2(-\sqrt{2}) = \sqrt{2}$). Furthermore, these automorphisms commute with each other. Since $\operatorname{Gal}(E/\mathbb Q)$ is an abelian group of order $8$ where every non-identity element has order $2$, it must be isomorphic to the elementary abelian group of order $8$:
    \[
        \operatorname{Gal}(E/\mathbb Q) \cong C_2 \times C_2 \times C_2.
    \]
\end{proof}
    
\begin{proof}[Solution to (b)]
    The subgroup $H$ is generated by $\sigma_2$, which sends $\sqrt{2} \mapsto -\sqrt{2}$ while fixing $\sqrt{3}$ and $\sqrt{5}$. An element $x \in E$ written in the standard basis is fixed by $\sigma_2$ if and only if the coefficients of all basis elements containing $\sqrt{2}$ (namely $\sqrt{2}, \sqrt{6}, \sqrt{10}, \sqrt{30}$) are zero. The remaining elements are generated by $\sqrt{3}$ and $\sqrt{5}$. Thus, the fixed field is:
    \[
        E^H = \mathbb Q(\sqrt{3}, \sqrt{5}).
    \]
\end{proof}

\begin{proof}[Solution to (c)]
    An element is fixed by $H$ if it is simultaneously fixed by $\sigma_2$ and $\sigma_3$. 
    Being fixed by $\sigma_2$ eliminates coefficients for base components containing $\sqrt{2}$, while being fixed by $\sigma_3$ eliminates coefficients for base components containing $\sqrt{3}$. The only remaining basis elements unaffected by both negations are $1$ and $\sqrt{5}$. Thus, the fixed field is:
    \[
        E^H = \mathbb Q(\sqrt{5}).
    \]
\end{proof}

\begin{proof}[Solution to (d)]
    The generator acts on the primary radicals as follows:
    \[
        \sigma_2\sigma_3(\sqrt{2}) = -\sqrt{2}, \quad \sigma_2\sigma_3(\sqrt{3}) = -\sqrt{3}, \quad \sigma_2\sigma_3(\sqrt{5}) = \sqrt{5}.
    \]
    Evaluating its effect on the product elements of the basis yields:
    \[
        \sigma_2\sigma_3(\sqrt{6}) = (-\sqrt{2})(-\sqrt{3}) = \sqrt{6}.
    \]
    Therefore, the basis elements explicitly fixed by $\sigma_2\sigma_3$ are $1$, $\sqrt{5}$, $\sqrt{6}$, and $\sqrt{30}$ (since $\sqrt{30} = \sqrt{5}\sqrt{6}$). Thus, the fixed field is:
    \[
        E^H = \mathbb Q(\sqrt{6}, \sqrt{5}).
    \]
\end{proof}

\begin{proof}[Solution to (e)]
    From part (d), we know $\sigma_2\sigma_3$ fixes $\mathbb Q(\sqrt{6}, \sqrt{5})$. Now we apply the second generator $\sigma_5$, which maps $\sqrt{5} \mapsto -\sqrt{5}$ and fixes $\sqrt{6}$. The only basis components from part (d) fixed by $\sigma_5$ are $1$ and $\sqrt{6}$. Thus, the fixed field is:
    \[
        E^H = \mathbb Q(\sqrt{6}).
    \]
    By the Fundamental Theorem of Galois Theory, the degree of the extension over the fixed field equals the order of the subgroup:
    \[
        [E:E^H] = |H| = 4.
    \]
    By the tower law property of extension degrees ($[E:\mathbb Q] = [E:E^H][E^H:\mathbb Q]$):
    \[
        [E^H:\mathbb Q] = \frac{8}{4} = 2.
    \]
\end{proof}

\begin{proof}[Solution to (f)]
        By Galois correspondence, $\operatorname{Gal}(E/K)$ consists of the automorphisms in $\operatorname{Gal}(E/\mathbb Q)$ that fix $K = \mathbb Q(\sqrt{30})$ elementwise. We test how the basic automorphisms act on $\sqrt{30}$:
        \begin{align*}
            \sigma_2(\sqrt{30}) &= (-\sqrt{2})\sqrt{3}\sqrt{5} = -\sqrt{30}, \\
            \sigma_3(\sqrt{30}) &= \sqrt{2}(-\sqrt{3})\sqrt{5} = -\sqrt{30}, \\
            \sigma_5(\sqrt{30}) &= \sqrt{2}\sqrt{3}(-\sqrt{5}) = -\sqrt{30}.
        \end{align*}
        Products of pairs of these maps will resolve the double negation: $\sigma_2\sigma_3(\sqrt{30}) = \sqrt{30}$, $\sigma_2\sigma_5(\sqrt{30}) = \sqrt{30}$, and $\sigma_3\sigma_5(\sqrt{30}) = \sqrt{30}$. Since $[E:K] = 4$, the subgroup must have order $4$. Thus:
        \[
            \operatorname{Gal}(E/K) = \{\text{id}, \sigma_2\sigma_3, \sigma_2\sigma_5, \sigma_3\sigma_5\}.
        \]
\end{proof}

\begin{proof}[Solution to (g)]
    We find which non-identity automorphisms fix $\sqrt{2} + \sqrt{3}$:
    \begin{align*}
        \sigma_2(\sqrt{2} + \sqrt{3}) &= -\sqrt{2} + \sqrt{3} \neq \sqrt{2} + \sqrt{3}, \\
        \sigma_3(\sqrt{2} + \sqrt{3}) &= \sqrt{2} - \sqrt{3} \neq \sqrt{2} + \sqrt{3}, \\
        \sigma_5(\sqrt{2} + \sqrt{3}) &= \sqrt{2} + \sqrt{3}.
    \end{align*}
    Since $\sigma_5$ fixes the generator of $K$, and $[E:K] = 2$ (as $K = \mathbb Q(\sqrt{2}, \sqrt{3})$ has degree $4$ over $\mathbb Q$), the Galois group must have order $2$. Therefore:
    \[
        \operatorname{Gal}(E/K) = \langle \sigma_5 \rangle = \{\text{id}, \sigma_5\}.
    \]
\end{proof}

\begin{prob}
    Let 
    \[
        K = \mathbb Q(\sqrt 2, \sqrt 5). 
    \]
    \begin{enumerate}
        \item Show that $K$ is the splitting field of 
            \[
                f(x) = (x^2 - 2)(x^2-5)
            \]
            over $\mathbb Q$, and prove that 
            \[
                [K:\mathbb Q] = 4. 
            \]
        \item Find $G = \operatorname{Gal}(K/\mathbb Q)$. 
            Describe it's elements explicitly by their 
            action on $\sqrt 2$ and $\sqrt 5$, and prove 
            that 
            \[
                G \cong C_2 \times C_2.
            \]
        \item List the subgroups of $G$.
        \item For each $H \leq G$, find the 
            fixed field $K^H$.
        \item Explain which intermediate extensions 
            are Galois over $\mathbb Q$.
    \end{enumerate}
\end{prob}
\begin{proof}[Proof of (a)]
    We begin by finding the roots of 
    $f(x)$ in $\overline{\mathbb Q}$.

    $x^2 - 2$ has zeros 
    $\pm \sqrt 2$ and $x^2 - 5$ has zeros 
    $\pm \sqrt 5$.
    These are contained in $K$. 
    Thus the splitting field 
    $L$ is a subset $L \subseteq K$. 

    Now, $K$ has $\{1,\sqrt 2, \sqrt 5, \sqrt{10}\}$ 
    as the basis as a vector space over 
    $\mathbb Q$. The first $3$ elements are 
    in the splitting field $L$ and $\sqrt{10} = \sqrt 2 \sqrt 5 \in L$ 
    so since every elements of $K$ is a linear combination 
    of these elements which are indeed contained in 
    the splitting field we have $K \subseteq L$. 
    Thus $K = L$.

    We claimed that $B = \{1,\sqrt 2, \sqrt 5, \sqrt{10}\}$ 
    is a basis for the extension and then 
    \[
        |B| = [K : \mathbb Q], 
    \]
    but we did this without proof. 
    We proceed with the tower law instead 
    and observe that as $x^2 - 2$ is the minimal 
    polynomial of $\sqrt 2$ over $\mathbb Q$ we have 
    \[
        [\mathbb Q(\sqrt 2) : \mathbb Q] = \deg (x^2 - 2) = 2. 
    \]
    We observe that $\sqrt 5 \not\in \mathbb Q(\sqrt 2)$, as 
    if this were the case we would have 
    $\sqrt 5 = a + b\sqrt 2$. If $b = 0$ then 
    we would have $\sqrt 5 \in \mathbb Q$ which 
    we do not. Working out the rest of the 
    cases we get contradictions where $\sqrt 2$ 
    is asserted to be rational which it is not. 
    Then $x^2 - 5$ is the minimal polynomial 
    of $\sqrt 5$ over $\mathbb Q(\sqrt 2)$ meaning 
    \[
        [K : \mathbb Q(\sqrt 2)] = 2 
    \]
    as before and using the tower law we get 
    \[
        [K : \mathbb Q] = 4.
    \]
\end{proof}

\begin{proof}[Solution and proof (b)]
    $G \leq S_4$ since $K$ is a splitting 
    field of some polynomial of degree $4$.  

    Specifically $G$ is the group of 
    automorphisms on $K$ which fix $\mathbb Q$ 
    and preserve 
    \[
        \psi(\sqrt 2)^2 - 2 = 0
    \]
    as with the minimal polynomial of 
    $\sqrt 5$. Thus we get 
    $\psi(\sqrt 2) = \pm \sqrt 2$ and 
    $\psi(\sqrt 5) = \pm \sqrt 5$ for 
    every $\psi \in G$.
    Thus, for each $\psi \in G$ we make 
    the choice of whether to flip the sign 
    of each root or not, meaning 
    $G \cong C_2 \times C_2$. 
\end{proof}

\begin{proof}[Solution of (c)]
    The nontrivial subgroups of $G$ are those which 
    one root is fixed, i.e. 
    \[
        \{0\} \times C_2, \ 
        C_2 \times \{0\}.
    \]
\end{proof}

\begin{proof}[Solution of (d)]
    For each $\psi \in G$ we have 
    \[
        \psi[\mathbb Q] = \mathbb Q.
    \]
    If $\psi$ fixes a basiselement $\alpha$ of 
    $K$, then $\psi$ will fix $\mathbb Q(\alpha)$. 
    Suppose $\alpha$ is a basiselement and 
    $\psi(\alpha) = \alpha$. Then 
    for each $a + b\alpha \in \mathbb Q(\alpha)$, we have 
    (since $\psi$ is a homomorphism), 
    \[
        \psi(a + b\alpha) 
        = \psi(a) + \psi(b)\psi(\alpha) 
        = a + b\alpha.
    \]
    Thus $\psi$ fixes $\mathbb Q(\alpha)$. 

    Then the fixed fields are 
    \[
        K, \ \mathbb Q(\sqrt 2), \
        \mathbb Q(\sqrt 5), \
        \mathbb Q(\sqrt{10}), \
        \mathbb Q.
    \]
\end{proof}
\begin{proof}[Solution of (e)]
    $G \cong C_2 \times C_2$ is abelian, hence 
    every subgroup is normal. Thus every 
    intermediate extension is Galois.
\end{proof}

\begin{prob}
    Work over $\mathbb F_2$. Let 
    \[
        f(x) = x^3 + x + 1, 
        g(x) = x^3 + x^2 + 1
    \]
    in $\mathbb F_2[x]$.

    Let 
    \[
        K = \mathbb F_2[x]/\langle f \rangle, \
        L = \mathbb F_2[x]/\langle g \rangle.
    \]

    Write 
    \[
        \alpha = x + \langle f \rangle \in K, 
        \ \beta = x + \langle g \rangle \in L.
    \]

    \begin{enumerate}[label=\alph*)]
        \item Show that $f$ and $g$ are irreducible over 
            $\mathbb F_2$.
            Conclude that they are fields. 
        \item Show that $K \cong L$ as fields. 
        \item Does the assignment $\alpha \mapsto \beta$ 
            necessarily define a $\mathbb F_2$-field homomorphism 
            \[
                K \to L? 
            \]
            Decide in this specific case. That is, check whether 
            there exists an $\mathbb F_2$-field homomorphism 
            $\phi : K \to L$ such that 
            \[
                \phi(\alpha) = \beta. 
            \]
        \item Find an explicit element of $L$ to which 
            $\alpha$ can be sent. In other words, find 
            $\gamma \in L$ such that 
            \[
                f(\gamma) = 0.
            \]
            Then explain why $\alpha \mapsto \gamma$ does 
            define a $\mathbb F_2$-field homomorphism $K \to L$.
        \item Let 
            \[
                \sigma : K \to K, \ \sigma(a) = a^2.
            \]
            Compute 
            \[
                \sigma(\alpha), \ \sigma^2(\alpha), \ 
                \sigma^3(\alpha).
            \]
            Then determine $\operatorname{Gal}(K/\mathbb F_2)$.
    \end{enumerate}
\end{prob}
\begin{proof}[Proof of (a)]
    Both $f,g$ have degree $3 \in \{2,3\}$ and are 
    thus irreducible over $\mathbb F_2$ if and only if they 
    have no zeros in $\mathbb F_2$.

    We quickly check that 
    \[
        f(0) = f(1) = g(0) = g(1) = 1.
    \]

    Since $\mathbb F_2$ is a field, the polynomial ring (in 
    one variable) is a 
    PID and thus a commutative ring. Therefore the 
    quotient ring is a field if and only if 
    the ideal we quotient by is maximal. 
    As $\mathbb F_2[x]$ is a PID we have that 
    the ideals are maximal if and only if they are prime. 
    Thus irreducibility implies the ideal is maximal which 
    in turn implies the quotient ring is a field. 

    Therefore we conclude that $\mathbb F_2[x]/\langle f \rangle$ 
    and $\mathbb F_2[x]/\langle g \rangle$ are fields. 
\end{proof}
\begin{proof}[Proof of (b)]
    We have that $f,g$ are irreducible and there is a standard 
    result which says that in this case 
    \[
        \left|\frac{\mathbb F_2[x]}{\langle f \rangle}\right|
        = |\mathbb F_2|^{\deg f} 
        = |\mathbb F_2|^{\deg g} = 
        \left|\frac{\mathbb F_2[x]}{\langle g \rangle}\right|
    \]
    and from here we can apply the classification of finite fields 
    to conclude that both quotient rings are then isomorphic.

    For the $|\mathbb F_2|^{\deg f}$ part we can elucidate the 
    claim as follows. We will do this for the quotient with 
    $\langle f \rangle$, and the approach is the same for 
    $g$. 

    Suppose $l(x) \in \mathbb F_2[x]$. Then 
    $l(x) = f(x)q(x) + r(x)$ where $0 \leq \deg r(x) < \deg f$. 
    Therefore each element in the quotient can be written 
    as a linear combination of $1, x, x^2$ with scalars in 
    $\mathbb F_2$, since the $f(x)q(x)$ term vanishes. 
    Thus there are $2^3$ choices. 
\end{proof}

\begin{proof}[Proof of (c)]
    Suppose $\phi : K \to L$ is a map such that 
    $\phi(\alpha) = \beta$. 
    The question is then if 
    \[
        f(\phi(\alpha)) = 0. 
    \]
    We have 
    \[
        f(\beta) = \beta^3 + \beta + 1,
    \]
    and
    \[
        \beta^3 = \beta^2 + 1 
    \]
    so the question is whether 
    \[
        \beta(\beta + 1) = 0.  
    \]
    This cannot be since $L$ is a field and 
    this would imply $\beta = 1$ or 
    $\beta = 0$, neither of which can be the case. 

    Thus $\alpha \mapsto \beta$ does not 
    constitute a $\mathbb F_2$-fixing homomorphism from $K \to L$.
     
\end{proof}
\begin{proof}[Solution of (d)]
    We need $\gamma \in L$ such that 
    $f(\gamma) = 0$.

    Let $\gamma := \beta^3$. 
    Then 
    \[
        f(\gamma) 
        = f(\beta^3) 
        = \beta^2 + (\beta^2 + 1) + 1 
        = 0.
    \]
    Define $\Phi : K \to L$ by 
    \[
        \Phi(h(\alpha)) = h(\gamma)
    \]
    for $h \in \mathbb F_2[x]$.
\end{proof}
\begin{proof}[Solution of (e)]
    We have 
    \[
        \sigma(\alpha) = \alpha^2.
    \]
    Then 
    \[
        \sigma^2(\alpha) = 
        \sigma(\alpha^2) = \alpha^4  
        = \alpha^2 + \alpha.
    \]
    Lastly 
    \[
        \sigma^3(\alpha) 
        = \sigma(\alpha^2 + \alpha) 
        = \alpha^4 + \alpha^2 
        = \alpha. 
    \]

    We always have that 
    $\mathbb F_{q^n} / \mathbb F_{q}$ is Galois 
    with $[\mathbb F_{q^n} : \mathbb F_q] = n$, 
    and furthermore the Galois group is 
    always the cyclic group of order $n$ generated 
    by the Frobenious automorphism $a \mapsto a^q$. 
    Thus 
    \[
        \operatorname{Gal}(K/\mathbb F_2) 
        = C_3 = \langle \sigma \rangle.
    \]
\end{proof}

\begin{prob}
    Let 
    \[
        G = \mathbb Z_{12} \times \mathbb Z_8.
    \]
    Let 
    \[
        H = \langle (4,2) \rangle \leq G.
    \]
    We want to determine $G/H$ up to isomorphism.

    \begin{enumerate}[label=\alph*)]
        \item Find the order of $(4,2) \in 
            \mathbb Z_{12} \times \mathbb Z_8$. 
            Conclude $|H|$ and compute 
            \[
                |G/H|.
            \]
        \item Show that in $G/H$, we have the relation 
            \[
                4 \overline{e_1} + 2 \overline{e_2} = 0, 
            \]
            where 
            \[
                \overline{e_1} = (1,0) + H, 
                \ \overline{e_2} = (0,1) + H.
            \]
            Record the original relations so 
            $G/H$ is generated by $\overline{e_1},\overline{e_2}$ 
            with $12\overline{e_1} = 0, 8 \overline{e_2} = 0$ and 
            the above relation.
        \item Prove that $G/H$ is not cyclic.
        \item Conclude that $G/H \cong \mathbb Z_4 \times 
            \mathbb Z_2$.
        \item Construct a surjective homomorphism 
            \[
                \phi : \mathbb Z_{12} \times 
                \mathbb Z_8 \to \mathbb Z_4 \times \mathbb Z_2 
            \]
            such that 
            \[
                \ker \phi = H.
            \]
            Use the first isomorphism theorem to conclude 
            the previous result.
    \end{enumerate}
\end{prob}

\begin{proof}[Solution of (a)]
    We have that $4$ has order $3$ in $\mathbb Z_{12}$ and 
    $2$ has the order $4$ in $\mathbb Z_8$.
    Then $|(4,2)| = \operatorname{lcm}(3,4) = 12$. 
    Thus $|H| = 12$. 

    Then, by Lagrange's theorem we have 
    \[
        |G/H| = 12 \cdot 8 / 12 = 8.
    \]
\end{proof}
\begin{proof}[Proof of (b)]
    Because LaTeX is annoying we will denote 
    $\overline{e_1} =: e_1$ and likewise for $\overline{e_2}$. 
    We have $H = \langle 4,2 \rangle$ so 
    $(4,2) H = H$, equivalently $(4,2) \in H$, equivalently 
    $(4,2)H = 0$ in the quotient. Then, since $(4,2)H = 
    4e_1 + 2e_2$ we see that the relation holds.
\end{proof}
\begin{proof}[Proof of (c)]
    If $G/H$ was cyclic it must contain some element of 
    order $|G/H| = 8$. Suppose $(a,b)H \in G/H$.  
    Then $(a,b) = ae_1 + be_2$. Then it is evident that 
    $(a,b)$ cannot have order $8$ since we already have the 
    relation given in (b).
\end{proof}
\begin{proof}[Proof of (d)]
    By the classification theorem for finite abelian groups we have 
    that the groups of order $|G/H| = 8$, up to isomorphism are 
    \[
        \prod_{i}\mathbb Z_{p_i^{e_i}}
    \]
    such that 
    \[
        \prod_{i} p_i^{e_i} = 8.
    \]
    These are $\mathbb Z_2^3, \ \mathbb Z_2 \times \mathbb Z_4, \ 
    \mathbb Z_8$.
    
    We have that $G/H$ has no element of order $8$ hence 
    $\mathbb Z_8$ is out of the question. 
    Furthermore, $(1,1)$ has order $4$ so $\mathbb Z_2^3$ is also 
    out of the question. Therefore 
    \[
        G \cong \mathbb Z_2 \times \mathbb Z_4.
    \]
\end{proof}
\begin{proof}[Proof of (e)]
    Let $\phi : \mathbb Z_{12} \times \mathbb Z_8 
    \to \mathbb Z_4 \times \mathbb Z_2$, by 
    \[
        \phi(a,b) = (a \mod 4, b \mod 2). 
    \]
    Then this map is clearly surjective and 
    \[
        \ker \phi = H. 
    \]
    Thus, by the first isomorphism theorem we have the desired 
    result.
\end{proof}

\begin{prob}
    Let $G = D_4$ be the symmetry group of the square. It has 
    $8$ elements: four rotations and four reflections. 

    We color the four vertices using black and white. 
    Two colorings are considered equivalent if you can obtain 
    one from the other by a symmetry of the square. 

    Let $X$ be the set of all such colorings. 
    Then $|X| = 16$.

    The group $D_4$ acts on $X$ by permuting the vertices. 

    We want to count the number of different colorings up to 
    the equivalence class created by this action.

    \begin{enumerate}[label=\alph*)]
        \item State Burnside's formula. 
        \item Label the vertices $1,2,3,4$ cyclically. 
            List the cycle of each symmetry of the square 
            as the permutation across vertices. 
        \item For a permutation $g$ of the vertices
            prove that the number of colorings fixed by $g$ is 
            $2^{c(g)}$ where $c(g)$ is the number of disjoint cycles 
            in the cycle decomposition of $g$.
        \item Use Burnside's formula to compute the number of  
            essentially different colorings of the square. 
        \item Now color the square using $m$ colors instead of $2$. 
            Find a formula for the number of essentially different 
            colorings in terms of $m$.
    \end{enumerate}
\end{prob}

\begin{proof}[Solution of (a)]
    Burnside's formula is 
    \[
        \#(X/G) = \frac{1}{|G|}\sum_{g\in G}|X^g|,
    \]
    where  
    \[
        X^g = \{x \in X : g \cdot x = x\}.
    \]
    $X^g$ is the set of $x \in X$ such that $g$ fixes $x$, meaning 
    it is like the opposite of $\operatorname{Stab}_G(x)$. Instead 
    of the $g \in G$ which leave $x$ untouched, it is the set 
    of $x$ which $g \in G$ leave untouched.
\end{proof}
\begin{proof}[Solution of (b)]
    Denote $r_{90}$ as rotating $90$ degrees, $r_{180}$ for 
    $180$, etc. 
    Then $r_{90}(1, 2, 3, 4) = (4,1,2,3)$ and 
    $r_{180}(1,2,3,4) = (3,4,1,2)$ and $r_{270}(1,2,3,4) 
    = (2,3,4,1)$. 

    For reflections we denote $s_{\uparrow}$ as a reflection 
    along vertical center-line, $s_{\rightarrow}$ as a reflection 
    along the horizontal, $s_{\nearrow}$ as one diagonal and 
    $s_{\searrow}$ as the other. 
    Then $s_{\uparrow}(1,2,3,4) = (3,4,2,1)$, 
    $s_{\rightarrow}(1,2,3,4) = (2,1,4,3)$, 
    $s_{\nearrow}(1,2,3,4) = (1,4,3,2)$ and 
    $s_{\searrow}(1,2,3,4) = (3,2,1,4)$.

    In cycle notation we get $r_{90} = (1 \ 2 \ 3 \ 4)$ and then 
    $r_{180} = r_{90}^2 = (1 \ 3)(2 \ 4)$ and similarly for $270$ we 
    get $r_{270} = r_{90} \circ r_{180} = (1 \ 4 \ 3 \ 2)$. 
    Then reflections are $s_{\uparrow} = (1 \ 4 \ 2 \ 3)$, 
    $s_{\rightarrow} = (1 \ 2)(3 \ 4)$, 
    $s_{\nearrow} = (2 \ 4)$ and $s_{\searrow} = (1 \ 3)$.
\end{proof}


\begin{prob}
    Let $R = \mathbb Z[x]$. 
    Define $\phi : R \to \mathbb F_3$ by 
    \[
        \phi(f) = f(1) \mod 3. 
    \]
    \begin{enumerate}
        \item Prove that $\phi$ is a ring 
            homomorphism and find 
            \[
                \ker \phi.
            \]
        \item Use the first isormorphism theorem 
            to identify the quotient 
            $\mathbb Z[x]/\langle 3, x-1 \rangle$.
            Decide whether $\langle 3, x-1 \rangle$ is 
            prime or maximal.
        \item[] Now let $J = \langle 3, x^2 + 1 \rangle 
            \subseteq \mathbb Z[x]$.
        \item Show that 
            \[
                \mathbb Z[x]/J \cong 
                \mathbb F_3[x]/\langle x^2 + 1 \rangle.
            \]
            Decide whether $J$ is prime or maximal.
        \item[] Let 
            \[
                K = \mathbb F_3[x]/\langle x^2 + 1\rangle,
            \]
            and let 
            \[
                \alpha = x + \langle x^2 + 1\rangle \in K.
            \]
        \item Show that $K$ is a field with $9$ 
            elements and basis $\{1,\alpha\}$ 
            over $\mathbb F_3$.
            Compute $\alpha^2, \alpha^3, \alpha^4$ and 
            determine the order of $\alpha$ in $K^\times$.
    \end{enumerate}
\end{prob}
\begin{proof}[Proof of a few problems]
    We have 
    \begin{align*}
        \phi(f+g) 
        &= (f+g)(1) \mod 3 \\ 
        &= (f(1) \mod 3 + g(1) \mod 3) \mod 3 \\ 
        &= \phi(f) + \phi(g). 
    \end{align*} 
    Similarly, 
    \begin{align*}
        \phi(fg) 
        &= f(1)g(1) \mod 3 \\ 
        &= f(1) \mod 3 \cdot g(1) \mod 3 \\ 
        &= \phi(f)\phi(g).
    \end{align*}
    We can also check $\phi(1) = 1$ is trivial.

    We want to find those $f \in R$ such 
    that $\phi(f) = 0$. Equivalently those 
    $f$ for which $f \equiv 0 \mod 3$ or 
    those $f$ which are divisible by $x-1$ 
    as $f(x) = g(x)(x-1) = 0$ for $x = 1$. 
    Thus 
    \[
        \langle 3, x-1\rangle \subseteq \ker \phi.
    \]
    Suppose $f \in \ker \phi$. 
    Then $\phi(f) \equiv 0 \mod 3$ so $f$ is 
    divisible by $3$ or $x-1$ so 
    \[
        \ker \phi \subseteq \langle 3, x-1\rangle. 
    \]

    Moving on... 

    \medskip 

    $\phi$ is surjective. Consider some 
    $n \in \mathbb F_3$. Then $f(x) = n$ 
    is sent to $n$ by $\phi$. 
    Therefore, by the first isomorphism theorem, 
    \[
        \frac{\mathbb Z[x]}{\langle 3, x-1\rangle} 
        \cong \mathbb F_3.
    \]
    Since $\mathbb F_3$ is a field we conclude 
    that $\langle 3,x-1 \rangle$ is maximal, because 
    a commutative ring (which $\mathbb Z[x]$ is) quotiented 
    by an ideal is a field if and only if 
    the ideal is maximal. 

    Moving on... 

    \medskip

    We will compare $\langle x^2 + 1 \rangle$ as an 
    ideal of $\mathbb R[x]$ vs. $\mathbb C[x]$.

    It is well known that 
    \[
        \frac{\mathbb R[x]}{\langle x^2 + 1 \rangle} 
        \cong \mathbb C 
    \]
    via the first isomorphism theorem applied to 
    the homomorphism $f \mapsto f(i)$.

    Over $\mathbb C$, $x^2 + 1$ splits 
    into $(x + i)(x-i)$ both of which 
    are polynomials in $\mathbb C[x]$ so 
    $\langle x^2 + 1 \rangle$ is not even 
    prime and thus $\mathbb C/\langle x^2 + 1 \rangle$ is 
    not an integral domain and hence ceirtainly not 
    isomorphic to $\mathbb C$.
\end{proof}

\begin{prob}
    Let $G,H$ be finite groups and let 
    $\phi : G \to H$ be a group homomorphism.

    Prove that 
    \[
        \frac{|G|}{|\ker \phi|} 
    \]
    divides $H$.

    Then prove that if $\gcd(|G|,|H|) = 1$, then 
    $\phi$ is the trivial homomorphism
\end{prob}
\begin{proof}
    By the first isomorphism theorem/fundamental theorem 
    of group homomorphisms, we have 
    \[
        \frac{|G|}{|\ker \phi|} 
        = \left|\frac{G}{\ker \phi}\right| 
        = |\operatorname{im} \phi|. 
    \]
    We note that $\operatorname{im}\phi$ is a subgroup 
    of $H$ (this can be proven by observing 
    that $\phi(e_G) = e_H$, $\phi(a), \phi(b) \in H 
    \Rightarrow \phi(a)\phi(b) = \phi(ab) \in H$, and so on), 
    so by Lagrange's theorem we have 
    that $|\operatorname{im} \phi|$ divides 
    $|H|$, which by the previous equalities, 
    gives the desired result.

    Assume $\gcd(|G|,|H|) = 1$.
    Then $|G|$ does not divide $|H|$, but 
    \[
        \frac{|G|}{|\ker \phi|}
    \]
    does divide $|H|$.
    Then $|G|$ does not split into factors 
    $a,b$ such that either divides $|H|$.
    Since  
    \[
        |G| = |\ker \phi| \cdot |\operatorname{im} \phi|
    \]
    and $|\operatorname{im} \phi|$ does divide $|H|$, it must 
    be the case that one of these factors is $1$ 
    and the other is $|G|$. We cannot not 
    have $|\operatorname{im} \phi| = |G|$ by assumption. Hence 
    $|\ker \phi| = |G|$ and $\phi$ is then the 
    trivial homomorphism $g \mapsto e_H$ for ever 
    $g \in G$.
\end{proof}

\begin{prob}
    List all finite abelian groups of 
    order $45 = 3^2 \cdot 5$ up to 
    isomorphism. Which are cyclic?
\end{prob}
\begin{proof}[Solution]
    By the classification theorem for finite 
    abelian groups, which is a special case of the 
    fundamental theorem of finitely generated abelian 
    groups, we have 
    \[
        G \cong \prod_{i}\mathbb Z_{p_i^{e_i}}
    \]
    such that 
    \[
        \prod_{i} p_i^{e_i} = 45.
    \]
    Then the unique ways to factor $45$ gives 
    us 
    \begin{align*}
        \mathbb Z_{3}^2 \times \mathbb Z_5 \\ 
        \mathbb Z_9 \times \mathbb Z_5. 
    \end{align*}
    For 
    \[
        H = \prod_i \mathbb Z_{q_i}
    \]
    we have that $H$ is cyclic only if 
    all $q_i$ are relatively prime. 
    Thus $\mathbb Z_9 \times \mathbb Z_5$ is 
    the only one that is cyclic, and 
    we can quickly see that it is 
    indeed generated by $(1,1)$. 
\end{proof}
\begin{prob}
    Let 
    \[
        \sigma = (1 \ 2 \ 5)(3 \ 4) \in S_5.
    \]
    Find the order of $\sigma$, decide whether 
    $\sigma \in A_5$ and decide whether 
    \[
        \langle \sigma \rangle \trianglelefteq 
        S_5.
    \]
\end{prob}
\begin{proof}
    Since $\sigma$ is the composition of 
    two disjoint cycles of lengths $2$ and 
    $3$ respectively we have 
    $\operatorname{ord}(\sigma) = 
    \operatorname{lcm}(2,3) = 6$.

    We have that $(1 \ 2 \ 5) = (1 \ 2)(2 \ 5)$ so 
    \[
        \sigma = (1 \ 2)(2 \ 5)(3 \ 4) \not\in A_5.
    \]
    If $\sigma$ was an element of the alternating group 
    we would automatically rule out normality. 

    We have $\sigma^3 = (3 \ 4) \in \langle \sigma \rangle$ 
    and we do the all-or-nothing test and ask whether 
    the subgroup contains all transpositions in $S_5$. 
    In $S_5$ there are $\binom{5}{2} = 10$ transpositions 
    and the only pure transposition in the subgroup is 
    the afforementioned $(3 \ 4)$ and the subgroup is 
    therefore not normal.
\end{proof}

\begin{prob}
    Let $D_4$ act on $X$ the set of colorings 
    of the vertices of a square using $m$ colors.   

    Use Burnside's formula to show that the number of 
    distinct (up to the equivalence class generated by 
    the action) colorings is 
    \[
        \frac{m^4 + 2m^3 + 3m^2 + 2m}{8}.
    \]
\end{prob}
\begin{proof}
Let the vertices of the square be labeled $1, 2, 3, 4$ in counterclockwise order. The dihedral group $D_4$ acts on the vertices as a subgroup of $S_4$, and this action induces an action on the set $X$ of all $m^4$ possible vertex colorings. 

By Burnside's Lemma, the number of distinct colorings up to symmetry (orbits) is given by:
\[
|X/D_4| = \frac{1}{|D_4|} \sum_{g \in D_4} |X^g|
\]
where $|D_4| = 8$, and $X^g$ denotes the set of colorings fixed by the group element $g$. A coloring is fixed by $g$ if and only if all vertices belonging to the same cycle in the disjoint cycle decomposition of $g$ are assigned the exact same color. If a permutation decomposes into $k$ disjoint cycles (including $1$-cycles), there are $m^k$ independent choices of colors, so $|X^g| = m^k$.

We analyze the $8$ elements of $D_4$ by their geometric actions and corresponding cycle structures:

\begin{enumerate}
    \item \textbf{The Identity ($e$):} 
    The identity element leaves all vertices fixed. Its cycle structure is $(1)(2)(3)(4)$, which yields $k = 4$ disjoint cycles. 
    \[ |X^e| = m^4 \]

    \item \textbf{Rotations by $\pm 90^\circ$ ($r, r^3$):} 
    A rotation by $90^\circ$ maps $1 \to 2 \to 3 \to 4 \to 1$. Both $r$ and its inverse $r^3$ form a single $4$-cycle: $(1\,2\,3\,4)$ and $(1\,4\,3\,2)$ respectively. Each has $k = 1$ cycle, requiring all 4 vertices to share a single color.
    \[ |X^r| = |X^{r^3}| = m^1 = m \]

    \item \textbf{Rotation by $180^\circ$ ($r^2$):} 
    This rotation maps opposite vertices to each other ($1 \leftrightarrow 3$ and $2 \leftrightarrow 4$). Its cycle structure is $(1\,3)(2\,4)$, giving $k = 2$ disjoint cycles. Vertices $1$ and $3$ must share a color, and vertices $2$ and $4$ must share a color.
    \[ |X^{r^2}| = m^2 \]

    \item \textbf{Reflections across Diagonal Axes ($s, r^2s$):} 
    There are $2$ reflections whose axes pass through opposite pairs of vertices. 
    \begin{itemize}
        \item One reflection fixes vertices $1$ and $3$ while swapping $2 \leftrightarrow 4$, yielding the cycle type $(1)(3)(2\,4)$.
        \item The other reflection fixes vertices $2$ and $4$ while swapping $1 \leftrightarrow 3$, yielding the cycle type $(2)(4)(1\,3)$.
    \end{itemize}
    Both elements yield $k = 3$ disjoint cycles.
    \[ |X^s| = |X^{r^2s}| = m^3 \]

    \item \textbf{Reflections across Midpoint Axes ($rs, r^3s$):} 
    There are $2$ reflections whose axes pass through the midpoints of opposite edges. These reflections leave no vertices fixed, instead swapping adjacent pairs of vertices simultaneously ($1 \leftrightarrow 2$ and $3 \leftrightarrow 4$, or $1 \leftrightarrow 4$ and $2 \leftrightarrow 3$). Their cycle structures are of the form $(\bullet\,\bullet)(\bullet\,\bullet)$, yielding $k = 2$ disjoint cycles.
    \[ |X^{rs}| = |X^{r^3s}| = m^2 \]
\end{enumerate}

Summing the sizes of the fixed-point sets across all $8$ elements yields:
\begin{align*}
\sum_{g \in D_4} |X^g| &= |X^e| + |X^r| + |X^{r^3}| + |X^{r^2}| + |X^s| + |X^{r^2s}| + |X^{rs}| + |X^{r^3s}| \\
&= m^4 + m + m + m^2 + m^3 + m^3 + m^2 + m^2 \\
&= m^4 + 2m^3 + 3m^2 + 2m
\end{align*}

Applying Burnside's formula by dividing this sum by $|D_4| = 8$, we find that the total number of distinct colorings up to symmetry is:
\[
|X/D_4| = \frac{m^4 + 2m^3 + 3m^2 + 2m}{8}
\]
\end{proof}

\begin{prob}
    Let $R = \mathbb Z \times \mathbb Z$.
    Find all units and zero divisors in 
    $R$.
    Decide whether the ideal 
    \[
        \mathbb Z \times \{0\}
    \]
    is prime and/or maximal.
\end{prob}
\begin{proof}
    The set of units in $\mathbb Z^2$ are 
    the $2$-combinations of units in 
    $\mathbb Z$, of which there are $1,-1$ 
    both of which are their own inverse. 
    Therefore the units of $\mathbb Z^2$ are 
    $(-1,-1), (-1,1), (1,-1), (1,1)$, all 
    with themselves as their multiplicative inverse. 

    Let $\phi : \mathbb Z^2 \to \mathbb Z$ by 
    \[
        \phi(a,b) = b. 
    \]
    Then $\phi$ is a surjection and $\ker \phi$ is 
    \[
        \ker \phi = \mathbb Z \times \{0\}.
    \]
    Then, by the first isomorphism theorem: 
    \[
        \frac{\mathbb Z^2}{\mathbb Z\times \{0\}} 
        \cong \mathbb Z. 
    \]
    $\mathbb Z$ is an integral domain so $\mathbb Z 
    \times \{0\}$ is prime, but not maximal as 
    $\mathbb Z$ is not a field. 
\end{proof}
\begin{prob}
    Define 
    \[
        \phi : \mathbb Z[x] \to \mathbb Z_5
    \]
    by 
    \[
        \phi(f) = f(3) \mod 5.  
    \]
    Prove that 
    \[
        \ker \phi = \langle 5, x-3\rangle. 
    \]
    Then identify $\mathbb Z[x]/\ker \phi$ and 
    decide whether $\ker \phi$ is prime or 
    maximal.
\end{prob}
\begin{proof}
    Suppose $f \in \ker \phi$. 
    Then $\phi (f) = 0$, equivalently 
    $f(3) \equiv 0$ in $\mathbb Z_5$.   
    Then $f(x) = q(x)(x-3)$ or 
    $f(x) = 5q(x)$ so 
    \[
        \ker \phi \subseteq \langle 5, x-3\rangle.
    \]
    Suppose $f(x) = q(x)(x-3)$. Then 
    $f(3) \equiv 0$, likewise if 
    $f(x) = 5q(x)$. Then 
    \[
        \langle 5,x-3 \rangle \subseteq \ker \phi.
    \]

    $\phi$ is surjective since, for $n \in \mathbb Z_5$ we have 
    $f(x) := n \Rightarrow \phi(n) \equiv n$. Thus by the first 
    isomorphism theorem: 
    \[
        \frac{\mathbb Z[x]}{\ker \phi} 
        \cong \mathbb Z_5.
    \]
\end{proof}

\begin{prob}
    Let 
    \[
        f(x) = x^3 + x + 1 \in \mathbb F_2[x].
    \]
    Show that $f$ is irreducible over $\mathbb F_2$. 
    Let 
    \[
        K = \mathbb F_2[x]/\langle f \rangle, \ 
        \alpha = x + \langle f \rangle. 
    \]
    Show that $K$ is a field with $8$ elements and 
    basis 
    \[
        \{1, \alpha, \alpha^2\} 
    \]
    over $\mathbb F_2$.

    Compute $\alpha^n$ for $3 \leq n \leq 7$.
\end{prob}
\begin{proof}
    $f$ is degree $3$ and is thus irreducible if and 
    only if it has no zeros in the ground-field. Checking: 
    \[
        f(0) = f(1) = 1
    \]
    we see that $f$ is irreducible over $\mathbb F_2$.

    Since $\mathbb F_2$ is a field we have that 
    the polynomial ring is a PID. Then, 
    $\langle f \rangle$ is maximal if and 
    only if it is prime. Thus, since all PID's 
    are commutative rings we get that 
    $K = \mathbb F_2[x]/\langle f \rangle$ is 
    a field if and only if $f$ is irreducible over 
    $\mathbb F_2$. Therefore $K$ is a field. 

    Now let $g(x) \in \mathbb F_2[x]$ so 
    $g(x) = q(x)f(x) + r(x)$ where 
    $0 \leq \deg r < \deg f$. 
    Then sinec $q(x)f(x)$ vanishes 
    over $\langle f \rangle$ we get that 
    $g$ is a linear combination of 
    $1,x,x^2$ with scalars in $\mathbb F_2$. 
    Thus there are $|\mathbb F_2|^{\deg f} = 2^3 = 8$ possibilities. 

    \begin{align*}
        \alpha^3 + \alpha + 1 
        &= 0 \\ 
        \alpha^3 
        &= \alpha + 1 \\ 
        \alpha^4 &= \alpha(\alpha + 1) = \alpha^2 + \alpha \\ 
        \alpha^5 &= \alpha(\alpha^2 + \alpha) = \alpha^2 + \alpha + 1 \\ 
        \alpha^6 &= \alpha^3 + \alpha^2 + \alpha = \alpha^2 + 1 \\ 
        \alpha^7 &= \alpha^3 + \alpha = \alpha + \alpha + 1 = 1.
    \end{align*}
\end{proof}

\begin{prob}
    Let $K$ be the field from the previous problem.
    Define $\sigma : K \to K$, $\sigma(a) = a^2$.

    Show that $\sigma$ is an automorphism. 
    Compute 
    \[
        \sigma(\alpha), \ \sigma^2(\alpha), \ \sigma^3(\alpha).
    \]
    Determine $\operatorname{Gal}(K/\mathbb F_2)$.
\end{prob}

\begin{proof}
    We have $K \cong \mathbb F_8$ since $|K| = 8$ and 
    $K$ is a field and finite field are classified 
    up to isomorphism. 

    Let $\psi : \mathbb F_8 \to \mathbb F_8$ by 
    \[
        \psi(a) = a^2. 
    \]
    Then suppose 
    \[
        \psi(a) = 0. 
    \]
    Then $a = 0$ so $\ker \psi = \{0\}$ and 
    is thus injective and since the domain 
    equals the codomain and they are finite bijectivity 
    follows. Now we show that this is indeed a homomorphism: 
    \begin{align*}
        \psi(ab) &= (ab)^2 \\ 
                 &= a^2b^2 \\ 
                 &= \psi(a)\psi(b) \\ 
        \psi(a+b)&= (a+b)^2 \\ 
                 &= a^2 + 2ab + b^2 \\ 
                 &\equiv a^2 + b^2 \\ 
                 &= \psi(a) + \psi(b) \\
        \psi(1)  &= 1^2 \\  &= 1.
    \end{align*}
    Let $\theta : \mathbb F_8 \to K$ be a field isomorphism, which 
    necessarily exists.
    Then $\sigma = \theta \circ \psi \circ \theta^{-1}$ is then 
    an automorphism on $K$.

    Now, 
    \begin{align*}
        \sigma(\alpha) 
        &= \alpha^2 \\ 
        \sigma^2(\alpha) \\ 
        &= \sigma(\alpha^2) \\ 
        &= \alpha^4 \\ 
        &= \alpha^2 + \alpha \\ 
        \sigma^3(\alpha) 
        &= \sigma(\alpha^4) \\ 
        &= \alpha^8 \\ 
        &= \alpha \cdot 1 \\ 
        &= \alpha. 
    \end{align*}
    Then $\operatorname{ord}(\sigma) = 3$. 

    For $\mathbb F_{q^n} / \mathbb F_q$ we have 
    that the extension is Galois with 
    $\operatorname{Gal}(K/\mathbb Q) = \mathbb Z/n\mathbb Z$ 
    generated by the Frobenious automorphism 
    $a \mapsto a^n$, and in fact 
    \[
        \langle \sigma \rangle = 
        \operatorname{Gal}(K/\mathbb F_2), 
    \]
    as we can see that $\sigma$ fixes $\{0,1\}$ and 
    $f(\sigma(\alpha)) = \sigma(f(\alpha)) = 0$. 
\end{proof}

\begin{prob}
    Let 
    \[
        K = \mathbb Q(\sqrt 2, \sqrt 7).
    \]
    Prove that 
    \[
        [K : \mathbb Q] = 4. 
    \]
\end{prob}
\begin{proof}
    We will first show that $\mathbb Q(\sqrt 2)$ is the 
    splitting field of $x^2 - 2$ over 
    $\mathbb Q$. 

    Let $L_{\sqrt 2}$ be the splitting field of $x^2 - 2$ over 
    $\mathbb Q$. Then 
    \[
        L_{\sqrt 2} \subseteq K
    \]
    since $K$ contains the roots $\pm \sqrt 2 \in L$ of $x^2 - 2$.

    Now since $\pm \sqrt 2 \in L_{\sqrt 2}$ and 
    $\mathbb Q\subseteq L_{\sqrt 2}$ we have that 
    $a + b\sqrt 2 \in L_{\sqrt 2}$ for any 
    $a,b \in \mathbb Q$ since $L_{\sqrt 2}$ is a field 
    and must therefore be closed under addition and multiplication 
    in a way which allows this. Therefeore 
    \[
        K \subseteq L_{\sqrt 2}. 
    \]

    For $\sqrt 2$, we have that $x^2 - 2$ is the minimal 
    polynomial and thus irreducible (can be shown with using 
    Eistenstein's criterion). Hence 
    \[
        [\mathbb Q(\sqrt 2) : \mathbb Q] = \deg(x^2 - 2) = 2. 
    \]

    We proceed now by showing that $\sqrt 2 \not \in \mathbb Q(\sqrt 2)$. 
    Sppose $\sqrt 7 \in \mathbb Q(\sqrt 2)$. Then 
    \begin{align*}
        \sqrt 7 &= a + b\sqrt 2 \\ 
        7 &= a^2 + 2ab\sqrt2 + 2b^2. 
    \end{align*}
    We cannot have $b = 0$ since then 
    \[
        7 = a^2, 
    \]
    but $7$ is not the square of any 
    element $\in \mathbb Q$. 
    If $a = 0$, then 
    \[
        7 = 2b^2 
    \]
    and 
    \[
        \frac{7}{2} = b^2,
    \]
    but once again $7/2$ is not the 
    square of any element $\in\mathbb Q$.
    Otherwise if both are non-zero then 
    \[
        \frac{7 - a^2 - 2b^2}{2ab} = \sqrt 2
    \]
    meaning $\sqrt 2 \in \mathbb Q$ which is 
    not the case. Therefore $\sqrt 7 \not \in 
    \mathbb Q(\sqrt 2)$, so $[K:\mathbb Q(\sqrt 2)] \neq 1$. 
    
    We use a similar argument as before that the minimal 
    polynomial of $\sqrt 7$ over $\mathbb Q(\sqrt 2)$ is 
    $x^2 - 7$ and therefore 
    \[
        [K : \mathbb Q(\sqrt 2)] = 2. 
    \]
    Using the tower law we get 
    \[
        [K : \mathbb Q] = 2 \cdot 2 = 4. 
    \]
\end{proof}
\begin{prob}
    Let  
    \[
        f(x) = x^3 - 5 \in \mathbb Q[x], 
    \]
    let 
    \[
        \alpha = \sqrt[3]5, 
    \]
    and let 
    \[
        \omega = e^{2i\pi/3}. 
    \]
    For the splitting field of $f$ over 
    $\mathbb Q$, compute it's degree over 
    $\mathbb Q$ and determine 
    \[
        \operatorname{Gal}(f/\mathbb Q).
    \]
\end{prob}
\begin{proof}
    The roots of $f$ over $\mathbb C$ are 
    \[
        \alpha, \omega \alpha, \omega^2 \alpha. 
    \]
    The splitting field is 
    $\mathbb Q(\alpha, \omega\alpha)$.

    Then $f$ is still the minimal polynomial of 
    $\alpha$ over $\mathbb Q$ so 
    \[
        [\mathbb Q(\alpha) : \mathbb Q] = 3. 
    \]

    We remark that $\omega \not\in \mathbb Q(\alpha)$ since 
    $\omega \in \mathbb C \supset \mathbb R \supset \mathbb Q(\alpha)$. 

    Now for $\omega$, the cube root of unity we compute 
    \[
        \prod_{d|n} (x^d - 1)^{\mu(n/d)}  
    \]
    which since $n = 3$ is prime simply gives 
    \[
        \sum_{n=0}^2 x^n = x^2 + x^1 + 1 
    \]
    as the minimal polynomial. 
    Therefore 
    \[
        [\mathbb Q(\alpha, \omega) : \mathbb Q(\alpha)]
        = 2. 
    \]
    Using the tower law we get 
    \[
        [\mathbb Q(\alpha,\omega): \mathbb Q] = 6. 
    \]
    The Galois group of a splitting field of 
    a degree $n$-polynomial is a subgroup of 
    $S_n$. The extension $f/\mathbb Q$ is Galois since 
    it is algebraic, normal (since all splitting fields are) and 
    sperable (since every extension of a characteristic $0$ field 
    which is algebraic is separable). 
    Then 
    \[
        [\mathbb Q(\alpha, \omega) : \mathbb Q] = 
        \operatorname{Gal}(f/\mathbb Q)
    \]
    so the Galois group is a subgroup of $S_3$ with 
    $6$ elements hence it is $S_3$.
\end{proof}

\begin{prob}
    Suppose $E/\mathbb Q$ is a finite Galois extension and 
    \[
        \operatorname{Gal}(E/\mathbb Q) \cong S_4.
    \]
    Prove that there exists an intermediate field $K$ with 
    \[
        [K : \mathbb Q] = 6. 
    \]
    Then decide whether $K$ is necessarily Galois 
    for your construction. 
\end{prob}
\begin{proof}
    For a subgroup $H \leq G := \operatorname{Gal}(E/\mathbb Q)$  
    we have that $[E:E^H] = |H|$. 
    We have $H:= \langle (1 \ 2 \ 3 \ 4) \leq S_4$ with 
    order $4 = 24/6$. Then since $E/\mathbb Q$ is 
    Galois we have
    \[
        [E : \mathbb Q] = |G| = 24. 
    \]
    Then 
    \[
        [E:E^{H}] = |H| = 4. 
    \]
    Then, by the tower law we have 
    \[
        [E^{H}:\mathbb Q] 
        = \frac{[E:\mathbb Q]}{[E:E^H]} = \frac{24}{4} = 6.
    \]
    Lastly $E^H/\mathbb Q$ is Galois if and only if 
    $H \trianglelefteq G$, which it is not. 
\end{proof}

\begin{prob}
    Let 
    \[
        G = \mathbb Z_{12} \times \mathbb Z_8
    \]
    and let 
    \[
        H = \langle (4,2) \rangle.
    \]
    Determine $G/H$ up to isomorphism.
\end{prob}

\begin{proof}
Let $G = \mathbb{Z}_{12} \times \mathbb{Z}_8$ and $H = \langle (4,2) \rangle$. To determine the structure of the quotient group $G/H$ using the First Isomorphism Theorem, we seek a surjective group homomorphism $\phi: \mathbb{Z}_{12} \times \mathbb{Z}_8 \to \mathbb{Z}_4 \times \mathbb{Z}_2$ such that $\ker(\phi) = H$.

First, let us analyze the cyclic subgroup $H$. By computing successive multiples of the generator $(4,2)$ in $G$, we find:
\begin{align*}
1 \cdot (4,2) &= (4,2) \\
2 \cdot (4,2) &= (8,4) \\
3 \cdot (4,2) &= (0,6) \\
4 \cdot (4,2) &= (4,0) \\
5 \cdot (4,2) &= (8,2) \\
6 \cdot (4,2) &= (0,4) \\
7 \cdot (4,2) &= (4,6) \\
8 \cdot (4,2) &= (8,0) \\
9 \cdot (4,2) &= (0,2) \\
10 \cdot (4,2) &= (4,4) \\
11 \cdot (4,2) &= (8,6) \\
12 \cdot (4,2) &= (0,0)
\end{align*}
Thus, $H$ is a subgroup of order $12$. By Lagrange's Theorem, the order of the quotient group must be:
\[
|G/H| = \frac{|G|}{|H|} = \frac{12 \times 8}{12} = 8
\]

We define a map $\phi: \mathbb{Z}_{12} \times \mathbb{Z}_8 \to \mathbb{Z}_4 \times \mathbb{Z}_2$ by:
\[
\phi(x, y) = (x - 2y \pmod{4}, \, y \pmod{2})
\]
We must check that $\phi$ satisfies the required properties:

\begin{enumerate}
    \item \textbf{Well-defined and Homomorphism:}
    The map is well-defined because if $x \equiv x' \pmod{12}$, then $x \equiv x' \pmod{4}$, and if $y \equiv y' \pmod{8}$, then $y \equiv y' \pmod{4}$ and $y \equiv y' \pmod{2}$. Furthermore, because reduction modulo $n$ and linear combinations are additive operations, it easily follows that for any $(x_1, y_1), (x_2, y_2) \in G$:
    \[
    \phi((x_1, y_1) + (x_2, y_2)) = \phi(x_1+x_2, y_1+y_2) = \phi(x_1, y_1) + \phi(x_2, y_2)
    \]
    Thus, $\phi$ is a group homomorphism.

    \item \textbf{Surjectivity:}
    Let $(a, b) \in \mathbb{Z}_4 \times \mathbb{Z}_2$. We can find a preimage $(x, y) \in G$ by setting $y = b \in \{0, 1\}$. Then we need $x - 2b \equiv a \pmod{4}$, which means we can choose $x = a + 2b \pmod{4}$. Since any value modulo $4$ can be trivially lifted to a valid residue modulo $12$, a preimage always exists, proving $\phi$ is surjective.

    \item \textbf{Kernel Verification:}
    By definition, $\ker(\phi) = \{ (x,y) \in G \mid \phi(x,y) = (0,0) \}$. This yields the system of congruences:
    \begin{align*}
    y &\equiv 0 \pmod{2} \implies y \in \{0, 2, 4, 6\} \\
    x - 2y &\equiv 0 \pmod{4} \implies x \equiv 2y \pmod{4}
    \end{align*}
    We check which elements $(x, y) \in \mathbb{Z}_{12} \times \mathbb{Z}_8$ satisfy this:
    \begin{itemize}
        \item If $y = 0$, then $x \equiv 0 \pmod{4} \implies x \in \{0, 4, 8\}$. This yields $(0,0), (4,0), (8,0)$.
        \item If $y = 2$, then $x \equiv 4 \equiv 0 \pmod{4} \implies x \in \{0, 4, 8\}$. This yields $(0,2), (4,2), (8,2)$.
        \item If $y = 4$, then $x \equiv 8 \equiv 0 \pmod{4} \implies x \in \{0, 4, 8\}$. This yields $(0,4), (4,4), (8,4)$.
        \item If $y = 6$, then $x \equiv 12 \equiv 0 \pmod{4} \implies x \in \{0, 4, 8\}$. This yields $(0,6), (4,6), (8,6)$.
    \end{itemize}
    Comparing this list of $12$ elements to our explicit generation of $H$, we see they are completely identical. Thus, $\ker(\phi) = H$.
\end{enumerate}

By the First Isomorphism Theorem, $G/\ker(\phi) \cong \text{Im}(\phi)$. Substituting our found structures yields:
\[
G/H \cong \mathbb{Z}_4 \times \mathbb{Z}_2
\]
\end{proof}

\end{document}
